### 2015-09-08 paper.tex
added paper skeleton and latex assets

\documentclass{ntmanuscript}

\usepackage[acronym,toc]{glossaries}
\newacronym{MIT}{MIT}{the Massachusetts Institute of Technology}
\newacronym{UW}{UW}{University of Wisconsin}
%\newacronym{<++>}{<++>}{<++>}
\makeglossaries

\usepackage[printwatermark]{xwatermark}
\usepackage{xcolor}
\usepackage{graphicx}

\newwatermark[allpages,color=black!5,angle=45,scale=6,xpos=-40,ypos=40]{DRAFT}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Facility and Time Step Discretization Effects in Fuel Cycle Simulation}

\author{Robert~W.~Carlsen, Paul~P.H.~Wilson}

% Institutes of the authors
\institute{University of Wisconsin - Madison, Department of Nuclear Engineering and Engineering Physics, Madison, WI 53706}

% Information concerning the person submitting the manuscript
\submitter{Robert W. Carlsen}

\submitteraddress{229 N Midvale Blvd Apt 1, Madison, WI 53705}
\submitteremail{rwcarlsen@gmail.com}

% No more than three keywords, though each can be a phrase
\keywords{nuclear fuel cycle, simulation, TODO:third}

\usepackage{cite} % order citations correctly

\usepackage{float} % used to help position figures with the H modifier

%%%%%%%%%%%%% stuff for syntax highlighting and code blocks %%%%%%%%%%%%%%%%
\usepackage{listings} % for highlighted code blocks
\usepackage{color}

% create listings pkg definition of diff language
\lstdefinelanguage{diff}{
  morecomment=[l][\color{blue}]{@@},     % group identifier
  morecomment=[l][\color{red}]-,         % deleted lines
  morecomment=[l][\color{green}]+,       % added lines
  morecomment=[l][\color{magenta}]{---}, % Diff header lines (must appear after +,-)
  morecomment=[l][\color{magenta}]{+++},
}
\definecolor{bkggray}{rgb}{0.93,0.93,0.93}
\lstset{
    backgroundcolor=\color{bkggray}
}
%%%%%%%%%%% end stuff for syntax highlighting and code blocks %%%%%%%%%%%%%%

\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{booktabs} % nice rules for tables
\usepackage{microtype} % if using PDF
\usepackage{xspace}
\usepackage{hyperref}
\newcommand{\units}[1] {\:\text{#1}}%
\newcommand{\SN}{S$_N$}%{S$_\text{N}$}%{$S_N$}%
\newcommand{\Cyclus}{\textsc{Cyclus}\xspace}%
\newcommand{\Cycamore}{\textsc{Cycamore}\xspace}%
\newcommand{\Class}[1] {\texttt{#1}}%
\newcommand{\TODO}[1] {{\color{red}\textbf{TODO: #1}}}%
\date{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\newpage
\begin{abstract}
As nuclear power expands, technical, economic, political, and environmental
analyses of nuclear fuel cycles by simulators increase in importance.  To date,
however, current tools are often fleet-based rather than discrete and privately
\end{abstract}

\section{Introduction}
\subsection{BLABLA}
\section{BLABLA}

\section{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ans}
\bibliography{refs}
\end{document}

### 2015-09-08 paper.tex
add figures and paste in content from dissertation

@@
 
+\usepackage{graphicx}
 \usepackage[printwatermark]{xwatermark}
 \usepackage{xcolor}
-\usepackage{graphicx}
 
@@
 \usepackage{tabularx}
-\usepackage{graphicx}
 \usepackage{booktabs} % nice rules for tables
@@
 \usepackage{hyperref}
+\usepackage{caption}
 \newcommand{\units}[1] {\:\text{#1}}%
@@
 \section{Introduction}
-\subsection{BLABLA}
-\section{BLABLA}
+
+One of the major motivations for the development of Cyclus was creating a
+simulation environment that is more expressive and less limited than the more
+traditional system dynamics based simulators. One of the defining differences
+between these two modeling paradigms involves the discretization of
+facilities. System dynamics simulators such as VISION simulate facility
+"fleets". In system dynamics fleet-based modeling:
+
+\begin{itemize}
+
+    \item groups of facilities (usually all facilities of a given type) are
+    treated as a single, aggregate entity.
+
+    \item resource flows are continuous (limited by time step discretization)
+
+\end{itemize}
+
+Cyclus' agent-based architecture in concert with its dynamic resource exchange
+has the ability to model fleet-based facilities.  However, Cyclus was designed
+with the intent of modeling facilities individually with discrete flows. The
+difference between these two modeling styles in terms of their affects on
+output of fuel cycle simulators in general is currently unknown. This
+analysis is designed to provide both quantitative and qualitative insight
+about how fleet-based continuous flow models differ from individual facility
+discrete flow models.
+
+Many fuel cycle simulators step through time in discrete steps.  The duration
+of a single time step can have significant impact on the dynamics of a
+simulation.  The time step duration can affect the magnitude of individual
+material transfers between facilities and correspondingly affects the noise
+level in facility material inventories.  It can also act as a minimum bound
+for outage durations. The effects of the time step duration will also be
+investigated together with fleet vs facility modeling.  Questions of
+interest include:
+
+\begin{itemize}
+
+    \item How well do fleet-based models with continuous material flow capture
+    optimal fuel cycle transitions compared to individual facility modeling?
+
+    \item What effects are the primary drivers for differences in fleet-based
+    and individual-based facility modeling?
+
+    \item Does the impact of time step duration changes differ in fleet-based
+    and individual-based facility modeling?
+
+\end{itemize}
+
+This work investigates in detail differences between fleet vs. individual
+reactor modeling and time step duration.
+
+\section{Fleet Modeling Overview}
+
+There are various low level phenomena that have potential to affect the larger
+outcome of simulations with respect to facility modeling decisions.  One
+potential difference caused by fleet-based modeling involves fuel sharing.
+Power production by a group of individual reactors for a fixed amount of fuel
+shortage may vary depending on how the existing fuel is distributed.  This is
+illustrated in Figure \ref{fig:fuel-sharing}.  Being short 3 batches might
+mean that one reactor has no batches and isn't operating or that 3 reactors
+are missing a single batch.  This will be referred to as the \emph{fuel
+sharing effect}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.70\textwidth]{exp2/fuel-sharing.pdf}
+    \caption[The fuel sharing effect]{
+        A simple diagram showing effects of constrained fuel supply
+        distribution choices. In both cases A and B, reactor 1 needs 3 new
+        fuel batches to operate and reactors 2 and 3 each need 1 fresh batch.
+        In case A, the two batches are given to reactors 2 and 3, resulting in
+        reactor 1 remaining offline. In case B, the available fuel batches are
+        given to reactor 1 resulting in all three reactors being offline.
+    }
+    \label{fig:fuel-sharing}
+\end{figure}
+
+Another difference between the fleet and individual modeling is the
+\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
+cores is short a half batch of fuel it will produce $\frac{5}{6}\cdot C$ of
+power for that time step where $C$ is the power capacity of a single reactor
+of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is short a
+half batch of fuel, then it will produce no power that time step.  Time step
+duration can also have this effect.  For both fleet and individual reactor
+modeling, power goes offline in for full time steps.  For example, if a
+3-batch fleet based reactor is short $\frac{1}{2}$ batch of fuel with a 3
+month time step, then it will produce $\frac{5}{6} \cdot 3 \text{mo} \cdot C $ of
+energy.  In contrast, a monthly time step might allow the same reactor to
+produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if there was
+actually a fuel shortage for only 1 month.
+
+Individually modeled reactors each have their own refueling cycles.  Depending
+on how the refueling cycles and outages are staggered, power production can
+vary significantly.  For example, if all reactors refuel at the same time,
+then all reactors would be offline refueling at the same time.  This is
+referred to as the \emph{cycle staggering effect} and does not occur in
+fleet-based facility models.  An example of poor cycle staggering is shown in
+Figure \ref{fig:sync-cycle}.  This effect is closely related to the
+\emph{inventory drawdown effect} described next.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/sync-cycle.eps}
+    \caption[The cycle staggering effect]{
+        In this scenario, reactors are deployed annualy with a refueling cycle
+        length (including outage) of 18 months.  So all reactors deployed
+        every 3rd year refuel together.  All initial 100 LWRs start out with
+        their cycles synced as well.  While it may look like separate curves,
+        this plot shows only a single time series.
+    }
+    \label{fig:sync-cycle}
+\end{figure}
+
+Another consequence of synchronized reactor cycles is fuel fabrication
+capacity idling for an entire cycle (e.g. 18 months), and then receiving
+requests for new batches for every reactor all at once.  The reactors draw out
+inventory from the fuel fabrication facility together in a large quantity of
+orders. In order to avoid running short of inventory for making fresh fuel,
+the fabrication facility must keep at enough material on-hand to satisfy all
+the reactors at once.  If the reactor cycles were staggered, then a much lower
+on-hand inventory would be sufficient to prevent fuel shortages. This is
+referred to as the \emph{inventory drawdown effect}.  Increasing time step
+duration can also cause similar effects as it reduces the time granularity
+over which refueling can occur, resulting in larger impulse drawdown on
+inventories. Another component of this effect is that incoming and outgoing
+material flows occur "simultaneously" on a given time step. Facilities do not
+know about incoming material when they make commitments for outgoing material.
+For longer time steps, larger incoming quanta of material are not available
+for making offers. In general longer time steps and individual reactor
+modeling create a need for larger floating inventory buffers in order to avoid
+temporary material shortages.
+
+There are also performance differences between the two modeling paradigms.  A
+single 200 year simulation with several hundred fleet-based reactors takes
+about 2 seconds to run on a typical desktop machine where the same simulation
+using individual reactor facilities takes about 20 seconds.  This order of
+magnitude difference is an important tradeoff to keep in mind when doing fuel
+cycle analysis in general, particularly when considering optimization or
+sensitivity studies.
+
+\subsection{Fleet Reactor Implementation}
+
+A fleet reactor facility was created for Cyclus.  It is possible to create
+fleet models for Cyclus that live anywhere along the spectrum from individual,
+discrete facilities to continuous flow fleet facilities.  The particular model
+created here falls more toward the aggregate, continuous flow side of the
+spectrum.
+
+The fleet reactor, although a single entity modeling many reactors, is
+actually designed to look like many individual reactors to the cyclus
+simulator kernel. This "trick" allows other plugins for managing facility
+deployments to transparently work with the fleet reactor without having to
+know anything about the fleet concept.  The fleet reactor has several
+important characteristics described below:
+
+\begin{itemize}
+
+    \item Reactor capacity is deployed and retired in single reactor units
+        (e.g. 1 GWe LWRs).
+
+    \item Material is discharged from the fleet's reactor cores continuously
+        (every time step) according to the following equation:
+
+        \begin{equation}
+            D =
+            \frac{B}{L} \cdot N_{r} \cdot \frac{C_{inv}}{C_{cap}}
+            \label{eqn:fleet-discharge}
+        \end{equation}
+
+        $D$ is the discharge rate (kg per time step), $B$ is the batch size
+        (kg) for a single reactor's batch, $L$ is the cycle length (including
+        refueling time) in time steps, $N_{r}$ is the number of
+        reactors in the fleet, $C_{inv}$ is the total amount of fuel in all
+        reactor cores (kg), and $C_{cap}$ is the total fuel capacity for all
+        reactor cores (kg).  The last fraction term accounts for any amount of
+        underfueling that may be occuring resulting in lower fuel burning
+        rates due to unfueled reactors.
+
+    \item Refueling occurs continuously with as much fresh fuel as is
+        available up to the total amount to fill all reactor cores in the
+        fleet.
+
+    \item Generated power is computed similarly to core discharge rates by
+        multiplying the total fleet power capacity by the same last term of
+        Equation \ref{eqn:fleet-discharge} to account for unfueled capacity in
+        the fleet.
+
+    \item When a reactor in the fleet is retired, a full core of material is
+        discharged - even if there is some fraction the fleet with unfueled
+        cores.
+
+\end{itemize}
+
+These characteristics result in a fleet reactor having perfect fuel sharing
+and perfect cycle staggering among the individual reactors it models.  First
+discharge of spent fuel from fleets begins immediately rather than after a
+complete refueling cycle. Fuel shortages also only cause proportional loss of
+capacity rather than whole-reactor quantized shutdown.
+
+\section{Scenario Description}
+
+The scenario starts with 100 LWRs and follows an exponential curve with a 1\%
+annual growth rate for 200 years.  The decommissioning of the initial 100
+reactors is staggered over years 15 to 55. Enrichment feeds LWRs with fresh
+fuel with infinite capacity. In year 35, fast reactors become available for
+deployment and no more thermal reactors are built.  LWR spent fuel separations
+begins in year 15 with 2000 MTHM/yr capacity and increases to 3000 MTHM/yr in
+year 25.  Fast reactor spent fuel separations and fuel fabrication have
+infinite capacity.  All spent fuel is stored/cooled for 7 years before it is
+available for separations. Figure \ref{fig:flow} shows the material flow
+relationships between the facilities.  Compositions used for the simulations
+were simple, containing primarily $^{238}U$, $^{235}U$, $^{239}Pu$,
+$^{241}Am$, and small quantities of a few lanthanides.
+
+Cyclus version ?.? (TODO:~update~version) was used for this analysis.
+Facility models used were the standard ones provided with Cyclus in the
+Cycamore library (also version ?.?).  A custom storage facility was used
+because Cycamore does not yet provide one.  A custom fleet-based reactor was
+also used for both reactor types rather than modeling facilities individually
+with the Cycamore reactor primarily for performance reasons. The fleet reactor
+has no sophisticated physics functionality and instead just transmutes all
+fresh fuel into a fixed spent fuel composition.  A templated Cyclus input file
+for the scenario is in Appendix TODO: exp1 input file in appendix. Additional
+parameters for the reactor types are shown in Table \ref{tab:reactors}.
+
+\vspace{2mm}
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ |r | c c | }
+        \hline                       
+            & LWR & FR \\
+        \hline                       
+        Lifetime (yr)         & 80 & 80 \\
+        Cycle length (months) & 18 & 14 \\
+        Batch size (kg)       & 29565 & 7490 \\
+        Batches per core      & 3 & 5 \\
+        \hline                       
+    \end{tabular}
+    \captionsetup{justification=centering}
+    \caption[Reactor facility parameters]{Reactor facility parameters.}
+    \label{tab:reactors}
+\end{table}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{exp2/flow.eps}
+    \caption[Material flow paths]{Potential material flow paths between facility types.}
+    \label{fig:flow}
+\end{figure}
+
+The described scenario together with a few perterbations result in four
+different cases for investigation:
+
+\begin{enumerate}
+
+    \item Monthly time steps with individual reactor modeling
+    \item Monthly time steps with fleet reactor modeling
+    \item Quarterly (3-month) time steps with individual reactor modeling
+    \item Quarterly (3-month) time steps with fleet reactor modeling
+
+\end{enumerate}
+
+A few other small changes to the scenario are also necessary.  For the cases
+using quarterly time steps with individual reactors, cycle and outage times
+must be adjusted to be a multiple of 3 months.  In order to make the
+comparison fair, the cases with monthly time steps are also adjusted to have
+identical cycle and outage times.  The invariants preserved with respect to
+reactor behavior are shown in Table \ref{tab:exp2-invar}.  Table
+\ref{tab:exp2-reactors} shows the computed/selected configuration for both
+reactor types.  In each of the four cases, the initial 100 LWRs are modified
+to have a zero length refueling outage with an explicit 900 MWe power
+capacity.  This was done to avoid having all the initial LWRs in the
+individual reactor cases refuel at the same time.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ |r | c c | }
+        \hline                       
+                                                                             & LWR      & SFR      \\
+        \hline                       
+                                                                             &          &          \\
+        Discharge Rate ($\frac{\text{kg} \cdot \text{HM}}{\text{month}}$)           & 1642.5   & 535      \\
+                                                                             &          &          \\
+        Burnup  ($\frac{\text{MWe} \cdot \text{month}}{\text{kg} \cdot \text{HM}}$) & 0.547945 & 0.672897 \\
+                                                                             &          &          \\
+        Effective Power  ($\text{MWe}$)                                      & 900      & 360      \\
+                                                                             &          &          \\
+        Core Size  ($\text{kg} \cdot \text{HM}$)                                    & 88695    & 40125    \\
+                                                                             &          &          \\
+        \hline                       
+    \end{tabular}
+    \captionsetup{justification=centering}
+    \caption[Reactor parameter invariants]{
+        Reactor configuration invariants maintained across all four
+        simulation cases.
+    }
+
+    \label{tab:exp2-invar}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular}{ |r | c c | c c | }
+        \hline                       
+                                          & \multicolumn{2}{c |}{LWR}       & \multicolumn{2}{c |}{SFR} \\
+                                          & Cases 1,3 & Cases 2,4 (fleet) & Cases 1,3 & Cases 2,4 (fleet)  \\
+        \hline                       
+        Cycle length (months)             & 15        & 18                & 12        & 15 \\
+        Refueling outage (months)         & 3         & 0                 & 3         & 0 \\
+        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 29,565            & 8,025     & 40,125 \\
+        Power capacity (MWe)              & 1,080     & 900               & 450       & 360 \\
+        \hline                       
+    \end{tabular}
+    \captionsetup{justification=centering}
+    \caption[Reactor parameters by case]{
+        Reactor configuration for each of the four simulation cases.
+    }
+
+    \label{tab:exp2-reactors}
+\end{table}
+
+The parameters for each case were also selected to keep capacity factors for
+individual reactors very roughly around 90\%.  The template input file for
+each of these cases is in Appendix TODO: exp2 input files in appendix.
+
+A longer 21-month deployment period is used because it provides better natural
+staggering for the reactor refueling cycles. The same deployment schedule is
+used for all four cases; power capacity is built along a curve starting at
+100GWe in year zero following the 1\% exponential growth curve.  This
+particular deployment schedule causes significant fuel shortages providing
+interesting features to investigate and compare.
+
+\section{Results}
+
+\subsection{Power Production}
+
+Figure \ref{fig:exp2-power} shows an overall view of the generated
+power over time for each of the four cases.  Figure
+\ref{fig:exp2-power-rel} normalizes the Figure
+\ref{fig:exp2-power} curves to the expected exponential power curve
+magnifying some interesting differences between the four cases.  For the first
+100 years, the four cases behave somewhat similarly, although the individually
+modeled reactors in cases 1 and 3 actually go online and offline for refueling
+causing more variance.  Around year 100, a fuel shortage begins and more
+significant differences between the four cases become apparent. This fuel
+shortage lasts until about year 140. The power generated in Case 4 is lower
+than in case 2 during the fuel shortage years. This is primarily due to the
+\emph{quantized shutdown effect}.  The longer 3-month time step in case 4
+results in reactor capacity going offline longer than necessary.
+
+During the initial 100 years, cases 1 and 3 have larger variance in power
+output than cases 2 and 4.  This is caused by minor refueling cycle
+synchronization.  The \emph{cycle staggering effect} becomes much more visible
+during and after the fuel shortage as evidenced by the larger swings in power
+generation from time step to time step.  Initially, the reactor cycles are
+staggered well.  During the shortage, several reactors that previously had
+staggered cycles are all offline together waiting for fuel.  On time steps
+where reactors retire (not just during the shortage), new deployments are made
+to replace them in addition to new deployments made to address power demand
+growth.  Because the initial reactors retire in waves somewhat close together,
+there are corresponding waves of deployment, and these waves are echoed every
+80 years (the reactor lifetime).  These waves of deployment cause waves in
+recycled fuel availability that cause many of the reactors offline during the
+shortage to receive fuel and come online together.  The net effect is that the
+fuel shortage degrades reactor cycle stagering overall.  As the simulation
+continues, however, new reactors continue to be deployed and reactors with
+synchornized cycles retire resulting in a gradual return to decent better
+staggering visible by the end of the 200 years.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/power.eps}
+    \caption[Generated power]{
+        Generated power for all four cases.  A fuel shortage occurs from about
+        year 100 to about year 140.  In cases 1 and 3 with individual reactor
+        modeling, refueling cycles become much more synchronized during and
+        after the fuel shortage resuting in large power fluctuations (the
+        \emph{cycle staggering effect}).
+    }
+    \label{fig:exp2-power}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/power-rel.eps}
+    \caption[Normalized power]{
+        Generated power is normalized to the expected 1\% exponential growth
+        curve for all four cases.  A fuel shortage occurs from about year 100
+        to about year 140.  The \emph{cycle staggering effect} can be seen as
+        divergence from 1.0 and is especially visible in cases 1 and 3 where
+        the shortage increases cycle synchronization significantly. The
+        \emph{quantized shutdown effect} can also be seen in the discrepency
+        between quarterly and monthly time steps for fleet reactor modeling
+        (cases 2 and 4) during the shortage; a longer time step causes more
+        reactor outage than necessary. The consistent jumping up and down for
+        the fleet reactor cases is due to the mis-alignment of reactors going
+        offline (multiples of 12 months) and the build period (21 months).
+    }
+    \label{fig:exp2-power-rel}
+\end{figure}
+
+\subsection{Fuel Shortage}
+
+Figure \ref{fig:exp2-unfueled} shows the fuel shortage more explicitly
+for each of the four cases. The individual reactor modeling cases result in
+more cumulative outage than corresponding fleet-based cases. The quarterly
+time step of cases 3 and 4 make the outage significantly more severe. Cases 1
+and 3 exhibit approximately twice as much offline power as their corresponding
+fleet based simulations in cases 2 and 4.  The monthly time step simulations
+also have roughly twice as much offline power as their corresponding quarterly
+cases.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/unfueled.eps}
+    \caption[Cumulative offline power due to fuel shortage]{
+        Cumulative offline power for reactors that had delayed cycle start due
+        to fuel shortage. The difference between individual and fleet based
+        modeling is primarily a result of the \emph{fuel sharing effect}. The
+        difference between monthly and quarterly time steps is primarily
+        caused by the \emph{inventory drawdown effect}.
+    }
+    \label{fig:exp2-unfueled}
+\end{figure}
+
+Quantifying imperfect fuel sharing is a bit difficult, but Figure
+\ref{fig:exp2-badshare} gives one way to approximate the \emph{fuel
+sharing effect}.  Fleet reactor modeling by design exhibits perfect fuel
+sharing by design (zero inefficiency) and so is not included in this figure.
+Every batch that is given to a reactor that ends up not being able to start it
+cycle at the scheduled time (due to fuel shortage) adds one to this cumulative
+for each time step the cycle start is delayed.
+
+One useful over-estimate of the fuel sharing inefficiency is to assume every
+one of these batches could have been given to a reactor enabling it to
+operate.  Multiplying each of these cumulative batches by reactor capacity
+(i.e. 450 MWe) provides a way to compare fuel sharing inefficiency with the
+overall fuel outage.  Correcting for capacity factor in is appropriate these
+cases because the outage is many reactor cycle lengths in duration. This
+results in a cumulative offline power caused by poor fuel sharing of roughly
+1,600 $\text{GWe} \cdot \text{months}$ and 5,400 $\text{GWe} \cdot
+\text{months}$ for cases 1 and 3 respectively.  These approximations are a
+significant portion of the difference between individual and fleet cases in
+Figure \ref{fig:exp2-unfueled} suggesting that most of the discrepency
+in cumulative outage between fleet and individual modeling is caused by
+inefficient fuel sharing in the individual reactor cases.
+
+The poor fuel sharing exhibited in the individual reactor cases would not
+occur if there were no new reactors being built during the shortage.
+Previously operating reactors only ever need one batch - receiving only a
+single batch allows them to operate.  Newly constructed reactors need multiple
+batches in order to begin operation, and if they receive less than a full core
+they cannot operate.  The real world doesn't necessarily optimize for fuel
+sharing efficiency very cleanly. There are other factors that can affect fuel
+sharing outcomes.  A few include:
+
+\begin{enumerate}
+
+    \item Multiple fuel quanta comprise a single batch (i.e. multiple
+        assemblies per batch). This functionality is natively supported in the
+        individual reactor model.
+
+    \item On-hand fresh fuel inventory maintained at reactors. This
+        functionality is natively supported in the individual reactor model.
+
+    \item Long-term fuel contracts between reactors operators and fuel
+        suppliers.
+
+\end{enumerate}
+
+Having multiple assemblies per batch (item 1 above) will further degrade fuel
+sharing efficiency.  Even reactors that are short only one batch could
+potentially receive some fuel and still be unable to operate.  In order to
+maintain optimal sharing, fresh fuel now needs to be sent in all-or-nothing
+multi-assembly quanta with preference for reactors that need fewer assemblies.
+Spare fresh fuel inventory (item 2 above) will have the global effect
+increasing total outage severity; on average more batches of fresh fuel will
+be sitting unused.  However, it can potentially reduce the frequency of
+outages for individual reactors in some cases.
+
+If they were to ever occur, real world fuel shortages would likely not result
+in optimal fuel sharing. However, modeling these the causes of suboptimal fuel
+sharing is probably best accomplished with a more direct, intentional approach
+rather than as a modeling artifact as seen here in cases 1 and 3.  One way to
+alleviate the poor fuel sharing is to modify the individual reactor model to
+lower the preference value for requests for fresh fuel depending on how many
+assemblies it needs to have a full core.  The more assemblies it needs, the
+lower it will set its request preference.  This will have the effect of
+allowing the resource exchange resolution in Cyclus to globally prefer sending
+fuel to reactors that need less to operate.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/badshare.eps}
+    \caption[Cumulative unnecessary idling fuel]{
+        Fuel sharing inefficiency approximated by the cumulative number of
+        fuel batches given to each reactor on each refueling period multiplied
+        by the number of time steps that reactor had to delay the start of its
+        next cycle (due to fuel shortage).  Fleet reactors have perfect fuel
+        sharing by design and so are not included here. The large difference
+        between cases 1 and 3 is mostly due to the \emph{inventory drawdown
+        effect}. 
+    }
+    \label{fig:exp2-badshare}
+\end{figure}
+
+\subsection{Inventory Drawdown}
+
+The large differences seen between monthly and quarterly time step cases in
+Figures \ref{fig:exp2-unfueled} and \ref{fig:exp2-badshare} are
+primarily caused by the \emph{inventory drawdown effect}.  Figure
+\ref{fig:exp2-puinv} helps to visualize this effect.  The black dots in
+the figure represent impulse flows of available Pu out of the fuel fab and
+into fast reactors.  Blue lines represent the flows into Pu inventory
+available for making fuel. The in flows are not shown for the individual
+reactor cases because they are somewhat messy and obscure otherinformation in
+the plots. The colored curves show Pu inventory available for making fast
+reactor fuel.  When the inventory curves meet and force down the flows, fuel
+shortages occur.
+
+Quarterly time steps have less frequent but larger impulse material flows.
+These larger flows, however, do not compensate for the larger withdrawals
+because facilities do not know about incoming inventory when they resolve
+their outflow for a particular time step.  This information lag effectively
+requires the floating Pu inventory to be maintained at a higher level.  This
+can be clearly seen in the case 4 plot which has a higher standing inventory
+than case 2 during the fuel shortage years.  The case 4 Pu inventory is unable
+to drop below approximately 100 tonnes, which indicates out-flows to be
+roughly equal to the Pu flow into inventory each time step during those years.
+This can be confirmed by the in-flow and inventory curves touching at those
+points.  Case 2 Pu inventory is similarly limited at about the 30 tonne level.
+As expected, the case 4 withdrawals are also three times higher case 2
+withdrawals.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/puinv.eps}
+    \caption[Pu inventory and flow]{
+        Pu inventory and flows for all four cases.  In-flows for cases 1 and 3
+        are omitted because they are very noisy and obscure useful other
+        information.  Larger time steps results in larger per time step
+        material flows. These larger flows mean more material is not available
+        for supplying to requesters.  This information lag causes more supply
+        constraints than occur with smaller time steps - illustrating the
+        \emph{inventory drawdown effect}.
+    }
+    \label{fig:exp2-puinv}
+\end{figure}
+
+Different modeling choices can have a significant impact on the outcome of a
+simulation.  With discrete reactor modeling many factors must be considered in
+order to ensure the integrity of conclusions.  Individual reactor outage
+modeling can be very useful for certain analyses, but it can reduce result
+quality if things such as cycle staggering are handled appropriately.  The
+fleet and discreet modeling done here is itself multi-dimensional.  If
+individual outages are not of interest, reactor power could be lowered and the
+outage reduced to zero duration making the capacity factor part of the
+reactor's power capacity.  The cycle staggering effects present when modeling
+reactors individually can provide valuable insight not possible with the
+fleet-based reactor modeling.  Starting reactors following shortage induced
+outages should not follow the natural pacing of fuel availability, otherwise
+refueling cycles become unsatisfactorily synchronized.
+
+Increasing the time steps increases facilities' standing inventory requirement
+for avoiding supply constraints.  Facilities that have unconstrained supply
+and very large inventory buffers, are not affected by the time step in this
+way.  Facilities can provide no more inventory than they can hold on a single
+time step.  So increasing time step duration for a facility with unconstrained
+supply generally means higher standing inventories are required in order to
+achieve equal throughput. Even when facilities do not have individual
+inventory or throughput limits, the time step can still have a significant
+effect. In recycle loops there is a constraint on how fast material is
+introduced through the loop (i.e. the discharge rate of spent fuel from
+reactors).  The \emph{inventory drawdown effect} can reduce the aggregate
+throughput of the recycling loop.
+
+The impact of modeling assumptions depends heavily on the metrics of interest.
+For example, someone interested in fuel-shortage driven reactor outages could
+observe a factor of four span among the four cases shown in the detail
+comparison.  However, someone more interested in the overall transition speed
+would not see as significant a discrepancy.
+
+\section{Conclusion}
 

### 2015-09-08 exp2/badshare.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+function querycmd {
+    cyan -db $1 -custom query.json badshare
+}
+
+function expand {
+    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
+}
+
+querycmd case1.sqlite  > tmp1.dat
+querycmd case3.sqlite | expand > tmp3.dat
+
+paste tmp1.dat tmp3.dat | awk 'NR==1 {print "Time\t\tCase1\t\tCase3"; next;} {print $1"\t\t"$2"\t\t"$4}' > badshare.dat
+
+rm tmp1.dat tmp3.dat
+

### 2015-09-08 exp2/power-constr.sh

@@ -1,0 +1,26 @@
+#!/bin/bash
+
+simdur=2400
+
+# number of months between facility deployments
+deploy_period=21
+
+# LWR capacity in MWe
+lwr_cap=900
+
+# number of reactors initially deployed
+n_initial=100
+
+# +/- fraction of power curve that defines min/max bounds
+errband=0.10
+
+# % annual growth
+growthrate=.01
+
+
+echo -n '"MinPower": ['
+seq $simdur | awk 'NR == 1 {mult=1} NR % '$deploy_period' == 1 {printf "%.10g, ", '$n_initial'*'$lwr_cap'*mult*(1-'$errband')} {mult *= (1+'$growthrate'/12)}' | sed 's/, *$//'
+echo '],'
+echo -n '"MaxPower": ['
+seq $simdur | awk 'NR == 1 {mult=1} NR % '$deploy_period' == 1 {printf "%.10g, ", '$n_initial'*'$lwr_cap'*mult*(1+'$errband')} {mult *= (1+'$growthrate'/12)}' | sed 's/, *$//'
+echo '],'

### 2015-09-08 exp2/power.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+function querycmd {
+    cyan -db $1 power
+}
+
+function expand {
+    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
+}
+
+querycmd case1.sqlite > tmp1.dat
+querycmd case2.sqlite > tmp2.dat
+querycmd case3.sqlite | expand > tmp3.dat # quarterly
+querycmd case4.sqlite | expand > tmp4.dat # quarterly
+
+paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > power.dat
+
+rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
+

### 2015-09-08 exp2/puflow.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+function querycmd {
+    cyan -db $1 flow -nucs pu239,pu240,pu241,pu242 -from fastfab -to fast_reactor
+}
+
+function expand {
+    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
+}
+
+querycmd case1.sqlite > tmp1.dat
+querycmd case2.sqlite > tmp2.dat
+querycmd case3.sqlite | expand > tmp3.dat
+querycmd case4.sqlite | expand > tmp4.dat
+
+paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > puflow.dat
+
+rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
+

### 2015-09-08 exp2/puflowin.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+function querycmd {
+    cyan -db $1 flow -nucs pu239,pu240,pu241,pu242 -to fastfab
+}
+
+function expand {
+    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
+}
+
+querycmd case1.sqlite > tmp1.dat
+querycmd case2.sqlite > tmp2.dat
+querycmd case3.sqlite | expand > tmp3.dat
+querycmd case4.sqlite | expand > tmp4.dat
+
+paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > puflowin.dat
+
+rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
+

### 2015-09-08 exp2/puinv.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+function querycmd {
+    cyan -db $1 inv -nucs pu239,pu240,pu241,pu242 fastfab
+}
+
+function expand {
+    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
+}
+
+querycmd case1.sqlite > tmp1.dat
+querycmd case2.sqlite > tmp2.dat
+querycmd case3.sqlite | expand > tmp3.dat
+querycmd case4.sqlite | expand > tmp4.dat
+
+paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > puinv.dat
+
+rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
+

### 2015-09-08 exp2/unfueled.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+function indivquerycmd {
+    cyan -db $1 -custom query.json unfueled
+}
+
+function fleetquerycmd {
+    cyan -db $1 -custom query.json unfueled-fleet
+}
+
+function expand {
+    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
+}
+
+# the individual reactor unfueled query returns a count of offline reactors -
+# convert this to MWe power
+function topower {
+    awk 'NR==1{print $0;next;} {print $1" "$2*450;}'
+}
+
+indivquerycmd case1.sqlite | topower > tmp1.dat
+fleetquerycmd case2.sqlite > tmp2.dat
+indivquerycmd case3.sqlite | expand | topower > tmp3.dat
+fleetquerycmd case4.sqlite | expand > tmp4.dat
+
+paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > unfueled.dat
+
+rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
+

### 2015-09-08 paper.tex
add publish (2-col ans trans style) target and tweak layout

@@
-\documentclass{ntmanuscript}
+\documentclass{style}
 
@@
 \usepackage{xcolor}
+\usepackage{placeins}
 
@@
 There are various low level phenomena that have potential to affect the larger
-outcome of simulations with respect to facility modeling decisions.  One
-potential difference caused by fleet-based modeling involves fuel sharing.
-Power production by a group of individual reactors for a fixed amount of fuel
-shortage may vary depending on how the existing fuel is distributed.  This is
-illustrated in Figure \ref{fig:fuel-sharing}.  Being short 3 batches might
-mean that one reactor has no batches and isn't operating or that 3 reactors
-are missing a single batch.  This will be referred to as the \emph{fuel
-sharing effect}.
+outcome of simulations with respect to facility modeling decisions.
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.70\textwidth]{exp2/fuel-sharing.pdf}
-    \caption[The fuel sharing effect]{
-        A simple diagram showing effects of constrained fuel supply
-        distribution choices. In both cases A and B, reactor 1 needs 3 new
-        fuel batches to operate and reactors 2 and 3 each need 1 fresh batch.
-        In case A, the two batches are given to reactors 2 and 3, resulting in
-        reactor 1 remaining offline. In case B, the available fuel batches are
-        given to reactor 1 resulting in all three reactors being offline.
-    }
-    \label{fig:fuel-sharing}
-\end{figure}
-
-Another difference between the fleet and individual modeling is the
+One difference between the fleet and individual modeling is the
 \emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
@@
 3-batch fleet based reactor is short $\frac{1}{2}$ batch of fuel with a 3
-month time step, then it will produce $\frac{5}{6} \cdot 3 \text{mo} \cdot C $ of
-energy.  In contrast, a monthly time step might allow the same reactor to
-produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if there was
-actually a fuel shortage for only 1 month.
+month time step, then it will produce $\frac{5}{6} \cdot 3 \text{mo} \cdot C $
+of energy.  In contrast, a monthly time step might allow the same reactor to
+produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if
+there was actually a fuel shortage for only 1 month.
 
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{exp2/sync-cycle.eps}
+    \includegraphics[width=1.0\columnwidth]{exp2/sync-cycle.eps}
     \caption[The cycle staggering effect]{
@@
     \label{fig:sync-cycle}
+\end{figure}
+
+Another difference between fleet-based and individual facility modeling
+involves fuel sharing.  Power production by a group of individual reactors for
+a fixed amount of fuel shortage may vary depending on how the existing fuel is
+distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  Being
+short 3 batches might mean that one reactor has no batches and isn't operating
+or that 3 reactors are missing a single batch.  This will be referred to as
+the \emph{fuel sharing effect}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\columnwidth]{exp2/fuel-sharing.pdf}
+    \caption[The fuel sharing effect]{
+        A simple diagram showing effects of constrained fuel supply
+        distribution choices. In both cases A and B, reactor 1 needs 3 new
+        fuel batches to operate and reactors 2 and 3 each need 1 fresh batch.
+        In case A, the two batches are given to reactors 2 and 3, resulting in
+        reactor 1 remaining offline. In case B, the available fuel batches are
+        given to reactor 1 resulting in all three reactors being offline.
+    }
+    \label{fig:fuel-sharing}
 \end{figure}
@@
 
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.47\textwidth]{exp2/flow.eps}
+    \caption[Material flow paths]{Potential material flow paths between facility types.}
+    \label{fig:flow}
+\end{figure}
+
 Cyclus version ?.? (TODO:~update~version) was used for this analysis.
@@
 \end{table}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{exp2/flow.eps}
-    \caption[Material flow paths]{Potential material flow paths between facility types.}
-    \label{fig:flow}
-\end{figure}
 
@@
 
-\begin{table}
+\begin{table*}
     \centering
@@
     \label{tab:exp2-reactors}
-\end{table}
+\end{table*}
 
@@
 
-\begin{figure}[h]
+\begin{figure*}[!h]
     \centering
-    \includegraphics[width=1.0\textwidth]{exp2/power.eps}
+    \includegraphics[width=0.8\textwidth]{exp2/power.eps}
     \caption[Generated power]{
@@
     \label{fig:exp2-power}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[h]
+\begin{figure*}[!h]
     \centering
-    \includegraphics[width=1.0\textwidth]{exp2/power-rel.eps}
+    \includegraphics[width=0.8\textwidth]{exp2/power-rel.eps}
     \caption[Normalized power]{
@@
     \label{fig:exp2-power-rel}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
     \centering
-    \includegraphics[width=1.0\textwidth]{exp2/unfueled.eps}
+    \includegraphics[width=1.0\columnwidth]{exp2/unfueled.eps}
     \caption[Cumulative offline power due to fuel shortage]{
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
     \centering
-    \includegraphics[width=1.0\textwidth]{exp2/badshare.eps}
+    \includegraphics[width=1.0\columnwidth]{exp2/badshare.eps}
     \caption[Cumulative unnecessary idling fuel]{
@@
 
-\begin{figure}[!h]
+\begin{figure*}[!h]
     \centering
@@
     \label{fig:exp2-puinv}
-\end{figure}
+\end{figure*}
 

### 2015-09-09 paper.tex
update refs, start fleshing out intro

@@
 
+%%%%%% acronyms %%%%%%
 \usepackage[acronym,toc]{glossaries}
@@
 
-\usepackage{graphicx}
 \usepackage[printwatermark]{xwatermark}
 \usepackage{xcolor}
-\usepackage{placeins}
-
 \newwatermark[allpages,color=black!5,angle=45,scale=6,xpos=-40,ypos=40]{DRAFT}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%% cover info %%%%%%%%%%%%%%%
 \title{Facility and Time Step Discretization Effects in Fuel Cycle Simulation}
-
 \author{Robert~W.~Carlsen, Paul~P.H.~Wilson}
 
-% Institutes of the authors
 \institute{University of Wisconsin - Madison, Department of Nuclear Engineering and Engineering Physics, Madison, WI 53706}
-
-% Information concerning the person submitting the manuscript
 \submitter{Robert W. Carlsen}
-
 \submitteraddress{229 N Midvale Blvd Apt 1, Madison, WI 53705}
@@
 
-\usepackage{cite} % order citations correctly
+%%%%%%%%% includes and commands %%%%%%%%%%
 
-\usepackage{float} % used to help position figures with the H modifier
-
-%%%%%%%%%%%%% stuff for syntax highlighting and code blocks %%%%%%%%%%%%%%%%
-\usepackage{listings} % for highlighted code blocks
-\usepackage{color}
-
-% create listings pkg definition of diff language
-\lstdefinelanguage{diff}{
-  morecomment=[l][\color{blue}]{@@},     % group identifier
-  morecomment=[l][\color{red}]-,         % deleted lines
-  morecomment=[l][\color{green}]+,       % added lines
-  morecomment=[l][\color{magenta}]{---}, % Diff header lines (must appear after +,-)
-  morecomment=[l][\color{magenta}]{+++},
-}
-\definecolor{bkggray}{rgb}{0.93,0.93,0.93}
-\lstset{
-    backgroundcolor=\color{bkggray}
-}
-%%%%%%%%%%% end stuff for syntax highlighting and code blocks %%%%%%%%%%%%%%
-
+\usepackage{graphicx}
+\usepackage{placeins}
 \usepackage{tabularx}
@@
 \usepackage{caption}
-\newcommand{\units}[1] {\:\text{#1}}%
-\newcommand{\SN}{S$_N$}%{S$_\text{N}$}%{$S_N$}%
+\usepackage{cite} % order citations correctly
+\usepackage{float} % used to help position figures with the H modifier
+
 \newcommand{\Cyclus}{\textsc{Cyclus}\xspace}%
 \newcommand{\Cycamore}{\textsc{Cycamore}\xspace}%
-\newcommand{\Class}[1] {\texttt{#1}}%
-\newcommand{\TODO}[1] {{\color{red}\textbf{TODO: #1}}}%
 \date{}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-One of the major motivations for the development of Cyclus was creating a
-simulation environment that is more expressive and less limited than the more
-traditional system dynamics based simulators. One of the defining differences
-between these two modeling paradigms involves the discretization of
-facilities. System dynamics simulators such as VISION simulate facility
-"fleets". In system dynamics fleet-based modeling:
+One of the major motivations for the development of the Cyclus fuel cycle
+simulator was creating a simulation environment that is more expressive than
+more traditional fuel cycle simulators.  Cyclus was designed with a
+plugin-style architecture that allows different models for facility types to
+be swapped in and out trivially.  As a Cyclus simulation walks through each
+time step, facilities engage in a market-like dynamic resource exchange (DRE).
+They broadcast requests for material and suppliers respond with bids. The DRE
+then creates a network flow problem that it optimizes to resolve which
+transactions actually occur.  The plugin architecture combined with the DRE
+creates a powerful and flexible environment for the natural comparison of
+different modeling paradigms.
+
+Traditionally, many fuel cycle simulators use a system dynamics approach
+\cite{forrester_industrial_1961} to modeling. These simulators model several
+"stocks" that have levels that change over time due to "flows" between them.
+The flows are determined by an equation that is a function of the state of the
+system (i.e.  the system's stocks and flows).  A modeled system is then solved
+in discrete or continuous time depending on the limitations of the software
+and constraints imposed by the model.  In system dynamics simulators, the
+number of stocks and potential flows between them is generally a static
+property of the modeled system, making it more difficult to swap different
+reactor models in and out of the same scenario without wiring them into the
+simulator manually.
+
+System dynamics based fuel cycle simulators generally use a collection of
+stocks to represent the state of groups of like facilities commonly referred
+to as "fleets".  As the simulation steps through time, the levels of stocks
+(e.g. reactor fresh fuel inventory, repository waste inventories, etc.) are
+adjusted according to the calculated flow values.  The VISION simulator
+\cite{jacobson_verifiable_2010} among others operates this way.  In
+fleet-based modeling:
 
@@
 has the ability to model fleet-based facilities.  However, Cyclus was designed
-with the intent of modeling facilities individually with discrete flows. The
+with the intent of modeling facilities individually with discrete flows
+enabling the investigation of interesting real-world effects such as
+competition, reactor cycle staggering, and individual reactor outages.  The
 difference between these two modeling styles in terms of their affects on
-output of fuel cycle simulators in general is currently unknown. This
-analysis is designed to provide both quantitative and qualitative insight
-about how fleet-based continuous flow models differ from individual facility
-discrete flow models.
+output of fuel cycle simulators in general is currently unknown. This analysis
+is designed to provide both quantitative and qualitative insight about how
+fleet-based continuous flow models differ from individual facility discrete
+flow models.
 
@@
 for outage durations. The effects of the time step duration will also be
-investigated together with fleet vs facility modeling.  Questions of
-interest include:
+investigated together with fleet vs facility modeling.  Questions of interest
+include:
 
@@
 This work investigates in detail differences between fleet vs. individual
-reactor modeling and time step duration.
+reactor modeling together with variations in time step duration.
 

### 2015-09-09 paper.tex
clean up fleet modeling discussion

@@
 \begin{abstract}
-As nuclear power expands, technical, economic, political, and environmental
-analyses of nuclear fuel cycles by simulators increase in importance.  To date,
-however, current tools are often fleet-based rather than discrete and privately
+TODO: Write this abstract.
 \end{abstract}
@@
 There are various low level phenomena that have potential to affect the larger
-outcome of simulations with respect to facility modeling decisions.
+outcome of simulations with respect to facility modeling decisions. Several
+effects are described below that result from fundamental differences between
+these two modeling paradigms.
 
-One difference between the fleet and individual modeling is the
+One difference between the fleet and individual facility modeling is the
 \emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
@@
 duration can also have this effect.  For both fleet and individual reactor
-modeling, power goes offline in for full time steps.  For example, if a
+modeling, power goes offline in full time step increments.  For example, if a
 3-batch fleet based reactor is short $\frac{1}{2}$ batch of fuel with a 3
@@
 produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if
-there was actually a fuel shortage for only 1 month.
+the fuel shortage actually ended after only 1 month.
 
@@
 vary significantly.  For example, if all reactors refuel at the same time,
-then all reactors would be offline refueling at the same time.  This is
-referred to as the \emph{cycle staggering effect} and does not occur in
-fleet-based facility models.  An example of poor cycle staggering is shown in
-Figure \ref{fig:sync-cycle}.  This effect is closely related to the
-\emph{inventory drawdown effect} described next.
+then all reactors would be offline at the same time.  This is referred to as
+the \emph{cycle staggering effect} and does not occur in fleet-based reactor
+models.  An example of poor cycle staggering is shown in Figure
+\ref{fig:sync-cycle}.  Poor staggering causes spikes of material supply and
+demand that can lead to suboptimal resource utilization.  A fuel fabrication
+facility might receive requests for new batches for every reactor all at once.
+The reactors draw out inventory from the fuel fabrication facility together in
+a large quantity of orders.  If the fabrication facility does not have enough
+on hand, some number of reactors will need to wait until enough fuel can be
+pumped through fabrication.  Avoiding such constraints would require the
+fabrication facility to maintain suboptimally large on-hand inventories as a
+contingency.
 
@@
 a fixed amount of fuel shortage may vary depending on how the existing fuel is
-distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  Being
-short 3 batches might mean that one reactor has no batches and isn't operating
-or that 3 reactors are missing a single batch.  This will be referred to as
-the \emph{fuel sharing effect}.
+distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  A system
+being short 3 batches might mean that one reactor has no batches and isn't
+operating or that 3 reactors are each missing a single batch.  This is
+referred to as the \emph{fuel sharing effect}.
 
@@
 
-Another consequence of synchronized reactor cycles is fuel fabrication
-capacity idling for an entire cycle (e.g. 18 months), and then receiving
-requests for new batches for every reactor all at once.  The reactors draw out
-inventory from the fuel fabrication facility together in a large quantity of
-orders. In order to avoid running short of inventory for making fresh fuel,
-the fabrication facility must keep at enough material on-hand to satisfy all
-the reactors at once.  If the reactor cycles were staggered, then a much lower
-on-hand inventory would be sufficient to prevent fuel shortages. This is
-referred to as the \emph{inventory drawdown effect}.  Increasing time step
-duration can also cause similar effects as it reduces the time granularity
-over which refueling can occur, resulting in larger impulse drawdown on
-inventories. Another component of this effect is that incoming and outgoing
-material flows occur "simultaneously" on a given time step. Facilities do not
-know about incoming material when they make commitments for outgoing material.
-For longer time steps, larger incoming quanta of material are not available
-for making offers. In general longer time steps and individual reactor
-modeling create a need for larger floating inventory buffers in order to avoid
-temporary material shortages.
+One important phenomenon related to time step duration is the \emph{inventory
+drawdown effect}.  Increasing time step duration reduces the frequency over
+which refueling can occur, resulting in larger impulse drawdown on
+inventories. This by itself is not problematic because this is balanced by
+larger inventory top-up quantities.  Another component of this effect is the
+simultaneity of incoming and outgoing material flows on a given time step.
+Facilities do not know about incoming material when they make commitments for
+outgoing material.  For longer time steps, larger incoming quanta of material
+are not available for making offers. In general longer time steps and
+individual reactor modeling create a need for larger floating inventory
+buffers in order to avoid temporary material shortages.
 

### 2015-09-11 exp2/badshare.sh
consolidate analysis into single script. Add no fuel shortage build schedule and a bit of analysis

@@ -1,17 +1,0 @@
-#!/bin/bash
-
-function querycmd {
-    cyan -db $1 -custom query.json badshare
-}
-
-function expand {
-    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
-}
-
-querycmd case1.sqlite  > tmp1.dat
-querycmd case3.sqlite | expand > tmp3.dat
-
-paste tmp1.dat tmp3.dat | awk 'NR==1 {print "Time\t\tCase1\t\tCase3"; next;} {print $1"\t\t"$2"\t\t"$4}' > badshare.dat
-
-rm tmp1.dat tmp3.dat
-

### 2015-09-11 exp2/gendata.sh

@@ -1,0 +1,84 @@
+#!/bin/bash
+
+function badshare {
+    cyan -db $1 -custom query.json badshare
+}
+
+function puflow {
+    cyan -db $1 flow -nucs pu239,pu240,pu241,pu242 -from fastfab -to fast_reactor
+}
+
+function puflowin {
+    cyan -db $1 flow -nucs pu239,pu240,pu241,pu242 -to fastfab
+}
+
+function puinv {
+    cyan -db $1 inv -nucs pu239,pu240,pu241,pu242 fastfab
+}
+
+function power {
+    cyan -db $1 power
+}
+
+function indivunfueled {
+    cyan -db $1 -custom query.json unfueled
+}
+
+function fleetunfueled {
+    cyan -db $1 -custom query.json unfueled-fleet
+}
+function expand {
+    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
+}
+
+# the individual reactor unfueled query returns a count of offline reactors -
+# convert this to MWe power
+function topower {
+    awk 'NR==1{print $0;next;} {print $1" "$2*450;}'
+}
+
+function finalize {
+    paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > $1
+    rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
+}
+
+puflow case1.sqlite > tmp1.dat
+puflow case2.sqlite > tmp2.dat
+puflow case3.sqlite | expand > tmp3.dat
+puflow case4.sqlite | expand > tmp4.dat
+finalize puflow.dat
+
+puflowin case1.sqlite > tmp1.dat
+puflowin case2.sqlite > tmp2.dat
+puflowin case3.sqlite | expand > tmp3.dat
+puflowin case4.sqlite | expand > tmp4.dat
+finalize puflowin.dat
+
+puinv case1.sqlite > tmp1.dat
+puinv case2.sqlite > tmp2.dat
+puinv case3.sqlite | expand > tmp3.dat
+puinv case4.sqlite | expand > tmp4.dat
+finalize puinv.dat
+
+power case1.sqlite > tmp1.dat
+power case2.sqlite > tmp2.dat
[…]

### 2015-09-11 exp2/power.sh

@@ -1,19 +1,0 @@
-#!/bin/bash
-
-function querycmd {
-    cyan -db $1 power
-}
-
-function expand {
-    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
-}
-
-querycmd case1.sqlite > tmp1.dat
-querycmd case2.sqlite > tmp2.dat
-querycmd case3.sqlite | expand > tmp3.dat # quarterly
-querycmd case4.sqlite | expand > tmp4.dat # quarterly
-
-paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > power.dat
-
-rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
-

### 2015-09-11 exp2/puflow.sh

@@ -1,19 +1,0 @@
-#!/bin/bash
-
-function querycmd {
-    cyan -db $1 flow -nucs pu239,pu240,pu241,pu242 -from fastfab -to fast_reactor
-}
-
-function expand {
-    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
-}
-
-querycmd case1.sqlite > tmp1.dat
-querycmd case2.sqlite > tmp2.dat
-querycmd case3.sqlite | expand > tmp3.dat
-querycmd case4.sqlite | expand > tmp4.dat
-
-paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > puflow.dat
-
-rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
-

### 2015-09-11 exp2/puflowin.sh

@@ -1,19 +1,0 @@
-#!/bin/bash
-
-function querycmd {
-    cyan -db $1 flow -nucs pu239,pu240,pu241,pu242 -to fastfab
-}
-
-function expand {
-    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
-}
-
-querycmd case1.sqlite > tmp1.dat
-querycmd case2.sqlite > tmp2.dat
-querycmd case3.sqlite | expand > tmp3.dat
-querycmd case4.sqlite | expand > tmp4.dat
-
-paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > puflowin.dat
-
-rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
-

### 2015-09-11 exp2/puinv.sh

@@ -1,19 +1,0 @@
-#!/bin/bash
-
-function querycmd {
-    cyan -db $1 inv -nucs pu239,pu240,pu241,pu242 fastfab
-}
-
-function expand {
-    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
-}
-
-querycmd case1.sqlite > tmp1.dat
-querycmd case2.sqlite > tmp2.dat
-querycmd case3.sqlite | expand > tmp3.dat
-querycmd case4.sqlite | expand > tmp4.dat
-
-paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > puinv.dat
-
-rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
-

### 2015-09-11 exp2/runscen.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+# runscen takes 3 arguments - the first is the build schedule file name,
+# the second is the case to run (i.e. 1, 2, 3, or 4), and the third is the
+# file name for the generated output database.
+function runscen {
+    db=$(cat $1 | cycobj -sched -scen scen-case${2}.json | grep -o ' \+[^ ]*\.sqlite')
+    mv $db $3
+    rm -f *-*-*-*
+}
+
+runscen buildsched-monthly-noshortage.dat 1 noshortage-case1.sqlite
+runscen buildsched-monthly-noshortage.dat 2 noshortage-case2.sqlite
+runscen buildsched-quarterly-noshortage.dat 3 noshortage-case3.sqlite
+runscen buildsched-quarterly-noshortage.dat 4 noshortage-case4.sqlite

### 2015-09-11 exp2/unfueled.sh

@@ -1,29 +1,0 @@
-#!/bin/bash
-
-function indivquerycmd {
-    cyan -db $1 -custom query.json unfueled
-}
-
-function fleetquerycmd {
-    cyan -db $1 -custom query.json unfueled-fleet
-}
-
-function expand {
-    awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
-}
-
-# the individual reactor unfueled query returns a count of offline reactors -
-# convert this to MWe power
-function topower {
-    awk 'NR==1{print $0;next;} {print $1" "$2*450;}'
-}
-
-indivquerycmd case1.sqlite | topower > tmp1.dat
-fleetquerycmd case2.sqlite > tmp2.dat
-indivquerycmd case3.sqlite | expand | topower > tmp3.dat
-fleetquerycmd case4.sqlite | expand > tmp4.dat
-
-paste tmp1.dat tmp2.dat tmp3.dat tmp4.dat | awk 'NR==1{print "Time\t\tCase1\t\tCase2\t\tCase3\t\tCase4";next;} {print $1"\t\t"$2"\t\t"$4"\t\t"$6"\t\t"$8}' > unfueled.dat
-
-rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
-

### 2015-09-11 exp2/gendata.sh
added direct separated pu inv comparison

@@ -27,6 +27,7 @@ function indivunfueled {
 function fleetunfueled {
     cyan -db $1 -custom query.json unfueled-fleet
 }
+
 function expand {
     awk 'NR==1{print $0;next;} {print $1*3-2" "$2; print $1*3-1" "$2; print $1*3-0" "$2;}'
 }
@@ -66,12 +67,6 @@ power case3.sqlite | expand > tmp3.dat
 power case4.sqlite | expand > tmp4.dat
 finalize power.dat
 
-power noshortage-case1.sqlite > tmp1.dat
-power noshortage-case2.sqlite > tmp2.dat
-power noshortage-case3.sqlite | expand > tmp3.dat
-power noshortage-case4.sqlite | expand > tmp4.dat
-finalize power-noshort.dat
-
 indivunfueled case1.sqlite | topower > tmp1.dat
 fleetunfueled case2.sqlite > tmp2.dat
 indivunfueled case3.sqlite | expand | topower > tmp3.dat
@@ -82,3 +77,4 @@ badshare case1.sqlite  > tmp1.dat
 badshare case3.sqlite | expand > tmp3.dat
 paste tmp1.dat tmp3.dat | awk 'NR==1 {print "Time\t\tCase1\t\tCase3"; next;} {print $1"\t\t"$2"\t\t"$4}' > badshare.dat
 rm tmp1.dat tmp3.dat
+

### 2015-09-14 paper.tex
consolidate analysis into single script. Add no fuel shortage build schedule and a bit of analysis
added direct separated pu inv comparison
paragraph about pu inventory comparison fig

@@
 identical cycle and outage times.  The invariants preserved with respect to
-reactor behavior are shown in Table \ref{tab:exp2-invar}.  Table
-\ref{tab:exp2-reactors} shows the computed/selected configuration for both
+reactor behavior are shown in Table \ref{tab:invar}.  Table
+\ref{tab:reactors} shows the computed/selected configuration for both
 reactor types.  In each of the four cases, the initial 100 LWRs are modified
@@
 
-    \label{tab:exp2-invar}
+    \label{tab:invar}
 \end{table}
@@
 
-    \label{tab:exp2-reactors}
+    \label{tab:reactors}
 \end{table*}
@@
 
-Figure \ref{fig:exp2-power} shows an overall view of the generated
+Figure \ref{fig:power} shows an overall view of the generated
 power over time for each of the four cases.  Figure
-\ref{fig:exp2-power-rel} normalizes the Figure
-\ref{fig:exp2-power} curves to the expected exponential power curve
+\ref{fig:power-rel} normalizes the Figure
+\ref{fig:power} curves to the expected exponential power curve
 magnifying some interesting differences between the four cases.  For the first
@@
     }
-    \label{fig:exp2-power}
+    \label{fig:power}
 \end{figure*}
@@
     }
-    \label{fig:exp2-power-rel}
+    \label{fig:power-rel}
 \end{figure*}
@@
 
-Figure \ref{fig:exp2-unfueled} shows the fuel shortage more explicitly
+Figure \ref{fig:unfueled} shows the fuel shortage more explicitly
 for each of the four cases. The individual reactor modeling cases result in
@@
     }
-    \label{fig:exp2-unfueled}
+    \label{fig:unfueled}
 \end{figure}
@@
 Quantifying imperfect fuel sharing is a bit difficult, but Figure
-\ref{fig:exp2-badshare} gives one way to approximate the \emph{fuel
+\ref{fig:badshare} gives one way to approximate the \emph{fuel
 sharing effect}.  Fleet reactor modeling by design exhibits perfect fuel
@@
 significant portion of the difference between individual and fleet cases in
-Figure \ref{fig:exp2-unfueled} suggesting that most of the discrepency
+Figure \ref{fig:unfueled} suggesting that most of the discrepency
 in cumulative outage between fleet and individual modeling is caused by
@@
     }
-    \label{fig:exp2-badshare}
+    \label{fig:badshare}
 \end{figure}
@@
 The large differences seen between monthly and quarterly time step cases in
-Figures \ref{fig:exp2-unfueled} and \ref{fig:exp2-badshare} are
+Figures \ref{fig:unfueled} and \ref{fig:badshare} are
 primarily caused by the \emph{inventory drawdown effect}.  Figure
-\ref{fig:exp2-puinv} helps to visualize this effect.  The black dots in
+\ref{fig:puinv} helps to visualize this effect.  The black dots in
 the figure represent impulse flows of available Pu out of the fuel fab and
@@
     }
-    \label{fig:exp2-puinv}
+    \label{fig:puinv}
 \end{figure*}
@@
 comparison.  However, someone more interested in the overall transition speed
-would not see as significant a discrepancy.
+would not see as significant a discrepancy.  Also notable is that these
+different modeling choices may have varying levels of significance depending
+on the scenario/context they are operating in.  Figure \ref{fig:puinv-compare}
+shows the same inventory curves from Figure \ref{fig:puinv} superimposed.
+Before the fuel shortage begins, differences between each of the four cases
+are relatively minor.  However, after the shortage begins around year 100,
+separated Pu inventories vary much more significantly between the different
+cases.
+
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{exp2/puinv-compare.eps}
+    \caption[TODO]{
+        TODO
+    }
+    \label{fig:puinv-compare}
+\end{figure*}
 

### 2015-09-14 paper.tex
fill out conclusion a bit

@@
 
+Also notable is that different modeling choices (e.g. time step duration,
+facility discretization, etc.) may have varying levels of significance
+depending on the scenario/context they are operating in.  Figure
+\ref{fig:puinv-compare} shows the same inventory curves from Figure
+\ref{fig:puinv} superimposed.  Before the fuel shortage begins, differences
+between each of the four cases are relatively minor.  However, after the
+shortage begins around year 100, separated Pu inventories vary much more
+significantly between the different cases.
+
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{exp2/puinv-compare.eps}
+    \caption[TODO]{
+        TODO
+    }
+    \label{fig:puinv-compare}
+\end{figure*}
+
+\section{Conclusion}
+
 Different modeling choices can have a significant impact on the outcome of a
@@
 comparison.  However, someone more interested in the overall transition speed
-would not see as significant a discrepancy.  Also notable is that these
-different modeling choices may have varying levels of significance depending
-on the scenario/context they are operating in.  Figure \ref{fig:puinv-compare}
-shows the same inventory curves from Figure \ref{fig:puinv} superimposed.
-Before the fuel shortage begins, differences between each of the four cases
-are relatively minor.  However, after the shortage begins around year 100,
-separated Pu inventories vary much more significantly between the different
-cases.
+would not see as significant a discrepancy.
 
-\begin{figure*}[!h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{exp2/puinv-compare.eps}
-    \caption[TODO]{
-        TODO
-    }
-    \label{fig:puinv-compare}
-\end{figure*}
-
-\section{Conclusion}
+In an unconstrained supply setting, varying the time step duration or
+discretization of facility modeling have a relatively minor impact on the
+overall simulation results.  Differences become more significant in material
+constrained regimes.  Robustness in constrained regimes is especially useful
+in the context of sensitivity and optimization analysis.
 

### 2015-09-14 paper.tex
add reference to evaluation and screening report

@@
 \section{Scenario Description}
+
+The scenario used is a transition from a the current once-through light water
+reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle known as
+EG-23 in the Department of Energy's (DOE) Advanced Nuclear Fuel Cycle Options
+report \cite{wigeland_nuclear_2014}. Scenario and transition details were
+patterned after ongoing work by the DOE in their Fuel Cycle Options campaign.
 

### 2015-09-14 paper.tex
more introduction wrangling

@@
 
-One of the major motivations for the development of the Cyclus fuel cycle
-simulator was creating a simulation environment that is more expressive than
-more traditional fuel cycle simulators.  Cyclus was designed with a
-plugin-style architecture that allows different models for facility types to
-be swapped in and out trivially.  As a Cyclus simulation walks through each
-time step, facilities engage in a market-like dynamic resource exchange (DRE).
-They broadcast requests for material and suppliers respond with bids. The DRE
-then creates a network flow problem that it optimizes to resolve which
-transactions actually occur.  The plugin architecture combined with the DRE
-creates a powerful and flexible environment for the natural comparison of
-different modeling paradigms.
+\subsection{Motivation and Previous Studies}
 
-Traditionally, many fuel cycle simulators use a system dynamics approach
+Due to the diversity of fuel cycle simulators' modeling assumptions direct
+comparison and benchmarking can be difficult.  In 2012 the Organisation for
+Economic Co-operation and Development (OECD) completed a benchmark study
+\cite{oecd2012benchmark} that is perhaps one of the most complete published
+comparisons performed.  Despite this, however, various results from the
+different simulators were often significantly different due to different
+modeling decisions such as reprocessing strategies, refueling behavior,
+reactor end-of-life handling, etc.
+
+Traditionally, many fuel cycle simulators have used a system dynamics approach
 \cite{forrester_industrial_1961} to modeling. These simulators model several
@@
 The flows are determined by an equation that is a function of the state of the
-system (i.e.  the system's stocks and flows).  A modeled system is then solved
+system (i.e. the system's stocks).  A constructed system is then solved
 in discrete or continuous time depending on the limitations of the software
@@
 (e.g. reactor fresh fuel inventory, repository waste inventories, etc.) are
-adjusted according to the calculated flow values.  The VISION simulator
-\cite{jacobson_verifiable_2010} among others operates this way.  In
-fleet-based modeling:
+adjusted according to the calculated flow values.  Simulators that model
+facility fleets built on a system dynamics foundation include VISION
+\cite{jacobson_verifiable_2010}, DANESS \cite{van_den_durpel_daness_2009},
+DYMOND (predecessor to DANESS), and ORION among others.  In fleet-based
+modeling:
 
@@
 
-    \item resource flows are continuous (limited by time step discretization)
+    \item resource flows are often (approximately) continuous (limited by time
+        step discretization)
 
@@
 
-Cyclus' agent-based architecture in concert with its dynamic resource exchange
-has the ability to model fleet-based facilities.  However, Cyclus was designed
-with the intent of modeling facilities individually with discrete flows
-enabling the investigation of interesting real-world effects such as
-competition, reactor cycle staggering, and individual reactor outages.  The
-difference between these two modeling styles in terms of their affects on
-output of fuel cycle simulators in general is currently unknown. This analysis
-is designed to provide both quantitative and qualitative insight about how
-fleet-based continuous flow models differ from individual facility discrete
-flow models.
+This work investigates in detail differences between fleet vs. individual
+reactor modeling.  The difference between these two modeling styles in terms
+of their affects on output of fuel cycle simulators in general is currently
+unknown. This analysis is designed to provide both quantitative and
+qualitative insight about how fleet-based continuous flow models differ from
+individual facility discrete flow models.
 
@@
 for outage durations. The effects of the time step duration will also be
-investigated together with fleet vs facility modeling.  Questions of interest
-include:
+investigated together with fleet vs facility modeling.
+
+Questions of interest include:
 
@@
 
-This work investigates in detail differences between fleet vs. individual
-reactor modeling together with variations in time step duration.
+\subsection{CYCLUS}
+
+One of the major motivations for the development of the Cyclus fuel cycle
+simulator was creating a simulation environment that is more expressive than
+more traditional fuel cycle simulators.  Cyclus was designed with a
+plugin-style architecture that allows different models for facility types to
+be swapped in and out trivially.  As a Cyclus simulation walks through each
+time step, facilities engage in a market-like dynamic resource exchange (DRE).
+They broadcast requests for material and suppliers respond with bids. The DRE
+then creates a network flow problem that it optimizes to resolve which
+transactions actually occur.  The plugin architecture combined with the DRE
+creates a powerful and flexible environment for the natural comparison of
+different modeling paradigms.
+
+Cyclus' agent-based architecture in concert with its dynamic resource exchange
+has the ability to model fleet-based facilities.  However, Cyclus was designed
+to enable modeling facilities individually with discrete flows enabling the
+investigation of interesting real-world effects such as competition, reactor
+cycle staggering, and individual reactor outages.
 
@@
 inventories. This by itself is not problematic because this is balanced by
-larger inventory top-up quantities.  Another component of this effect is the
-simultaneity of incoming and outgoing material flows on a given time step.
-Facilities do not know about incoming material when they make commitments for
-outgoing material.  For longer time steps, larger incoming quanta of material
-are not available for making offers. In general longer time steps and
-individual reactor modeling create a need for larger floating inventory
-buffers in order to avoid temporary material shortages.
+correspondingly larger inventory top-up quantities.  The other component of
+this effect is the simultaneity of incoming and outgoing material flows on a
+given time step.  Within a particular time step facilities do not know about
+incoming material when they make commitments for outgoing material.  For
+longer time steps, larger incoming quanta of material are not available for
+making offers. In general longer time steps and individual reactor modeling
+create a need for larger floating inventory buffers in order to avoid
+material shortages.
 
@@
 cycle analysis in general, particularly when considering optimization or
-sensitivity studies.
+sensitivity studies where running hundreds of thousands of simulations may be
+desirable.
 
@@
 
-A fleet reactor facility was created for Cyclus.  It is possible to create
-fleet models for Cyclus that live anywhere along the spectrum from individual,
-discrete facilities to continuous flow fleet facilities.  The particular model
-created here falls more toward the aggregate, continuous flow side of the
-spectrum.
+A fleet reactor facility was created for Cyclus.  The flexibility of Cyclus
+makes it relatively straight forward to create fleet models for Cyclus that
+live anywhere along the spectrum from individual, discrete facilities to
+continuous flow fleet facilities.  In order to contrast well with individual,
+single-facility granularity modeling, the fleet model created here falls more
+toward the aggregate, continuous flow side of the spectrum.
 

### 2015-09-14 paper.tex
restructure sections

@@
 
-\subsection{Motivation and Previous Studies}
+\subsection{Motivation and Background}
 
@@
 
-\section{Fleet Modeling Overview}
+\section{Methodology}
+
+A single fuel cycle transition scenario is used with a few modeling variations
+resulting in four different cases:
+
+\begin{itemize}
+
+    \item \emph{Case 1}: Monthly time steps with individual reactor modeling
+    \item \emph{Case 2}: Monthly time steps with fleet reactor modeling
+    \item \emph{Case 3}: Quarterly (3-month) time steps with individual reactor modeling
+    \item \emph{Case 4}: Quarterly (3-month) time steps with fleet reactor modeling
+
+\end{itemize}
+
+Conducting this analysis involved developing some additional capability
+in/around Cyclus.  Three primary pieces to the methodology are:
+
+\begin{itemize}
+
+    \item A fleet-based reactor model.
+
+    \item A simulation scenario with both initial conditions and transition
+        details.
+
+    \item Theory and metrics for comparing differences between the four cases.
+
+\end{itemize}
+
+Because the standard package included with Cyclus does not include a
+fleet-based reactor model, one was necessarily developed. A simulation
+scenario was created that is roughly based on prior and ongoing work in the
+Department of Energy (DOE).  A few Phenomena related to the different modeling
+choices in each of the four cases are identified and used as a basis for
+comparison.  Each of these three things is described in more detail in the
+following sections.
+
+\subsection{Fleet Reactor Implementation}
+
+A fleet reactor facility was created for Cyclus.  The flexibility of Cyclus
+makes it relatively straight forward to create plug-in models for Cyclus that
+live anywhere along the spectrum from individual, discrete facilities to
+continuous flow fleet facilities.  In order to contrast well with individual,
+single-facility granularity modeling, the fleet model created here falls more
+toward the aggregate, continuous flow side of the spectrum.
+
+The fleet reactor, although a single entity modeling many reactors, is
+actually designed to look like many individual reactors to the cyclus
+simulator kernel. This "trick" allows other plugins for managing facility
+deployments to transparently work with the fleet reactor without having to
+know anything about the fleet concept.  The fleet reactor has several
+important characteristics described below:
+
+\begin{itemize}
+
+    \item Reactor capacity is deployed and retired in single reactor units
+        (e.g. 1 GWe LWRs).
+
+    \item Material is discharged from the fleet's reactor cores continuously
+        (every time step) according to the following equation:
+
+        \begin{equation}
+            D =
+            \frac{B}{L} \cdot N_{r} \cdot \frac{C_{inv}}{C_{cap}}
+            \label{eqn:fleet-discharge}
+        \end{equation}
+
+        $D$ is the discharge rate (kg per time step), $B$ is the batch size
+        (kg) for a single reactor's batch, $L$ is the cycle length (including
+        refueling time) in time steps, $N_{r}$ is the number of
+        reactors in the fleet, $C_{inv}$ is the total amount of fuel in all
+        reactor cores (kg), and $C_{cap}$ is the total fuel capacity for all
+        reactor cores (kg).  The last fraction term accounts for any amount of
+        underfueling that may be occuring resulting in lower fuel burning
+        rates due to unfueled reactors.
+
+    \item Refueling occurs continuously with as much fresh fuel as is
+        available up to the total amount to fill all reactor cores in the
+        fleet.
+
+    \item Generated power is computed similarly to core discharge rates by
+        multiplying the total fleet power capacity by the same last term of
+        Equation \ref{eqn:fleet-discharge} to account for unfueled capacity in
+        the fleet.
+
+    \item When a reactor in the fleet is retired, a full core of material is
+        discharged - even if there is some fraction the fleet with unfueled
+        cores.
+
+\end{itemize}
+
+These characteristics result in a fleet reactor having perfect fuel sharing
+and perfect cycle staggering among the individual reactors it models.  First
+discharge of spent fuel from fleets begins immediately rather than after a
+complete refueling cycle. Fuel shortages also only cause proportional loss of
+capacity rather than whole-reactor quantized shutdown.
+
+\subsection{Scenario Description}
+
+The scenario used is a transition from a the current once-through light water
+reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle known as
+EG-23 in the DOE's Advanced Nuclear Fuel Cycle Options report
+\cite{wigeland_nuclear_2014}. Scenario and transition details were patterned
+after ongoing work by the DOE in their Fuel Cycle Options campaign.
+
+The scenario starts with 100 LWRs and follows an exponential curve with a 1\%
+annual growth rate for 200 years.  The decommissioning of the initial 100
+reactors is staggered over years 15 to 55.  Enrichment feeds LWRs with fresh
+fuel with infinite capacity. In year 35, fast reactors become available for
+deployment and no more thermal reactors are built.  LWR spent fuel separations
+begins in year 15 with 2000 MTHM/yr capacity and increases to 3000 MTHM/yr in
+year 25.  Fast reactor spent fuel separations and fuel fabrication have
+infinite capacity.  All spent fuel is stored/cooled for 7 years before it is
+available for separations. Figure \ref{fig:flow} shows the material flow
+relationships between the facilities.  Compositions used for the simulations
+were simple, containing primarily $^{238}U$, $^{235}U$, $^{239}Pu$,
+$^{241}Am$, and small quantities of a few lanthanides.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.47\textwidth]{exp2/flow.eps}
+    \caption[Material flow paths]{Potential material flow paths between facility types.}
+    \label{fig:flow}
+\end{figure}
+
+Cyclus version ?.? (TODO:~update~version) was used for this analysis.
+Facility models used were the standard ones provided with Cyclus in the
+Cycamore library (also version ?.?).  A custom storage facility was used
+because Cycamore does not yet provide one.  The standard cycamore reactor was
+used for both reactor types (LWR and SFR) in cases 1 and 3, and the custom
+fleet-based reactor described earlier also used in cases 2 and 4.
+Additional parameters for the reactor types are shown in Table
+\ref{tab:reactors}.
+
+\vspace{2mm}
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ |r | c c | }
+        \hline                       
+            & LWR & FR \\
+        \hline                       
+        Lifetime (yr)         & 80 & 80 \\
+        Cycle length (months) & 18 & 14 \\
+        Batch size (kg)       & 29565 & 7490 \\
+        Batches per core      & 3 & 5 \\
+        \hline                       
+    \end{tabular}
+    \captionsetup{justification=centering}
+    \caption[Reactor facility parameters]{Reactor facility parameters. The cycle length includes the refueling outage.}
+    \label{tab:reactors}
+\end{table}
+
+The described scenario is used as closely as possible in each of the four
+cases.  However, a few other small case-dependent changes to the scenario
+relating to reactor configuration are necessary.  For the cases using
+quarterly time steps with individual reactors, cycle and outage times must be
+adjusted to be a multiple of 3 months.  In order to make the comparison more
+fair, the cases with monthly time steps are also adjusted to have identical
+cycle and outage times.  The invariants preserved with respect to reactor
+behavior are shown in Table \ref{tab:invar}.  Table \ref{tab:reactors} shows
+the computed/selected configuration for both reactor types for all four cases.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ |r | c c | }
+        \hline                       
+                                                                             & LWR      & SFR      \\
+        \hline                       
+                                                                             &          &          \\
+        Discharge Rate ($\frac{\text{kg} \cdot \text{HM}}{\text{month}}$)           & 1642.5   & 535      \\
+                                                                             &          &          \\
+        Burnup  ($\frac{\text{MWe} \cdot \text{month}}{\text{kg} \cdot \text{HM}}$) & 0.547945 & 0.672897 \\
+                                                                             &          &          \\
+        Effective Power  ($\text{MWe}$)                                      & 900      & 360      \\
+                                                                             &          &          \\
+        Core Size  ($\text{kg} \cdot \text{HM}$)                                    & 88695    & 40125    \\
+                                                                             &          &          \\
+        \hline                       
+    \end{tabular}
+    \captionsetup{justification=centering}
+    \caption[Reactor parameter invariants]{
+        Reactor configuration invariants maintained across all four
+        simulation cases.
+    }
+
+    \label{tab:invar}
+\end{table}
+
+\begin{table*}
+    \centering
+    \begin{tabular}{ |r | c c | c c | }
+        \hline                       
+                                          & \multicolumn{2}{c |}{LWR}       & \multicolumn{2}{c |}{SFR} \\
+                                          & Cases 1,3 & Cases 2,4 (fleet) & Cases 1,3 & Cases 2,4 (fleet)  \\
+        \hline                       
+        Cycle length (months)             & 15        & 18                & 12        & 15 \\
+        Refueling outage (months)         & 3         & 0                 & 3         & 0 \\
+        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 29,565            & 8,025     & 40,125 \\
+        Power capacity (MWe)              & 1,080     & 900               & 450       & 360 \\
+        \hline                       
+    \end{tabular}
+    \captionsetup{justification=centering}
+    \caption[Reactor parameters by case]{
+        Reactor configuration for each of the four simulation cases.
+    }
+
+    \label{tab:reactors}
+\end{table*}
+
+In each of the four cases, the initial 100 LWRs are modified to have a zero
+length refueling outage with an explicit 900 MWe power capacity.  This was
+done to avoid having all the initial LWRs in the individual reactor cases
+refuel at the same time.  The parameters for each case were also selected to
+keep capacity factors for individual reactors very roughly around 90\%.
+
+A longer 21-month deployment period is used because it provides better natural
+staggering for the reactor refueling cycles. The same deployment schedule is
+used for all four cases; power capacity is built along a curve starting at
+100GWe in year zero following the 1\% exponential growth curve.  This
+particular deployment schedule causes significant fuel shortages providing
+interesting features to investigate and compare.
+
+The input files and other assets used in the analysis for each of the four
+cases can be found in TODO: cite figshare tarball.
+
+\subsection{Modeling Effects}
 
@@
 desirable.
-
-\subsection{Fleet Reactor Implementation}
-
-A fleet reactor facility was created for Cyclus.  The flexibility of Cyclus
-makes it relatively straight forward to create fleet models for Cyclus that
-live anywhere along the spectrum from individual, discrete facilities to
-continuous flow fleet facilities.  In order to contrast well with individual,
-single-facility granularity modeling, the fleet model created here falls more
-toward the aggregate, continuous flow side of the spectrum.
-
-The fleet reactor, although a single entity modeling many reactors, is
-actually designed to look like many individual reactors to the cyclus
-simulator kernel. This "trick" allows other plugins for managing facility
-deployments to transparently work with the fleet reactor without having to
-know anything about the fleet concept.  The fleet reactor has several
-important characteristics described below:
-
-\begin{itemize}
-
-    \item Reactor capacity is deployed and retired in single reactor units
-        (e.g. 1 GWe LWRs).
-
-    \item Material is discharged from the fleet's reactor cores continuously
-        (every time step) according to the following equation:
-
-        \begin{equation}
-            D =
-            \frac{B}{L} \cdot N_{r} \cdot \frac{C_{inv}}{C_{cap}}
-            \label{eqn:fleet-discharge}
-        \end{equation}
-
-        $D$ is the discharge rate (kg per time step), $B$ is the batch size
-        (kg) for a single reactor's batch, $L$ is the cycle length (including
-        refueling time) in time steps, $N_{r}$ is the number of
-        reactors in the fleet, $C_{inv}$ is the total amount of fuel in all
-        reactor cores (kg), and $C_{cap}$ is the total fuel capacity for all
-        reactor cores (kg).  The last fraction term accounts for any amount of
-        underfueling that may be occuring resulting in lower fuel burning
-        rates due to unfueled reactors.
-
-    \item Refueling occurs continuously with as much fresh fuel as is
-        available up to the total amount to fill all reactor cores in the
-        fleet.
-
-    \item Generated power is computed similarly to core discharge rates by
-        multiplying the total fleet power capacity by the same last term of
-        Equation \ref{eqn:fleet-discharge} to account for unfueled capacity in
-        the fleet.
-
-    \item When a reactor in the fleet is retired, a full core of material is
-        discharged - even if there is some fraction the fleet with unfueled
-        cores.
-
-\end{itemize}
-
-These characteristics result in a fleet reactor having perfect fuel sharing
-and perfect cycle staggering among the individual reactors it models.  First
-discharge of spent fuel from fleets begins immediately rather than after a
-complete refueling cycle. Fuel shortages also only cause proportional loss of
-capacity rather than whole-reactor quantized shutdown.
-
-\section{Scenario Description}
-
-The scenario used is a transition from a the current once-through light water
-reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle known as
-EG-23 in the Department of Energy's (DOE) Advanced Nuclear Fuel Cycle Options
-report \cite{wigeland_nuclear_2014}. Scenario and transition details were
-patterned after ongoing work by the DOE in their Fuel Cycle Options campaign.
-
-The scenario starts with 100 LWRs and follows an exponential curve with a 1\%
-annual growth rate for 200 years.  The decommissioning of the initial 100
-reactors is staggered over years 15 to 55.  Enrichment feeds LWRs with fresh
-fuel with infinite capacity. In year 35, fast reactors become available for
-deployment and no more thermal reactors are built.  LWR spent fuel separations
-begins in year 15 with 2000 MTHM/yr capacity and increases to 3000 MTHM/yr in
-year 25.  Fast reactor spent fuel separations and fuel fabrication have
-infinite capacity.  All spent fuel is stored/cooled for 7 years before it is
-available for separations. Figure \ref{fig:flow} shows the material flow
-relationships between the facilities.  Compositions used for the simulations
-were simple, containing primarily $^{238}U$, $^{235}U$, $^{239}Pu$,
-$^{241}Am$, and small quantities of a few lanthanides.
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.47\textwidth]{exp2/flow.eps}
-    \caption[Material flow paths]{Potential material flow paths between facility types.}
-    \label{fig:flow}
-\end{figure}
-
-Cyclus version ?.? (TODO:~update~version) was used for this analysis.
-Facility models used were the standard ones provided with Cyclus in the
-Cycamore library (also version ?.?).  A custom storage facility was used
-because Cycamore does not yet provide one.  A custom fleet-based reactor was
-also used for both reactor types rather than modeling facilities individually
-with the Cycamore reactor primarily for performance reasons. The fleet reactor
-has no sophisticated physics functionality and instead just transmutes all
-fresh fuel into a fixed spent fuel composition.  A templated Cyclus input file
-for the scenario is in Appendix TODO: exp1 input file in appendix. Additional
-parameters for the reactor types are shown in Table \ref{tab:reactors}.
-
-\vspace{2mm}
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ |r | c c | }
-        \hline                       
-            & LWR & FR \\
-        \hline                       
-        Lifetime (yr)         & 80 & 80 \\
-        Cycle length (months) & 18 & 14 \\
-        Batch size (kg)       & 29565 & 7490 \\
-        Batches per core      & 3 & 5 \\
-        \hline                       
-    \end{tabular}
-    \captionsetup{justification=centering}
-    \caption[Reactor facility parameters]{Reactor facility parameters.}
-    \label{tab:reactors}
-\end{table}
-
-The described scenario together with a few perterbations result in four
-different cases for investigation:
-
-\begin{enumerate}
-
-    \item Monthly time steps with individual reactor modeling
-    \item Monthly time steps with fleet reactor modeling
-    \item Quarterly (3-month) time steps with individual reactor modeling
-    \item Quarterly (3-month) time steps with fleet reactor modeling
-
-\end{enumerate}
-
-A few other small changes to the scenario are also necessary.  For the cases
-using quarterly time steps with individual reactors, cycle and outage times
-must be adjusted to be a multiple of 3 months.  In order to make the
-comparison fair, the cases with monthly time steps are also adjusted to have
-identical cycle and outage times.  The invariants preserved with respect to
-reactor behavior are shown in Table \ref{tab:invar}.  Table
-\ref{tab:reactors} shows the computed/selected configuration for both
-reactor types.  In each of the four cases, the initial 100 LWRs are modified
-to have a zero length refueling outage with an explicit 900 MWe power
-capacity.  This was done to avoid having all the initial LWRs in the
-individual reactor cases refuel at the same time.
-
-\begin{table}
-    \centering
-    \begin{tabular}{ |r | c c | }
-        \hline                       
-                                                                             & LWR      & SFR      \\
-        \hline                       
-                                                                             &          &          \\
-        Discharge Rate ($\frac{\text{kg} \cdot \text{HM}}{\text{month}}$)           & 1642.5   & 535      \\
-                                                                             &          &          \\
-        Burnup  ($\frac{\text{MWe} \cdot \text{month}}{\text{kg} \cdot \text{HM}}$) & 0.547945 & 0.672897 \\
-                                                                             &          &          \\
-        Effective Power  ($\text{MWe}$)                                      & 900      & 360      \\
-                                                                             &          &          \\
-        Core Size  ($\text{kg} \cdot \text{HM}$)                                    & 88695    & 40125    \\
-                                                                             &          &          \\
-        \hline                       
-    \end{tabular}
-    \captionsetup{justification=centering}
-    \caption[Reactor parameter invariants]{
-        Reactor configuration invariants maintained across all four
-        simulation cases.
-    }
-
-    \label{tab:invar}
-\end{table}
-
-\begin{table*}
-    \centering
-    \begin{tabular}{ |r | c c | c c | }
-        \hline                       
-                                          & \multicolumn{2}{c |}{LWR}       & \multicolumn{2}{c |}{SFR} \\
-                                          & Cases 1,3 & Cases 2,4 (fleet) & Cases 1,3 & Cases 2,4 (fleet)  \\
-        \hline                       
-        Cycle length (months)             & 15        & 18                & 12        & 15 \\
-        Refueling outage (months)         & 3         & 0                 & 3         & 0 \\
-        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 29,565            & 8,025     & 40,125 \\
-        Power capacity (MWe)              & 1,080     & 900               & 450       & 360 \\
-        \hline                       
-    \end{tabular}
-    \captionsetup{justification=centering}
-    \caption[Reactor parameters by case]{
-        Reactor configuration for each of the four simulation cases.
-    }
-
-    \label{tab:reactors}
-\end{table*}
-
-The parameters for each case were also selected to keep capacity factors for
-individual reactors very roughly around 90\%.  The template input file for
-each of these cases is in Appendix TODO: exp2 input files in appendix.
-
-A longer 21-month deployment period is used because it provides better natural
-staggering for the reactor refueling cycles. The same deployment schedule is
-used for all four cases; power capacity is built along a curve starting at
-100GWe in year zero following the 1\% exponential growth curve.  This
-particular deployment schedule causes significant fuel shortages providing
-interesting features to investigate and compare.
 

### 2015-09-16 paper.tex
extra paragraph about pu comparison fig

@@
 
-Also notable is that different modeling choices (e.g. time step duration,
-facility discretization, etc.) may have varying levels of significance
-depending on the scenario/context they are operating in.  Figure
-\ref{fig:puinv-compare} shows the same inventory curves from Figure
-\ref{fig:puinv} superimposed.  Before the fuel shortage begins, differences
-between each of the four cases are relatively minor.  However, after the
-shortage begins around year 100, separated Pu inventories vary much more
-significantly between the different cases.
+Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
+\ref{fig:puinv} superimposed. Counter-intuitively, lower separated plutonium
+inventory during the shortage generally indicates fewer unfueled reactors.
+Lower inventory levels mean available Pu is being utilized more efficiently
+rather than sitting idle.  Cases 3 and 4 with their longer time steps show
+higher inventories indicating a more severe shortage.  Post shortage, they
+also fall a bit behind on building up Pu stocks due to the delayed
+contributions of more unfueled reactors to the separated Pu pool.
 
@@
 \end{figure*}
+
+Also notable is that different modeling choices (e.g. time step duration,
+facility discretization, etc.) may have varying levels of significance
+depending on the scenario/context they are operating in.  As shown in Figure
+\ref{fig:puinv-compare} Before the fuel shortage begins, differences between
+each of the four cases are relatively minor.  However, after the shortage
+begins around year 100, separated Pu inventories vary much more significantly
+between the different cases.
 

### 2015-09-17 paper.tex
add ORION and Cyclus citations

@@
 \cite{jacobson_verifiable_2010}, DANESS \cite{van_den_durpel_daness_2009},
-DYMOND (predecessor to DANESS), and ORION among others.  In fleet-based
-modeling:
+DYMOND (predecessor to DANESS), and ORION \cite{gregg_orion_2012} among
+others.  In fleet-based modeling:
 
@@
 One of the major motivations for the development of the Cyclus fuel cycle
-simulator was creating a simulation environment that is more expressive than
-more traditional fuel cycle simulators.  Cyclus was designed with a
-plugin-style architecture that allows different models for facility types to
-be swapped in and out trivially.  As a Cyclus simulation walks through each
-time step, facilities engage in a market-like dynamic resource exchange (DRE).
-They broadcast requests for material and suppliers respond with bids. The DRE
-then creates a network flow problem that it optimizes to resolve which
-transactions actually occur.  The plugin architecture combined with the DRE
-creates a powerful and flexible environment for the natural comparison of
-different modeling paradigms.
+simulator \cite{cyclus_2015} was creating a simulation environment that is
+more expressive than more traditional fuel cycle simulators.  Cyclus was
+designed with a plugin-style architecture that allows different models for
+facility types to be swapped in and out trivially.  As a Cyclus simulation
+walks through each time step, facilities engage in a market-like dynamic
+resource exchange (DRE).  They broadcast requests for material and suppliers
+respond with bids. The DRE then creates a network flow problem that it
+optimizes to resolve which transactions actually occur.  The plugin
+architecture combined with the DRE creates a powerful and flexible environment
+for the natural comparison of different modeling paradigms.
 

### 2015-09-17 exp2/gendata.sh
working on reproducibility instructions, provenance, etc.

@@ -43,36 +43,44 @@ function finalize {
     rm tmp1.dat tmp2.dat tmp3.dat tmp4.dat
 }
 
+# generate time series of pu flow out of the fuelfab.
 puflow case1.sqlite > tmp1.dat
 puflow case2.sqlite > tmp2.dat
 puflow case3.sqlite | expand > tmp3.dat
 puflow case4.sqlite | expand > tmp4.dat
 finalize puflow.dat
 
+# generate time series of pu flow into the fuelfab.
 puflowin case1.sqlite > tmp1.dat
 puflowin case2.sqlite > tmp2.dat
 puflowin case3.sqlite | expand > tmp3.dat
 puflowin case4.sqlite | expand > tmp4.dat
 finalize puflowin.dat
 
+# generate time series of pu inventory of the fuelfab
 puinv case1.sqlite > tmp1.dat
 puinv case2.sqlite > tmp2.dat
 puinv case3.sqlite | expand > tmp3.dat
 puinv case4.sqlite | expand > tmp4.dat
 finalize puinv.dat
 
+# generate time series of total power produced by all reactors
 power case1.sqlite > tmp1.dat
 power case2.sqlite > tmp2.dat
 power case3.sqlite | expand > tmp3.dat
 power case4.sqlite | expand > tmp4.dat
 finalize power.dat
 
+# generate time series of offline reactor capacity due to fuel shortage.
 indivunfueled case1.sqlite | topower > tmp1.dat
 fleetunfueled case2.sqlite > tmp2.dat
 indivunfueled case3.sqlite | expand | topower > tmp3.dat
 fleetunfueled case4.sqlite | expand > tmp4.dat
 finalize unfueled.dat
 
+# generate time series of offline reactor capacity due to inefficient fuel
+# sharing - fleet reactors have perfect sharing and so this metric has no
+# meaning in those cases.
 badshare case1.sqlite  > tmp1.dat
 badshare case3.sqlite | expand > tmp3.dat
 paste tmp1.dat tmp3.dat | awk 'NR==1 {print "Time\t\tCase1\t\tCase3"; next;} {print $1"\t\t"$2"\t\t"$4}' > badshare.dat

### 2015-09-17 exp2/runscen.sh

@@ -1,15 +1,15 @@
 #!/bin/bash
 
-# runscen takes 3 arguments - the first is the build schedule file name,
-# the second is the case to run (i.e. 1, 2, 3, or 4), and the third is the
-# file name for the generated output database.
+# runscen takes 2 arguments - the first is the build schedule file name,
+# the second is the case to run (i.e. 1, 2, 3, or 4)
 function runscen {
     db=$(cat $1 | cycobj -sched -scen scen-case${2}.json | grep -o ' \+[^ ]*\.sqlite')
-    mv $db $3
+    mv $db case${3}.sqlite
     rm -f *-*-*-*
 }
 
-runscen buildsched-monthly-noshortage.dat 1 noshortage-case1.sqlite
-runscen buildsched-monthly-noshortage.dat 2 noshortage-case2.sqlite
-runscen buildsched-quarterly-noshortage.dat 3 noshortage-case3.sqlite
-runscen buildsched-quarterly-noshortage.dat 4 noshortage-case4.sqlite
+runscen buildsched-monthly.dat 1
+runscen buildsched-monthly.dat 2
+runscen buildsched-quarterly.dat 3
+runscen buildsched-quarterly.dat 4
+

### 2015-09-17 exp2/gendata.sh
finish readme and correct comment in gendata

@@ -78,7 +78,7 @@ indivunfueled case3.sqlite | expand | topower > tmp3.dat
 fleetunfueled case4.sqlite | expand > tmp4.dat
 finalize unfueled.dat
 
-# generate time series of offline reactor capacity due to inefficient fuel
+# generate time series of unutilized batch-months due to inefficient fuel
 # sharing - fleet reactors have perfect sharing and so this metric has no
 # meaning in those cases.
 badshare case1.sqlite  > tmp1.dat

### 2015-09-17 paper.tex
add references to figshare tarball of code, data, etc.

@@
 complete refueling cycle. Fuel shortages also only cause proportional loss of
-capacity rather than whole-reactor quantized shutdown.
+capacity rather than whole-reactor quantized shutdown.  The full
+implementation of the fleet-based reactor is publicy available for download
+\cite{Carlsen2015}.
 
@@
 The input files and other assets used in the analysis for each of the four
-cases can be found in TODO: cite figshare tarball.
+cases are publicly available for download \cite{Carlsen2015}.
 
@@
 \section{Results}
+
+All results data, analysis tools and instructions are publicly available for
+download at \url{http://dx.doi.org/10.6084/m9.figshare.1546775}
+\cite{Carlsen2015}.
 

### 2015-09-18 paper.tex
explain how to measure modeling effects. tweak plots

@@
 produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if
-the fuel shortage actually ended after only 1 month.
+the fuel shortage actually ended after only 1 month.  Quantifying this effect
+directly is difficult.  One way of observing it is to compare fleet reactor
+scenarios with that have different time step durations (i.e. cases 2 and 4).
+Because the fleet reactors have perfect fuel sharing (described below) and no
+outage noise, differences observed will likely be due to quantized shutdown.
 
@@
 fabrication facility to maintain suboptimally large on-hand inventories as a
-contingency.
+contingency.  This effect can be naturally observed as the fluctuations in the
+difference between deployed power capacity and generated power.
 
@@
 operating or that 3 reactors are each missing a single batch.  This is
-referred to as the \emph{fuel sharing effect}.
+referred to as the \emph{fuel sharing effect}.  Fleet reactors, by design,
+have perfect fuel sharing - all fuel is distributed so that the minimum
+possible amount of capacity is offline during shortages. The most correct
+measure of fuel sharing inefficiency is the difference between the actual
+generated power and the maximum amount of power that could have been generated
+among all possible fuel distribution alternatives.  This is difficult to
+actually compute. One possible approximation is to count the number of fresh
+fuel batches distributed to reactors that ended up not having enough fuel to
+operate (i.e. a full core) that time step.  This approximation assumes that
+every "wasted" batch of fuel, could have been given to a reactor that only
+needed one batch to operate.
 
@@
 making offers. In general longer time steps and individual reactor modeling
-create a need for larger floating inventory buffers in order to avoid
-material shortages.
+create a need for larger floating inventory buffers in order to avoid material
+shortages.  Although direct measures of this effect are difficult, some of its
+consequences can be quantified by comparing scenarios with different time step
+durations (e.g. cases 2 and 4).
 

### 2015-09-18 paper.tex
improve puinv chart a bit
a bit more discussion about drawdown results

@@
 consequences can be quantified by comparing scenarios with different time step
-durations (e.g. cases 2 and 4).
+durations (e.g. cases 2 and 4) such as higher standing inventories during
+shortages.
 
@@
 available for making fuel. The in flows are not shown for the individual
-reactor cases because they are somewhat messy and obscure otherinformation in
+reactor cases because they are somewhat messy and obscure other information in
 the plots. The colored curves show Pu inventory available for making fast
@@
 their outflow for a particular time step.  This information lag effectively
-requires the floating Pu inventory to be maintained at a higher level.  This
-can be clearly seen in the case 4 plot which has a higher standing inventory
-than case 2 during the fuel shortage years.  The case 4 Pu inventory is unable
-to drop below approximately 100 tonnes, which indicates out-flows to be
-roughly equal to the Pu flow into inventory each time step during those years.
-This can be confirmed by the in-flow and inventory curves touching at those
-points.  Case 2 Pu inventory is similarly limited at about the 30 tonne level.
-As expected, the case 4 withdrawals are also three times higher case 2
-withdrawals.
+requires the floating Pu inventory to be maintained at a higher level during
+the shortage.  This can be clearly seen in the case 4 plot which has a higher
+standing inventory than case 2 during the fuel shortage years.  The case 4 Pu
+inventory is unable to drop below approximately 100 tonnes, which indicates
+out-flows to be roughly equal to the Pu flow into inventory each time step
+during those years.  This can be confirmed by the in-flow and inventory curves
+touching at those points.  Case 2 Pu inventory is similarly limited at about
+the 30 tonne level.  As expected, the case 4 withdrawals are also three times
+higher case 2 withdrawals.
+
+Throughout most of the shortage, the aggregate outflow of Pu from separated
+inventory is roughly the same in cases 2 and 4.  However, due to the longer
+time steps in case 4, the shortage starts slightly sooner because Pu inventory
+must be maintained at a higher limit as described above.  The small difference
+in outage start time has feed-back effects.  Not only are extra reactors
+offline, but those extra reactors are also discharging less aggregate spent
+fuel for recycling.  This farther increases the discrepency between the two
+cases. Eventually the shortage in case 2 ends slightly before it does in case
+4. This effect feedback occurs because the reactors in question have a
+conversion ratio greater than 1.0.
 

### 2015-09-18 paper.tex
add caption to pu comparison fig

@@
     \includegraphics[width=1.0\textwidth]{exp2/puinv.eps}
-    \caption[Pu inventory and flow]{
-        Pu inventory and flows for all four cases.  In-flows for cases 1 and 3
-        are omitted because they are very noisy and obscure useful other
-        information.  Larger time steps results in larger per time step
-        material flows. These larger flows mean more material is not available
-        for supplying to requesters.  This information lag causes more supply
-        constraints than occur with smaller time steps - illustrating the
-        \emph{inventory drawdown effect}.
+    \caption[Separated Pu inventory and flow]{
+        Separated Pu inventory and flows for all four cases.  In-flows for
+        cases 1 and 3 are omitted because they are very noisy and obscure
+        useful other information.  Larger time steps results in larger per
+        time step material flows. These larger flows mean more material is not
+        available for supplying to requesters.  This information lag causes
+        more supply constraints than occur with smaller time steps -
+        illustrating the \emph{inventory drawdown effect}.
     }
@@
     \includegraphics[width=0.8\textwidth]{exp2/puinv-compare.eps}
-    \caption[TODO]{
-        TODO
+    \caption{
+        Separated Pu inventories for all four cases.  Differences in modeling
+        choices are not particularly significant until about year 100 when
+        fuel shortages begin.  These differences gradually decay away starting
+        in year 140 after the fuel shortages have ebbed.  Higher aggregate Pu
+        inventory  during the shortage is correlated with a longer during and
+        post shortage recovery period.
     }

### 2015-09-18 paper.tex
add a bit more cyclus discussion to intro

@@
 \newpage
+
 \begin{abstract}
@@
 \section{Introduction}
+
+TODO: add good intro/overview paragraph
 
@@
 designed with a plugin-style architecture that allows different models for
-facility types to be swapped in and out trivially.  As a Cyclus simulation
-walks through each time step, facilities engage in a market-like dynamic
-resource exchange (DRE).  They broadcast requests for material and suppliers
-respond with bids. The DRE then creates a network flow problem that it
-optimizes to resolve which transactions actually occur.  The plugin
-architecture combined with the DRE creates a powerful and flexible environment
-for the natural comparison of different modeling paradigms.
+facility types, known as \emph{archetypes} to be swapped in and out trivially.
 
+As a Cyclus simulation walks through each time step, facilities engage in a
+market-like dynamic resource exchange (DRE).  They broadcast requests for
+material and suppliers respond with bids. The DRE then creates a network flow
+problem that is optimized to resolve which transactions actually occur.  The
+plugin architecture combined with the DRE creates a powerful and flexible
+environment for the natural comparison of different modeling paradigms.
 Cyclus' agent-based architecture in concert with its dynamic resource exchange
@@
 cycle staggering, and individual reactor outages.
+
+Cyclus was designed to operate natively in unix/linux type environments with
+flexible ascriptable usage and input file format.  Cyclus holds its diverse
+generated data into a powerful open-source database format.  In addition to
+information about material flows and facility deployments, Cyclus' single-file
+database contains version information about the archetypes used in a
+simulation, the simulation input file, version of Cyclus used to run the
+simulation, and the installed versions of Cyclus dependencies. This and other
+features of Cyclus lend themselves well to automation, research provenance and
+reproducibility, and large-scale computing.
 

### 2015-09-21 paper.tex
fix bullets from dissertation context to paper context

@@
 
-    \item How well do fleet-based models with continuous material flow capture
-    optimal fuel cycle transitions compared to individual facility modeling?
+    \item What effects drive differences in fleet-based and individual-based
+        facility modeling?
 
-    \item What effects are the primary drivers for differences in fleet-based
-    and individual-based facility modeling?
+    \item How does changing the time step duration affect simulations?
 
-    \item Does the impact of time step duration changes differ in fleet-based
+    \item Does the impact of changing time step duration differ in fleet-based
         and individual-based facility modeling?

### 2015-09-28 paper.tex
clarify drawdown explanation
discuss jit ordering and its affect on fuel utilization

@@
 create a need for larger floating inventory buffers in order to avoid material
-shortages.  Although direct measures of this effect are difficult, some of its
+shortages.  Another way to think about this effect is that a single material
+object can only move between a single pair of facilities per time step. Larger
+time steps result in longer minimum bounds on time to traverse paths between
+facilities. Although direct measures of this effect are difficult, some of its
 consequences can be quantified by comparing scenarios with different time step
@@
 also fall a bit behind on building up Pu stocks due to the delayed
-contributions of more unfueled reactors to the separated Pu pool.
+contributions of more unfueled reactors to the separated Pu pool.  The
+reactors in these scenarios request fresh fuel only when they need it. While
+this is an unrealistic behavior for real-world facilities, it is closer to a
+globally optimum fuel management strategy.  Keeping on-hand fresh fuel batches
+would also have the same effect of increasing idling plutonium quantities and
+making the fuel shortage worse overall.
 

### 2015-09-28 paper.tex
clarify flows in case1

@@
 The large differences seen between monthly and quarterly time step cases in
-Figures \ref{fig:unfueled} and \ref{fig:badshare} are
-primarily caused by the \emph{inventory drawdown effect}.  Figure
-\ref{fig:puinv} helps to visualize this effect.  The black dots in
-the figure represent impulse flows of available Pu out of the fuel fab and
-into fast reactors.  Blue lines represent the flows into Pu inventory
-available for making fuel. The in flows are not shown for the individual
-reactor cases because they are somewhat messy and obscure other information in
-the plots. The colored curves show Pu inventory available for making fast
-reactor fuel.  When the inventory curves meet and force down the flows, fuel
-shortages occur.
+Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily caused by the
+\emph{inventory drawdown effect}.  Figure \ref{fig:puinv} helps to visualize
+this effect.  The black dots in the figure represent impulse flows of
+available Pu out of the fuel fab and into fast reactors.  Blue lines represent
+the flows into Pu inventory available for making fuel. The in flows are not
+shown for the individual reactor cases because they are somewhat messy and
+obscure other information in the plots. The colored curves show Pu inventory
+available for making fast reactor fuel.  When the inventory curves meet and
+force down the out-flows, fuel shortages occur.  While this is most clear for
+cases 2 and 4, the individual reactor scenarios (cases 1 and 3) also have many
+points during the shortage where the outflow point lies exactly on top of the
+inventory point - indicating that all available inventory is being
+transferred.
 
@@
 the 30 tonne level.  As expected, the case 4 withdrawals are also three times
-higher case 2 withdrawals.
+higher case 2 withdrawals exactly matching the time step duration ratio
+between the two cases.
 
@@
 
+The striations visible in the monthly flows for case 1 in Figure
+\ref{fig:puinv} are a characteristic of the synchronization of reactor
+refueling caused by the shortage.  They are a sort of resonances of that the
+inventory level jumps between, and they gradually disappear as the refueling
+cycles become more staggered.
+
 \begin{figure*}[!h]
@@
     \caption[Separated Pu inventory and flow]{
+
         Separated Pu inventory and flows for all four cases.  In-flows for
-        cases 1 and 3 are omitted because they are very noisy and obscure
-        useful other information.  Larger time steps results in larger per
-        time step material flows. These larger flows mean more material is not
+        cases 1 and 3 are omitted because they are noisy and obscure other
+        useful information.  A larger time step results in larger per time
+        step material flows. These larger flows mean more material is not
         available for supplying to requesters.  This information lag causes
         more supply constraints than occur with smaller time steps -
-        illustrating the \emph{inventory drawdown effect}.
+        illustrating the \emph{inventory drawdown effect}. Another result of
+        this effect is the larger standing Pu inventory during the shortage
+        with larger time steps that can be seen in case 4 than in case 2.
+
     }

### 2015-09-29 paper.tex
add abstract and finish intro

@@
 \begin{abstract}
-TODO: Write this abstract.
+
+Due to the diversity of fuel cycle simulators' modeling assumptions direct
+comparison and benchmarking can be difficult.  In 2012 the Organisation for
+Economic Co-operation and Development (OECD) completed a benchmark study that
+is perhaps one of the most complete published comparisons performed.  Despite
+this, however, various results from the simulators were often significantly
+different due to inconsistency in modeling decisions such as reprocessing
+strategies, refueling behavior, reactor end-of-life handling, etc.  This work
+attempts to identify and quantify the effects of a few different modeling
+choices that may sometimes be taken for granted in the fuel cycle modeling
+domain.  Four scenarios were compared using combinations of fleet-based or
+individual reactors with monthly or quarterly (3-month) time steps.  When
+under supply-constraint pressure, the four cases exhibit noticably different
+behavior.  Otherwise, they generally behaved very similarly.  Fleet-based
+modeling is more efficient in supply-constrained environments at the expense
+of losing insight on issues such as reactor refueling cycle staggering.
+Finer-grained time steps also enable more efficient material use in
+supply-constrained environments. Large simulations with fleet-based reactors
+run much more quickly than their individual reactor counterparts.
+
 \end{abstract}
@@
 \section{Introduction}
-
-TODO: add good intro/overview paragraph
-
-\subsection{Motivation and Background}
 
@@
 modeling decisions such as reprocessing strategies, refueling behavior,
-reactor end-of-life handling, etc.
+reactor end-of-life handling, etc.  This work attempts to identify and
+quantify the effects of a few different modeling choices that may sometimes be
+taken for granted in the fuel cycle modeling domain: facility modeling
+discretization (i.e. individual facilities vs. aggregate fleets) and time step
+duration.  A better understanding of how these affect fuel cycle dynamics can
+allow for making more deliberate decisions with respect to tradeoffs such as
+performance vs. realism.
+
+\subsection{Motivation and Background}
 

### 2015-09-29 paper.tex
edits up to scenario description

@@
 this, however, various results from the simulators were often significantly
-different due to inconsistency in modeling decisions such as reprocessing
+different due to inconsistency in modeling decisions involving reprocessing
 strategies, refueling behavior, reactor end-of-life handling, etc.  This work
@@
 choices that may sometimes be taken for granted in the fuel cycle modeling
-domain.  Four scenarios were compared using combinations of fleet-based or
-individual reactors with monthly or quarterly (3-month) time steps.  When
-under supply-constraint pressure, the four cases exhibit noticably different
-behavior.  Otherwise, they generally behaved very similarly.  Fleet-based
-modeling is more efficient in supply-constrained environments at the expense
-of losing insight on issues such as reactor refueling cycle staggering.
+domain.  Four scenarios were compared using combinations of either fleet-based
+or individual reactors with either monthly or quarterly (3-month) time steps.
+When under supply-constraint pressure, the four cases exhibit noticably
+different behavior.  Otherwise, they generally behaved very similarly.
+Fleet-based modeling is more efficient in supply-constrained environments at
+the expense of losing insight on issues such as realistic sub-optimal fuel
+distribution and challenges in reactor refueling cycle staggering.
 Finer-grained time steps also enable more efficient material use in
@@
 comparisons performed.  Despite this, however, various results from the
-different simulators were often significantly different due to different
-modeling decisions such as reprocessing strategies, refueling behavior,
+different simulators were often significantly different due to inconsistent
+modeling decisions involving reprocessing strategies, refueling behavior,
 reactor end-of-life handling, etc.  This work attempts to identify and
@@
 \cite{forrester_industrial_1961} to modeling. These simulators model several
-"stocks" that have levels that change over time due to "flows" between them.
-The flows are determined by an equation that is a function of the state of the
-system (i.e. the system's stocks).  A constructed system is then solved
-in discrete or continuous time depending on the limitations of the software
-and constraints imposed by the model.  In system dynamics simulators, the
-number of stocks and potential flows between them is generally a static
-property of the modeled system, making it more difficult to swap different
-reactor models in and out of the same scenario without wiring them into the
-simulator manually.
+"stocks" representing isotopes or isotope groups that have levels that change
+over time due to "flows" between them.  The flows are determined by equations
+that are a function of the state of the system (i.e. the system's stocks).  A
+constructed system is then solved in discrete or continuous time depending on
+the limitations of the software and constraints imposed by the model.  In
+system dynamics simulators, the number of stocks and potential flows between
+them is generally a static property of the modeled system, making it somewhat
+difficult to swap different reactor models in and out of the same scenario
+without wiring them into the simulator manually.
 
 System dynamics based fuel cycle simulators generally use a collection of
-stocks to represent the state of groups of like facilities commonly referred
+stocks to represent the state of groups of like-facilities commonly referred
 to as "fleets".  As the simulation steps through time, the levels of stocks
@@
 
-This work investigates in detail differences between fleet vs. individual
+This work investigates in detail some differences between fleet vs. individual
 reactor modeling.  The difference between these two modeling styles in terms
@@
 material transfers between facilities and correspondingly affects the noise
-level in facility material inventories.  It can also act as a minimum bound
-for outage durations. The effects of the time step duration will also be
-investigated together with fleet vs facility modeling.
+level in facility material inventories and flows.  It can also act as a
+minimum bound for outage durations. The effects of the time step duration will
+also be investigated together with fleet vs facility modeling.
 
@@
 market-like dynamic resource exchange (DRE).  They broadcast requests for
-material and suppliers respond with bids. The DRE then creates a network flow
-problem that is optimized to resolve which transactions actually occur.  The
-plugin architecture combined with the DRE creates a powerful and flexible
-environment for the natural comparison of different modeling paradigms.
-Cyclus' agent-based architecture in concert with its dynamic resource exchange
-has the ability to model fleet-based facilities.  However, Cyclus was designed
-to enable modeling facilities individually with discrete flows enabling the
-investigation of interesting real-world effects such as competition, reactor
-cycle staggering, and individual reactor outages.
+material and suppliers respond with bids. Requesters then express preferences
+between all their bids, and the DRE then creates a network flow problem that
+is optimized to resolve which transactions actually occur.  The plugin
+architecture combined with the DRE creates a powerful and flexible environment
+for the natural comparison of different modeling paradigms.  Cyclus'
+agent-based architecture in concert with its DRE has the ability to model
+fleet-based facilities.  Additionally, however, Cyclus can model facilities
+individually with discrete flows enabling the investigation of interesting
+real-world effects such as competition, reactor cycle staggering, and
+individual reactor outages.
 
 Cyclus was designed to operate natively in unix/linux type environments with
-flexible ascriptable usage and input file format.  Cyclus holds its diverse
+flexible and scriptable usage and input file format.  Cyclus holds its diverse
 generated data into a powerful open-source database format.  In addition to
@@
 
-Because the standard package included with Cyclus does not include a
-fleet-based reactor model, one was necessarily developed. A simulation
-scenario was created that is roughly based on prior and ongoing work in the
-Department of Energy (DOE).  A few Phenomena related to the different modeling
-choices in each of the four cases are identified and used as a basis for
-comparison.  Each of these three things is described in more detail in the
-following sections.
+Because the standard package of archetypes included with Cyclus does not
+include a fleet-based reactor model, one was necessarily developed. A
+simulation scenario was created that is roughly based on prior and ongoing
+work in the Department of Energy (DOE) fuel cycle options campaign.  A few
+phenomena related to the different modeling choices in each of the four cases
+are identified and used as a basis for comparison.  Each of these things is
+described in more detail in the following sections.
 
@@
 A fleet reactor facility was created for Cyclus.  The flexibility of Cyclus
-makes it relatively straight forward to create plug-in models for Cyclus that
+makes it relatively straight forward to create plug-in archetypes that
 live anywhere along the spectrum from individual, discrete facilities to
@@
 deployments to transparently work with the fleet reactor without having to
-know anything about the fleet concept.  The fleet reactor has several
+know anything about the fleet paradigm.  The fleet reactor has several
 important characteristics described below:
@@
         (kg) for a single reactor's batch, $L$ is the cycle length (including
-        refueling time) in time steps, $N_{r}$ is the number of
-        reactors in the fleet, $C_{inv}$ is the total amount of fuel in all
-        reactor cores (kg), and $C_{cap}$ is the total fuel capacity for all
-        reactor cores (kg).  The last fraction term accounts for any amount of
+        refueling time) in time steps, $N_{r}$ is the number of reactors in
+        the fleet, $C_{inv}$ is the total amount of fuel in all reactor cores
+        (kg), and $C_{cap}$ is the total fuel capacity for all reactor cores
+        (kg).  The last fractional term accounts for any amount of
         underfueling that may be occuring resulting in lower fuel burning
@@
         available up to the total amount to fill all reactor cores in the
-        fleet.
+        fleet.  No extra fuel is kept on-hand - it is all ordered
+        just-in-time.
 
     \item Generated power is computed similarly to core discharge rates by
-        multiplying the total fleet power capacity by the same last term of
-        Equation \ref{eqn:fleet-discharge} to account for unfueled capacity in
-        the fleet.
+        multiplying the total fleet power capacity by the same last fractional
+        term of Equation \ref{eqn:fleet-discharge} to account for unfueled
+        capacity in the fleet.
 
@@
 
-These characteristics result in a fleet reactor having perfect fuel sharing
-and perfect cycle staggering among the individual reactors it models.  First
-discharge of spent fuel from fleets begins immediately rather than after a
-complete refueling cycle. Fuel shortages also only cause proportional loss of
-capacity rather than whole-reactor quantized shutdown.  The full
-implementation of the fleet-based reactor is publicy available for download
-\cite{Carlsen2015}.
+One characteristic of this fleet reactor implementation is that it has perfect
+fuel sharing and perfect cycle staggering among the individual reactors it
+models.  The first discharge of spent fuel from fleets begins immediately on
+the first time step of operation rather than after a complete refueling cycle.
+Fuel shortages also only cause proportional loss of capacity rather than
+whole-reactor quantized shutdown.  The full implementation of the fleet-based
+reactor is publicy available for download and use \cite{Carlsen2015}.
 

### 2015-09-29 paper.tex
more edits

@@
 choices that may sometimes be taken for granted in the fuel cycle modeling
-domain.  Four scenarios were compared using combinations of either fleet-based
+domain.  Four scenarios are compared using combinations of either fleet-based
 or individual reactors with either monthly or quarterly (3-month) time steps.
-When under supply-constraint pressure, the four cases exhibit noticably
-different behavior.  Otherwise, they generally behaved very similarly.
-Fleet-based modeling is more efficient in supply-constrained environments at
-the expense of losing insight on issues such as realistic sub-optimal fuel
-distribution and challenges in reactor refueling cycle staggering.
-Finer-grained time steps also enable more efficient material use in
-supply-constrained environments.  Large simulations with fleet-based reactors
-run much more quickly than their individual reactor counterparts.
+The Cyclus fuel cycle simulator is used as a level playing field to compare
+the scenarios.  When under supply-constraint pressure, the four cases exhibit
+noticably different behavior.  Otherwise, they generally behaved very
+similarly.  Fleet-based modeling is more efficient in supply-constrained
+environments at the expense of losing insight on issues such as realistically
+sub-optimal fuel distribution and challenges in reactor refueling cycle
+staggering.  Finer-grained time steps also enable more efficient material use
+in supply-constrained environments.  Large simulations with fleet-based
+reactors run much more quickly than their individual reactor counterparts.
 
 \end{abstract}
+
+\pagebreak
 
@@
 
-    \item groups of facilities (usually all facilities of a given type) are
+    \item Groups of facilities (usually all facilities of a given type) are
         treated as a single, aggregate entity.
 
-    \item resource flows are often (approximately) continuous (limited by time
-        step discretization)
+    \item Resource flows are often (approximately) continuous (limited by time
+        step discretization).
 
@@
 level in facility material inventories and flows.  It can also act as a
-minimum bound for outage durations. The effects of the time step duration will
-also be investigated together with fleet vs facility modeling.
+minimum bound for outage durations. The effects of the time step duration are
+also investigated together with fleet vs facility modeling.
 
@@
 Cyclus was designed to operate natively in unix/linux type environments with
-flexible and scriptable usage and input file format.  Cyclus holds its diverse
-generated data into a powerful open-source database format.  In addition to
-information about material flows and facility deployments, Cyclus' single-file
-database contains version information about the archetypes used in a
-simulation, the simulation input file, version of Cyclus used to run the
-simulation, and the installed versions of Cyclus dependencies. This and other
-features of Cyclus lend themselves well to automation, research provenance and
-reproducibility, and large-scale computing.
+flexible and scriptable usage and input file format.  Cyclus stores its
+diverse generated data in a powerful open-source database format.  In addition
+to information about material flows and facility deployments, Cyclus'
+single-file database contains the simulation input file, version information
+about the archetypes used in a simulation, the version of Cyclus used to run
+the simulation, and the installed versions of Cyclus dependencies. This and
+other features of Cyclus lend themselves well to automation, research
+provenance and reproducibility, and large-scale computing.
 
@@
 
-The scenario used is a transition from a the current once-through light water
+The scenario used approximates a transition from a the current once-through light water
 reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle known as
@@
 infinite capacity.  All spent fuel is stored/cooled for 7 years before it is
-available for separations. Figure \ref{fig:flow} shows the material flow
+available for separation. Figure \ref{fig:flow} shows the material flow
 relationships between the facilities.  Compositions used for the simulations
-were simple, containing primarily $^{238}U$, $^{235}U$, $^{239}Pu$,
-$^{241}Am$, and small quantities of a few lanthanides.
+were simple containing primarily $^{238}U$, $^{235}U$, $^{239}Pu$, $^{241}Am$,
+and small quantities of a few lanthanides.  Full composition details can be
+found in the Cyclus input files used \cite{Carlsen2015}.
 
@@
 
-Cyclus version ?.? (TODO:~update~version) was used for this analysis.
-Facility models used were the standard ones provided with Cyclus in the
-Cycamore library (also version ?.?).  A custom storage facility was used
-because Cycamore does not yet provide one.  The standard cycamore reactor was
-used for both reactor types (LWR and SFR) in cases 1 and 3, and the custom
-fleet-based reactor described earlier also used in cases 2 and 4.
-Additional parameters for the reactor types are shown in Table
-\ref{tab:reactors}.
+A pre-release of Cyclus version 1.4 was used for this analysis.  Archetypes
+used were the standard ones provided with Cyclus in the included Cycamore
+library - also a pre-release of version 1.4.  A custom storage facility was
+used because Cycamore did not yet provide one.  The standard cycamore reactor
+was used for both LWR and SRF reactor types in cases 1 and 3.  Correspondingly
+the custom fleet-based reactor described earlier was used for both reactor
+types in cases 2 and 4.  Additional configuration for the reactor types are
+shown in Table \ref{tab:reactors}.
 
@@
         \hline                       
-            & LWR & FR \\
+            & LWR & SFR \\
         \hline                       
@@
     \captionsetup{justification=centering}
-    \caption[Reactor facility parameters]{Reactor facility parameters. The cycle length includes the refueling outage.}
+    \caption[Reactor facility parameters]{Reactor facility parameters. Cycle length includes the refueling outage.}
     \label{tab:reactors}
@@
 The described scenario is used as closely as possible in each of the four
-cases.  However, a few other small case-dependent changes to the scenario
-relating to reactor configuration are necessary.  For the cases using
-quarterly time steps with individual reactors, cycle and outage times must be
-adjusted to be a multiple of 3 months.  In order to make the comparison more
-fair, the cases with monthly time steps are also adjusted to have identical
-cycle and outage times.  The invariants preserved with respect to reactor
-behavior are shown in Table \ref{tab:invar}.  Table \ref{tab:reactors} shows
-the computed/selected configuration for both reactor types for all four cases.
+cases.  A few other small case-dependent changes to the scenario relating to
+reactor configuration are necessary.  For the cases using quarterly time steps
+with individual reactors, cycle and outage times must be adjusted to be a
+multiple of 3 months.  In order to make the comparison more fair, the cases
+with monthly time steps are also adjusted to have identical cycle and outage
+times.  The invariants preserved with respect to reactor behavior between all
+four cases are shown in Table \ref{tab:invar}.  Table \ref{tab:reactors} shows
+the computed/selected configuration for both LWR and SFR reactor types for all
+four cases.
 

### 2015-09-29 paper.tex
edits

@@
 year 25.  Fast reactor spent fuel separations and fuel fabrication have
-infinite capacity.  All spent fuel is stored/cooled for 7 years before it is
+infinite capacity.  All spent fuel is stored/cooled for 7 years before it becomes
 available for separation. Figure \ref{fig:flow} shows the material flow
@@
 the custom fleet-based reactor described earlier was used for both reactor
-types in cases 2 and 4.  Additional configuration for the reactor types are
+types in cases 2 and 4.  Additional configuration for the reactor types is
 shown in Table \ref{tab:reactors}.
@@
 The described scenario is used as closely as possible in each of the four
-cases.  A few other small case-dependent changes to the scenario relating to
+cases.  However, a few small case-dependent changes to the scenarios relating to
 reactor configuration are necessary.  For the cases using quarterly time steps
@@
 In each of the four cases, the initial 100 LWRs are modified to have a zero
-length refueling outage with an explicit 900 MWe power capacity.  This was
-done to avoid having all the initial LWRs in the individual reactor cases
-refuel at the same time.  The parameters for each case were also selected to
-keep capacity factors for individual reactors very roughly around 90\%.
+length refueling outage with an explicit 900 MWe power capacity.  Because the
+individual reactor model doesn't have any way to start a reactor mid-cycle,
+this was done to avoid having all initial LWRs refuel at the same time.  The
+parameters for each case were also selected to keep capacity factors for
+individual reactors very roughly around 90\%.
 
-A longer 21-month deployment period is used because it provides better natural
-staggering for the reactor refueling cycles. The same deployment schedule is
-used for all four cases; power capacity is built along a curve starting at
-100GWe in year zero following the 1\% exponential growth curve.  This
-particular deployment schedule causes significant fuel shortages providing
-interesting features to investigate and compare.
+Instead of annual deployments, a longer 21-month deployment period is used
+because it provides better natural staggering for the reactor refueling
+cycles. The same deployment schedule is used for all four cases; power
+capacity is built along a curve starting at 100GWe in year zero following the
+1\% exponential growth curve.  This particular deployment schedule is chosen
+because it causes significant fuel shortages providing interesting features to
+investigate and compare.
 
-The input files and other assets used in the analysis for each of the four
-cases are publicly available for download \cite{Carlsen2015}.
+All input files and other assets used in the analysis for each of the four
+cases are publicly available for download and use \cite{Carlsen2015}.
 

### 2015-09-29 paper.tex
more edits

@@
 environments at the expense of losing insight on issues such as realistically
-sub-optimal fuel distribution and challenges in reactor refueling cycle
+suboptimal fuel distribution and challenges in reactor refueling cycle
 staggering.  Finer-grained time steps also enable more efficient material use
@@
 There are various low level phenomena that have potential to affect the larger
-outcome of simulations with respect to facility modeling decisions. Several
-effects are described below that result from fundamental differences between
-these two modeling paradigms.
+outcome of simulations with respect to facility and time step modeling
+decisions. Several effects are described below that arise from these
+different modeling choices.
 
@@
 \emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
-cores is short a half batch of fuel it will produce $\frac{5}{6}\cdot C$ of
-power for that time step where $C$ is the power capacity of a single reactor
-of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is short a
-half batch of fuel, then it will produce no power that time step.  Time step
-duration can also have this effect.  For both fleet and individual reactor
-modeling, power goes offline in full time step increments.  For example, if a
-3-batch fleet based reactor is short $\frac{1}{2}$ batch of fuel with a 3
-month time step, then it will produce $\frac{5}{6} \cdot 3 \text{mo} \cdot C $
-of energy.  In contrast, a monthly time step might allow the same reactor to
-produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if
-the fuel shortage actually ended after only 1 month.  Quantifying this effect
-directly is difficult.  One way of observing it is to compare fleet reactor
-scenarios with that have different time step durations (i.e. cases 2 and 4).
-Because the fleet reactors have perfect fuel sharing (described below) and no
-outage noise, differences observed will likely be due to quantized shutdown.
+cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot C$
+of power for that time step where $C$ is the power capacity of a single
+reactor of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is
+short a half batch of fuel, then it will produce no power ($0 \cdot C$) that
+time step.  Time step duration can also have this effect.  For both fleet and
+individual reactor modeling, power goes offline in full time step increments.
+For example, if a 3-batch fleet based reactor is short $\frac{1}{2}$ batch of
+fuel with a 3 month time step, then it will produce $\frac{2.5}{3} \cdot 3
+\text{mo} \cdot C $ of energy.  In contrast, a monthly time step might allow
+the same reactor to produce $[\frac{5}{6} \cdot 1 \text{mo} + 1 \cdot 2
+\text{mo}] \cdot C $ of energy if the fuel shortage actually ended after only
+1 month.  Quantifying this effect directly is difficult.  One way of observing
+it is to compare fleet reactor scenarios that have different time step
+durations (i.e. cases 2 and 4).  Because the fleet reactors have perfect fuel
+sharing (described below) and no outage noise, differences observed will
+likely be due to quantized shutdown.
 
@@
 vary significantly.  For example, if all reactors refuel at the same time,
-then all reactors would be offline at the same time.  This is referred to as
-the \emph{cycle staggering effect} and does not occur in fleet-based reactor
+then all reactors are offline at the same time.  This is referred to as the
+\emph{cycle staggering effect} and does not occur in fleet-based reactor
 models.  An example of poor cycle staggering is shown in Figure
-\ref{fig:sync-cycle}.  Poor staggering causes spikes of material supply and
-demand that can lead to suboptimal resource utilization.  A fuel fabrication
-facility might receive requests for new batches for every reactor all at once.
-The reactors draw out inventory from the fuel fabrication facility together in
-a large quantity of orders.  If the fabrication facility does not have enough
-on hand, some number of reactors will need to wait until enough fuel can be
-pumped through fabrication.  Avoiding such constraints would require the
-fabrication facility to maintain suboptimally large on-hand inventories as a
-contingency.  This effect can be naturally observed as the fluctuations in the
-difference between deployed power capacity and generated power.
+\ref{fig:sync-cycle}.  This effect can be naturally observed as the
+fluctuations in the difference between deployed power capacity and generated
+power.
+
+Poor reactor cycle staggering can also cause spikes of material supply and
+demand that can lead to suboptimal resource utilization.  Consider a fuel
+fabrication facility that receives requests for new batches for every reactor
+all at once.  The reactors draw out inventory from the fuel fabrication
+facility together in a large quantity of orders.  If the fabrication facility
+does not have enough material on hand, some number of reactors will need to
+wait until enough fuel can be pumped through fabrication.  Avoiding such
+constraints would require the fabrication facility to maintain suboptimally
+large on-hand inventories as a contingency.  Even with infinite material
+availability, poor staggering can result in supply constraints caused by
+finite facility throughput. Even if the fuel fabrication facility above has
+infinite material supply that it can keep on hand, it might not be able to
+fabricate fuel for all reactor requests all at once even though it has
+sufficient capacity if they request fuel uniformly over time.
 

### 2015-09-29 paper.tex
add discussion about time step duration effects on performance

@@
 involves fuel sharing.  Power production by a group of individual reactors for
-a fixed amount of fuel shortage may vary depending on how the existing fuel is
+a fixed amount of fuel shortage may vary depending on how available fuel is
 distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  A system
@@
 longer time steps, larger incoming quanta of material are not available for
-making offers. In general longer time steps and individual reactor modeling
-create a need for larger floating inventory buffers in order to avoid material
-shortages.  Another way to think about this effect is that a single material
-object can only move between a single pair of facilities per time step. Larger
-time steps result in longer minimum bounds on time to traverse paths between
-facilities. Although direct measures of this effect are difficult, some of its
-consequences can be quantified by comparing scenarios with different time step
-durations (e.g. cases 2 and 4) such as higher standing inventories during
-shortages.
+making offers. In general longer time steps create a need for larger floating
+inventory buffers in order to avoid material shortages.  Another aspect of
+this effect is that a single material object can only move between a single
+pair of facilities per time step. Larger time steps result in longer minimum
+bounds on time to traverse paths between facilities. Although direct measures
+of this effect are difficult, some of its consequences (e.g. higher standing
+inventories during shortages) can be quantified by comparing scenarios with
+different time step durations (e.g. cases 2 and 4).
 
@@
 sensitivity studies where running hundreds of thousands of simulations may be
-desirable.
+desirable.  The time step duration has less impact on performance because the
+amount of work being done in the simulation is the same; in the monthly time
+step cases most facilities (i.e. all reactors) are only perform actions every
+refueling cycle rather than every time step.
 

### 2015-09-29 paper.tex
more edits

@@
 All results data, analysis tools and instructions are publicly available for
-download at \url{http://dx.doi.org/10.6084/m9.figshare.1546775}
+download and use at \url{http://dx.doi.org/10.6084/m9.figshare.1546775}
 \cite{Carlsen2015}.
@@
 
-Figure \ref{fig:power} shows an overall view of the generated
-power over time for each of the four cases.  Figure
-\ref{fig:power-rel} normalizes the Figure
-\ref{fig:power} curves to the expected exponential power curve
-magnifying some interesting differences between the four cases.  For the first
-100 years, the four cases behave somewhat similarly, although the individually
-modeled reactors in cases 1 and 3 actually go online and offline for refueling
+Figure \ref{fig:power} shows an overall view of the generated power over time
+for each of the four cases.  Figure \ref{fig:power-rel} normalizes the Figure
+\ref{fig:power} curves to the expected exponential power curve magnifying some
+interesting differences between the four cases.  For the first 100 years, the
+four cases behave somewhat similarly, although the individually modeled
+reactors in cases 1 and 3 actually go offline and back online for refueling
 causing more variance.  Around year 100, a fuel shortage begins and more
 significant differences between the four cases become apparent. This fuel
-shortage lasts until about year 140. The power generated in Case 4 is lower
-than in case 2 during the fuel shortage years. This is primarily due to the
-\emph{quantized shutdown effect}.  The longer 3-month time step in case 4
+shortage lasts until about year 140. The power generated in Case 4 is slightly
+lower than in case 2 during the fuel shortage years. This is primarily due to
+the \emph{quantized shutdown effect}.  The longer 3-month time step in case 4
 results in reactor capacity going offline longer than necessary.
@@
         modeling, refueling cycles become much more synchronized during and
-        after the fuel shortage resuting in large power fluctuations (the
-        \emph{cycle staggering effect}).
+        after the fuel shortage resuting in large power fluctuations - an artifact of the
+        \emph{cycle staggering effect}.
     }
@@
         to about year 140.  The \emph{cycle staggering effect} can be seen as
-        divergence from 1.0 and is especially visible in cases 1 and 3 where
+        excessive divergence above and below $1.0$ in cases 1 and 3 where
         the shortage increases cycle synchronization significantly. The
@@
         reactor outage than necessary. The consistent jumping up and down for
-        the fleet reactor cases is due to the mis-alignment of reactors going
+        the fleet reactor cases (even pre-shortage) is due to the mis-alignment of reactors going
         offline (multiples of 12 months) and the build period (21 months).
@@
 
-Figure \ref{fig:unfueled} shows the fuel shortage more explicitly
-for each of the four cases. The individual reactor modeling cases result in
-more cumulative outage than corresponding fleet-based cases. The quarterly
-time step of cases 3 and 4 make the outage significantly more severe. Cases 1
+Figure \ref{fig:unfueled} shows the fuel shortage more explicitly for each of
+the four cases. The individual reactor modeling cases result in more
+cumulative outage than corresponding fleet-based cases. The quarterly time
+step of cases 3 and 4 also make the outage significantly more severe. Cases 1
 and 3 exhibit approximately twice as much offline power as their corresponding
@@
 sharing by design (zero inefficiency) and so is not included in this figure.
-Every batch that is given to a reactor that ends up not being able to start it
+Every batch that is given to a reactor that ends up not being able to start its
 cycle at the scheduled time (due to fuel shortage) adds one to this cumulative
@@
 One useful over-estimate of the fuel sharing inefficiency is to assume every
-one of these batches could have been given to a reactor enabling it to
+one of the "wasted" batches could have been given to a reactor enabling it to
 operate.  Multiplying each of these cumulative batches by reactor capacity
 (i.e. 450 MWe) provides a way to compare fuel sharing inefficiency with the
-overall fuel outage.  Correcting for capacity factor in is appropriate these
-cases because the outage is many reactor cycle lengths in duration. This
-results in a cumulative offline power caused by poor fuel sharing of roughly
-1,600 $\text{GWe} \cdot \text{months}$ and 5,400 $\text{GWe} \cdot
-\text{months}$ for cases 1 and 3 respectively.  These approximations are a
-significant portion of the difference between individual and fleet cases in
-Figure \ref{fig:unfueled} suggesting that most of the discrepency
-in cumulative outage between fleet and individual modeling is caused by
-inefficient fuel sharing in the individual reactor cases.
+overall fuel outage.  This results in a cumulative offline power caused by
+poor fuel sharing of roughly 2,000 $\text{GWe} \cdot \text{months}$ and 6,600
+$\text{GWe} \cdot \text{months}$ for cases 1 and 3 respectively.  These
+approximations are a significant portion of the difference between individual
+and fleet cases in Figure \ref{fig:unfueled} suggesting that most of the
+discrepency in cumulative outage between fleet and individual modeling is
+caused by inefficient fuel sharing in the individual reactor cases.
 
@@
 they cannot operate.  The real world doesn't necessarily optimize for fuel
-sharing efficiency very cleanly. There are other factors that can affect fuel
-sharing outcomes.  A few include:
+sharing efficiency very cleanly. There are many factors that can affect
+real-world fuel sharing outcomes.  A few include:
 
@@
 maintain optimal sharing, fresh fuel now needs to be sent in all-or-nothing
-multi-assembly quanta with preference for reactors that need fewer assemblies.
-Spare fresh fuel inventory (item 2 above) will have the global effect
-increasing total outage severity; on average more batches of fresh fuel will
-be sitting unused.  However, it can potentially reduce the frequency of
-outages for individual reactors in some cases.
+multi-assembly quanta with higher preference to reactors that need fewer
+assemblies.  Spare fresh fuel inventory (item 2 above) will have the global
+effect of increasing total outage severity; on average more batches of fresh
+fuel will be sitting unused.  However, it can potentially reduce the frequency
+of outages for individual reactors in some cases.
 

### 2015-09-29 paper.tex
more edits

@@
 shown for the individual reactor cases because they are somewhat messy and
-obscure other information in the plots. The colored curves show Pu inventory
+obscure other information in the plots. The other colored curves show Pu inventory
 available for making fast reactor fuel.  When the inventory curves meet and
@@
 Quarterly time steps have less frequent but larger impulse material flows.
-These larger flows, however, do not compensate for the larger withdrawals
+The larger in-flows, however, do not compensate for the larger withdrawals
 because facilities do not know about incoming inventory when they resolve
@@
 Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
-\ref{fig:puinv} superimposed. Counter-intuitively, lower separated plutonium
+\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated plutonium
 inventory during the shortage generally indicates fewer unfueled reactors.
@@
 contributions of more unfueled reactors to the separated Pu pool.  The
-reactors in these scenarios request fresh fuel only when they need it. While
+reactors in all the scenarios request fresh fuel only when they need it. While
 this is an unrealistic behavior for real-world facilities, it is closer to a
@@
 would also have the same effect of increasing idling plutonium quantities and
-making the fuel shortage worse overall.
+would make the fuel shortage worse overall.
 
@@
 depending on the scenario/context they are operating in.  As shown in Figure
-\ref{fig:puinv-compare} Before the fuel shortage begins, differences between
+\ref{fig:puinv-compare} before the fuel shortage begins, differences between
 each of the four cases are relatively minor.  However, after the shortage
-begins around year 100, separated Pu inventories vary much more significantly
+begins near year 100, separated Pu inventories vary much more significantly
 between the different cases.
@@
 modeling can be very useful for certain analyses, but it can reduce result
-quality if things such as cycle staggering are handled appropriately.  The
-fleet and discreet modeling done here is itself multi-dimensional.  If
-individual outages are not of interest, reactor power could be lowered and the
-outage reduced to zero duration making the capacity factor part of the
-reactor's power capacity.  The cycle staggering effects present when modeling
-reactors individually can provide valuable insight not possible with the
-fleet-based reactor modeling.  Starting reactors following shortage induced
-outages should not follow the natural pacing of fuel availability, otherwise
-refueling cycles become unsatisfactorily synchronized.
+quality if things such as cycle staggering are not handled appropriately.  The
+fleet and discreet modeling done here is itself multi-dimensional. If fuel
+sharing is of interest, but cycle staggering is not, individual reactors can
+be used with the power capacity lowered and the outage reduced to zero
+duration making the capacity factor included in the reactor's power capacity.
+Modeling reactors individually can provide valuable insight not possible with
+the fleet-based reactor modeling.  For example, bringing reactors back online
+following shortage-induced outages should not follow the natural pacing of
+fuel availability, otherwise refueling cycles become unsatisfactorily
+synchronized.
 
 Increasing the time steps increases facilities' standing inventory requirement
-for avoiding supply constraints.  Facilities that have unconstrained supply
-and very large inventory buffers, are not affected by the time step in this
-way.  Facilities can provide no more inventory than they can hold on a single
-time step.  So increasing time step duration for a facility with unconstrained
-supply generally means higher standing inventories are required in order to
-achieve equal throughput. Even when facilities do not have individual
+for avoiding supply constraints.  Facilities can provide no more inventory
+than they can hold on a single time step.  So increasing time step duration
+for a facility with unconstrained supply generally means higher standing
+inventories are required in order to achieve equal throughput. Facilities that
+have unconstrained supply and very large inventory buffers, are not affected
+meaningfully by the time step. Even when facilities do not have individual
 inventory or throughput limits, the time step can still have a significant
@@
 introduced through the loop (i.e. the discharge rate of spent fuel from
-reactors).  The \emph{inventory drawdown effect} can reduce the aggregate
+reactors).  The \emph{inventory drawdown effect} can impact the aggregate
 throughput of the recycling loop.
@@
 The impact of modeling assumptions depends heavily on the metrics of interest.
-For example, someone interested in fuel-shortage driven reactor outages could
+For example, someone interested in fuel-shortage driven reactor outages would
 observe a factor of four span among the four cases shown in the detail
-comparison.  However, someone more interested in the overall transition speed
-would not see as significant a discrepancy.
-
-In an unconstrained supply setting, varying the time step duration or
-discretization of facility modeling have a relatively minor impact on the
-overall simulation results.  Differences become more significant in material
-constrained regimes.  Robustness in constrained regimes is especially useful
-in the context of sensitivity and optimization analysis.
+comparison.  However, someone more interested in the overall transition
+mechanics would not see as significant a discrepancy.  In an unconstrained
+supply setting, varying the time step duration or discretization of facility
+modeling have a relatively minor impact on the overall simulation results.
+Differences become more significant in material constrained regimes.
+Robustness in constrained regimes is especially useful in the context of
+sensitivity and optimization analysis.
 

### 2015-09-30 paper.tex
adjust y range and legend placement for puinv figure
add todos
relabel four cases with MI,MF,QI,QF mneumonics. Add a bit more discussion of striations

@@
 \begin{document}
+
+TODO: add table titles
+TODO: adjust figure fonts to meet NT guidlines
 
@@
 
-\begin{itemize}
+\begin{enumerate}
 
-    \item \emph{Case 1}: Monthly time steps with individual reactor modeling
-    \item \emph{Case 2}: Monthly time steps with fleet reactor modeling
-    \item \emph{Case 3}: Quarterly (3-month) time steps with individual reactor modeling
-    \item \emph{Case 4}: Quarterly (3-month) time steps with fleet reactor modeling
+    \item \emph{Case MI}: Monthly time steps with individual reactor modeling
+    \item \emph{Case MF}: Monthly time steps with fleet reactor modeling
+    \item \emph{Case QI}: Quarterly (3-month) time steps with individual reactor modeling
+    \item \emph{Case QF}: Quarterly (3-month) time steps with fleet reactor modeling
 
-\end{itemize}
+\end{enumerate}
+
+The case labels consist of two letters each: "M" for monthly time steps, "Q"
+for quarterly time steps, "I" for individual reactors, and "F" for fleet
+reactors.
 
@@
 used because Cycamore did not yet provide one.  The standard cycamore reactor
-was used for both LWR and SRF reactor types in cases 1 and 3.  Correspondingly
+was used for both LWR and SRF reactor types in cases MI and QI.  Correspondingly
 the custom fleet-based reactor described earlier was used for both reactor
-types in cases 2 and 4.  Additional configuration for the reactor types is
+types in cases MF and QF.  Additional configuration for the reactor types is
 shown in Table \ref{tab:reactors}.
@@
                                           & \multicolumn{2}{c |}{LWR}       & \multicolumn{2}{c |}{SFR} \\
-                                          & Cases 1,3 & Cases 2,4 (fleet) & Cases 1,3 & Cases 2,4 (fleet)  \\
+                                          & Cases MI,QI & Cases MF,QF (fleet) & Cases MI,QI & Cases MF,QF (fleet)  \\
         \hline                       
@@
 it is to compare fleet reactor scenarios that have different time step
-durations (i.e. cases 2 and 4).  Because the fleet reactors have perfect fuel
+durations (i.e. cases MF and QF).  Because the fleet reactors have perfect fuel
 sharing (described below) and no outage noise, differences observed will
@@
 inventories during shortages) can be quantified by comparing scenarios with
-different time step durations (e.g. cases 2 and 4).
+different time step durations (e.g. cases MF and QF).
 
@@
 four cases behave somewhat similarly, although the individually modeled
-reactors in cases 1 and 3 actually go offline and back online for refueling
+reactors in cases MI and QI actually go offline and back online for refueling
 causing more variance.  Around year 100, a fuel shortage begins and more
 significant differences between the four cases become apparent. This fuel
-shortage lasts until about year 140. The power generated in Case 4 is slightly
-lower than in case 2 during the fuel shortage years. This is primarily due to
-the \emph{quantized shutdown effect}.  The longer 3-month time step in case 4
+shortage lasts until about year 140. The power generated in case QF is slightly
+lower than in case MF during the fuel shortage years. This is primarily due to
+the \emph{quantized shutdown effect}.  The longer 3-month time step in case QF 
 results in reactor capacity going offline longer than necessary.
 
-During the initial 100 years, cases 1 and 3 have larger variance in power
-output than cases 2 and 4.  This is caused by minor refueling cycle
+During the initial 100 years, cases MI and QI have larger variance in power
+output than cases MF and QF.  This is caused by minor refueling cycle
 synchronization.  The \emph{cycle staggering effect} becomes much more visible
@@
         Generated power for all four cases.  A fuel shortage occurs from about
-        year 100 to about year 140.  In cases 1 and 3 with individual reactor
+        year 100 to about year 140.  In cases MI and QI with individual reactor
         modeling, refueling cycles become much more synchronized during and
@@
         to about year 140.  The \emph{cycle staggering effect} can be seen as
-        excessive divergence above and below $1.0$ in cases 1 and 3 where
+        excessive divergence above and below $1.0$ in cases MI and QI where
         the shortage increases cycle synchronization significantly. The
@@
         between quarterly and monthly time steps for fleet reactor modeling
-        (cases 2 and 4) during the shortage; a longer time step causes more
+        (cases MF and QF) during the shortage; a longer time step causes more
         reactor outage than necessary. The consistent jumping up and down for
@@
 cumulative outage than corresponding fleet-based cases. The quarterly time
-step of cases 3 and 4 also make the outage significantly more severe. Cases 1
-and 3 exhibit approximately twice as much offline power as their corresponding
-fleet based simulations in cases 2 and 4.  The monthly time step simulations
+step of cases QI and QF also make the outage significantly more severe. Cases MI
+and QI exhibit approximately twice as much offline power as their corresponding
+fleet based simulations in cases MF and QF.  The monthly time step simulations
 also have roughly twice as much offline power as their corresponding quarterly
@@
 poor fuel sharing of roughly 2,000 $\text{GWe} \cdot \text{months}$ and 6,600
-$\text{GWe} \cdot \text{months}$ for cases 1 and 3 respectively.  These
+$\text{GWe} \cdot \text{months}$ for cases MI and QI respectively.  These
 approximations are a significant portion of the difference between individual
@@
 sharing is probably best accomplished with a more direct, intentional approach
-rather than as a modeling artifact as seen here in cases 1 and 3.  One way to
+rather than as a modeling artifact as seen here in cases MI and QI.  One way to
 alleviate the poor fuel sharing is to modify the individual reactor model to
@@
         sharing by design and so are not included here. The large difference
-        between cases 1 and 3 is mostly due to the \emph{inventory drawdown
+        between cases MI and QI is mostly due to the \emph{inventory drawdown
         effect}. 
@@
 force down the out-flows, fuel shortages occur.  While this is most clear for
-cases 2 and 4, the individual reactor scenarios (cases 1 and 3) also have many
+cases MF and QF, the individual reactor scenarios (cases MI and QI) also have many
 points during the shortage where the outflow point lies exactly on top of the
@@
 requires the floating Pu inventory to be maintained at a higher level during
-the shortage.  This can be clearly seen in the case 4 plot which has a higher
-standing inventory than case 2 during the fuel shortage years.  The case 4 Pu
+the shortage.  This can be clearly seen in the case QF plot which has a higher
+standing inventory than case MF during the fuel shortage years.  The case QF Pu
 inventory is unable to drop below approximately 100 tonnes, which indicates
@@
 during those years.  This can be confirmed by the in-flow and inventory curves
-touching at those points.  Case 2 Pu inventory is similarly limited at about
-the 30 tonne level.  As expected, the case 4 withdrawals are also three times
-higher case 2 withdrawals exactly matching the time step duration ratio
+touching at those points.  Case MF Pu inventory is similarly limited at about
+the 30 tonne level.  As expected, the case QF withdrawals are also three times
+higher case MF withdrawals exactly matching the time step duration ratio
 between the two cases.
@@
 Throughout most of the shortage, the aggregate outflow of Pu from separated
-inventory is roughly the same in cases 2 and 4.  However, due to the longer
-time steps in case 4, the shortage starts slightly sooner because Pu inventory
+inventory is roughly the same in cases MF and QF.  However, due to the longer
+time steps in case QF, the shortage starts slightly sooner because Pu inventory
 must be maintained at a higher limit as described above.  The small difference
@@
 fuel for recycling.  This farther increases the discrepency between the two
-cases. Eventually the shortage in case 2 ends slightly before it does in case
-4. This effect feedback occurs because the reactors in question have a
+cases. Eventually the shortage in case MF ends slightly before it does in case
+QF. This effect feedback occurs because the reactors in question have a
 conversion ratio greater than 1.0.
 
-The striations visible in the monthly flows for case 1 in Figure
+The striations visible in the monthly flows for case MI in Figure
 \ref{fig:puinv} are a characteristic of the synchronization of reactor
-refueling caused by the shortage.  They are a sort of resonances of that the
-inventory level jumps between, and they gradually disappear as the refueling
-cycles become more staggered.
+refueling caused by the shortage.  They begin at the end of the fuel shortage
+window when many reactors come online together and are resonances that the
+inventory level jumps between.  These resonances gradually disappear as the
+simulation progresses farther in time beyond the shortage and refueling cycles
+become more staggered.
 
@@
     \caption[Separated Pu inventory and flow]{
-
         Separated Pu inventory and flows for all four cases.  In-flows for
-        cases 1 and 3 are omitted because they are noisy and obscure other
+        cases MI and QI are omitted because they are noisy and obscure other
         useful information.  A larger time step results in larger per time
@@
         this effect is the larger standing Pu inventory during the shortage
-        with larger time steps that can be seen in case 4 than in case 2.
-
+        with larger time steps that can be seen in case QF than in case MF.
     }
@@
 Lower inventory levels mean available Pu is being utilized more efficiently
-rather than sitting idle.  Cases 3 and 4 with their longer time steps show
+rather than sitting idle.  Cases QI and QF with their longer time steps show
 higher inventories indicating a more severe shortage.  Post shortage, they
@@
 
+\pagebreak
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-09-30 paper.tex
More puinv plot discussion. make some figs wider.

@@
     \centering
-    \includegraphics[width=0.8\textwidth]{exp2/power.eps}
+    \includegraphics[width=1.0\textwidth]{exp2/power.eps}
     \caption[Generated power]{
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{exp2/power-rel.eps}
+    \includegraphics[width=1.0\textwidth]{exp2/power-rel.eps}
     \caption[Normalized power]{
@@
 available Pu out of the fuel fab and into fast reactors.  Blue lines represent
-the flows into Pu inventory available for making fuel. The in flows are not
+the flows into Pu inventory available for making fuel. The in-flows are not
 shown for the individual reactor cases because they are somewhat messy and
 obscure other information in the plots. The other colored curves show Pu inventory
-available for making fast reactor fuel.  When the inventory curves meet and
+available for fabricating fast reactor fuel.  When the inventory curves meet and
 force down the out-flows, fuel shortages occur.  While this is most clear for
@@
 cases. Eventually the shortage in case MF ends slightly before it does in case
-QF. This effect feedback occurs because the reactors in question have a
+QF. This feedback effect occurs because the reactors in question have a
 conversion ratio greater than 1.0.
 
-The striations visible in the monthly flows for case MI in Figure
+The horizontal striations visible in the monthly flows for case MI in Figure
 \ref{fig:puinv} are a characteristic of the synchronization of reactor
@@
 become more staggered.
+
+The in-flows shown for cases MF and QF show a periodic pulsing that begins
+during the shortage in about year 120.  This pusling actually has nothing to
+do with the shortage and is caused by the retirement of fast reactors.  The
+first fast reactors are deployed in year 35 and have an 80 year lifetime.
+When they retire, they discharge an entire core's worth of fuel rather than a
+single batch.  After a 7-year cooling period, the first discharged full cores
+begin to make it through the recycle loop in year 122.  Another interesting
+feature visible in all four cases in Figure \ref{fig:puinv} is the minority
+fraction of out-flow points that are higher than the bulk majority. These
+higher out-flows occur once every 21 months when new reactors are built. The
+new reactors require a full core of fresh fuel rather than the single batch
+needed when just refueling.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{exp2/puinv-compare.eps}
+    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare.eps}
     \caption{

### 2015-10-01 paper.tex
fix grammar mistake

@@
 continues, however, new reactors continue to be deployed and reactors with
-synchornized cycles retire resulting in a gradual return to decent better
+synchornized cycles retire resulting in a gradual return to better
 staggering visible by the end of the 200 years.

### 2015-10-01 paper.tex
rewording conclusion a bit

@@
 quality if things such as cycle staggering are not handled appropriately.  The
-fleet and discreet modeling done here is itself multi-dimensional. If fuel
-sharing is of interest, but cycle staggering is not, individual reactors can
-be used with the power capacity lowered and the outage reduced to zero
-duration making the capacity factor included in the reactor's power capacity.
+fleet and discreet models used here are just two points in a multi-dimensional
+spectrum of modeling choices. If fuel sharing is of interest, but cycle
+staggering is not, individual reactors can be used with the power capacity
+lowered and the outage reduced to zero duration making the capacity factor
+included in the reactor's power capacity.
+
 Modeling reactors individually can provide valuable insight not possible with
 the fleet-based reactor modeling.  For example, bringing reactors back online
-following shortage-induced outages should not follow the natural pacing of
-fuel availability, otherwise refueling cycles become unsatisfactorily
-synchronized.
+following shortage-induced outages should not necessarily follow the natural
+pacing of fuel availability, otherwise refueling cycles become
+unsatisfactorily synchronized.  If individual reactor modeling fidelity is not
+required, then performance benefits in both simulation run time and data
+analysis may suggest a fleet-based modeling approach.
 
-Increasing the time steps increases facilities' standing inventory requirement
-for avoiding supply constraints.  Facilities can provide no more inventory
-than they can hold on a single time step.  So increasing time step duration
-for a facility with unconstrained supply generally means higher standing
-inventories are required in order to achieve equal throughput. Facilities that
-have unconstrained supply and very large inventory buffers, are not affected
-meaningfully by the time step. Even when facilities do not have individual
-inventory or throughput limits, the time step can still have a significant
-effect. In recycle loops there is a constraint on how fast material is
-introduced through the loop (i.e. the discharge rate of spent fuel from
-reactors).  The \emph{inventory drawdown effect} can impact the aggregate
-throughput of the recycling loop.
+Increasing time step duration increases facilities' standing inventory
+requirement during supply constrained periods.  Facilities can provide no more
+inventory than they can hold on a single time step.  So increasing time step
+duration for a facility with unconstrained supply generally means higher
+standing inventories are required in order to achieve equal throughput.
+Facilities with large inventory buffer that are unconstrained are not affected
+significantly by the time step duration when operating in isolation.  However,
+even when facilities do not have individual inventory or throughput limits,
+the time step can still have a significant effect. In recycle loops there is a
+constraint on how fast material is introduced through the loop (i.e. the
+discharge rate of spent fuel from reactors).  The \emph{inventory drawdown
+effect} can impact the aggregate throughput of the recycling loop by affecting
+minimum bounds on idling material quantities.
 
@@
 
+TODO: thank ross maybe for vision-sourced inspiration about modeling effects
+TODO: thank neup and add neup logo
+
 \pagebreak

### 2015-10-01 paper.tex
add closing paragraph and neup logo

@@
 
+The value of realism in fuel cycle modeling is greatly be enhanced by an
+associated understanding of how outcomes are affected by added fidelity.
+Cyclus' flexibility for accomodating different modeling choices uniquely
+enables many interesting comparisons.  Continued work in this direction has
+great potential to to improve the integrity of fuel cycle analysis generally.
+
 \section{Acknowledgements}
@@
 TODO: thank ross maybe for vision-sourced inspiration about modeling effects
-TODO: thank neup and add neup logo
+
+This research is being performed in part using funding received from the DOE
+Office of Nuclear Energy's Nuclear Energy University Programs.  The authors
+thank the NEUP for its generous support.
+
+\begin{center}
+\includegraphics[width=1.5in]{neup_logo_large.jpg}
+\end{center}
 

### 2015-10-02 paper.tex
edits from dad up to scenario description section

@@
 
-Due to the diversity of fuel cycle simulators' modeling assumptions direct
+Due to the diversity of fuel cycle simulator modeling assumptions, direct
 comparison and benchmarking can be difficult.  In 2012 the Organisation for
@@
 strategies, refueling behavior, reactor end-of-life handling, etc.  This work
-attempts to identify and quantify the effects of a few different modeling
-choices that may sometimes be taken for granted in the fuel cycle modeling
-domain.  Four scenarios are compared using combinations of either fleet-based
-or individual reactors with either monthly or quarterly (3-month) time steps.
-The Cyclus fuel cycle simulator is used as a level playing field to compare
-the scenarios.  When under supply-constraint pressure, the four cases exhibit
-noticably different behavior.  Otherwise, they generally behaved very
-similarly.  Fleet-based modeling is more efficient in supply-constrained
-environments at the expense of losing insight on issues such as realistically
-suboptimal fuel distribution and challenges in reactor refueling cycle
-staggering.  Finer-grained time steps also enable more efficient material use
-in supply-constrained environments.  Large simulations with fleet-based
-reactors run much more quickly than their individual reactor counterparts.
+attempts to identify and quantify the effects of selected modeling choices
+that may sometimes be taken for granted in the fuel cycle modeling domain.
+Four scenarios are compared using combinations of either fleet-based or
+individual reactors with either monthly or quarterly (3-month) time steps.
+The Cyclus fuel cycle simulator's plug-in facility cability along with its
+market-like dynamic material routing allow it to be used as a level playing
+field for comparing the scenarios.  When under supply-constraint pressure, the
+four cases exhibit noticably different behavior.  Fleet-based modeling is more
+efficient in supply-constrained environments at the expense of losing insight
+on issues such as realistically suboptimal fuel distribution and challenges in
+reactor refueling cycle staggering.  Finer-grained time steps also enable more
+efficient material use in supply-constrained environments.  Large simulations
+with fleet-based reactors run much more quickly than their individual reactor
+counterparts.
 
@@
 
-Due to the diversity of fuel cycle simulators' modeling assumptions direct
-comparison and benchmarking can be difficult.  In 2012 the Organisation for
+The diversity of assumptions embedded within various fuel cycle simulators pose challenges for direct
+comparison and benchmarking.  In 2012 the Organisation for
 Economic Co-operation and Development (OECD) completed a benchmark study
@@
 reactor end-of-life handling, etc.  This work attempts to identify and
-quantify the effects of a few different modeling choices that may sometimes be
-taken for granted in the fuel cycle modeling domain: facility modeling
+quantify the effects of selected modeling choices relative to facility modeling
 discretization (i.e. individual facilities vs. aggregate fleets) and time step
@@
 allow for making more deliberate decisions with respect to tradeoffs such as
-performance vs. realism.
+computational investment vs. realism.
 
@@
 the limitations of the software and constraints imposed by the model.  In
-system dynamics simulators, the number of stocks and potential flows between
+system dynamics based simulators, the number of stocks and potential flows between
 them is generally a static property of the modeled system, making it somewhat
@@
 (e.g. reactor fresh fuel inventory, repository waste inventories, etc.) are
-adjusted according to the calculated flow values.  Simulators that model
+adjusted according to the calculated flow values.  Fuel cycle simulators that model
 facility fleets built on a system dynamics foundation include VISION
 \cite{jacobson_verifiable_2010}, DANESS \cite{van_den_durpel_daness_2009},
-DYMOND (predecessor to DANESS), and ORION \cite{gregg_orion_2012} among
+DYMOND, and ORION \cite{gregg_orion_2012} among
 others.  In fleet-based modeling:
@@
 reactor modeling.  The difference between these two modeling styles in terms
-of their affects on output of fuel cycle simulators in general is currently
-unknown. This analysis is designed to provide both quantitative and
+of their affects on output of fuel cycle simulators in general is not well
+known. This analysis is designed to provide both quantitative and
 qualitative insight about how fleet-based continuous flow models differ from
@@
 level in facility material inventories and flows.  It can also act as a
-minimum bound for outage durations. The effects of the time step duration are
+minimum bound for outage durations and is a minimum bound in general for state
+changes within the entire system. The effects of the time step duration are
 also investigated together with fleet vs facility modeling.
@@
 designed with a plugin-style architecture that allows different models for
-facility types, known as \emph{archetypes} to be swapped in and out trivially.
+facility types, known as \emph{archetypes}, to be swapped in and out easily.
 
@@
 market-like dynamic resource exchange (DRE).  They broadcast requests for
-material and suppliers respond with bids. Requesters then express preferences
-between all their bids, and the DRE then creates a network flow problem that
-is optimized to resolve which transactions actually occur.  The plugin
-architecture combined with the DRE creates a powerful and flexible environment
-for the natural comparison of different modeling paradigms.  Cyclus'
-agent-based architecture in concert with its DRE has the ability to model
-fleet-based facilities.  Additionally, however, Cyclus can model facilities
-individually with discrete flows enabling the investigation of interesting
-real-world effects such as competition, reactor cycle staggering, and
-individual reactor outages.
+material and potential supplying facilities respond with bids. The requesting facilities then
+express preferences between all their bids, and the DRE then generates a network
+flow problem that is optimized to resolve which transactions actually occur.
+The plugin architecture combined with the DRE creates a powerful and flexible
+environment for the natural comparison of different modeling paradigms.
+Cyclus' module plug-in based architecture in concert with its DRE has the ability to
+model fleet-based facilities.  Additionally, however, Cyclus can model
+facilities individually with discrete flows enabling the investigation of
+interesting real-world effects such as competition, reactor cycle staggering,
+and individual reactor outages.
 
@@
 flexible and scriptable usage and input file format.  Cyclus stores its
-diverse generated data in a powerful open-source database format.  In addition
+comprehensive output data in a powerful open-source database format.  In addition
 to information about material flows and facility deployments, Cyclus'
@@
 phenomena related to the different modeling choices in each of the four cases
-are identified and used as a basis for comparison.  Each of these things is
+are identified and used as a basis for comparison.  These are
 described in more detail in the following sections.
@@
 
-A fleet reactor facility was created for Cyclus.  The flexibility of Cyclus
-makes it relatively straight forward to create plug-in archetypes that
-live anywhere along the spectrum from individual, discrete facilities to
+A fleet reactor archetype was created for Cyclus.  The flexibility of Cyclus
+makes it relatively straight forward to create plug-in archetypes to model
+anywhere along the spectrum from individual, discrete facilities to
 continuous flow fleet facilities.  In order to contrast well with individual,
-single-facility granularity modeling, the fleet model created here falls more
+single-facility granularity modeling, the fleet archetype created here falls more
 toward the aggregate, continuous flow side of the spectrum.
@@
         (kg), and $C_{cap}$ is the total fuel capacity for all reactor cores
-        (kg).  The last fractional term accounts for any amount of
-        underfueling that may be occuring resulting in lower fuel burning
-        rates due to unfueled reactors.
+        (kg).  The last fractional term accounts for lower fuel burning rates
+        due to unfueled reactors.
 
@@
     \item When a reactor in the fleet is retired, a full core of material is
-        discharged - even if there is some fraction the fleet with unfueled
+        discharged - even if there is some fraction of the fleet with unfueled
         cores.
@@
 fuel sharing and perfect cycle staggering among the individual reactors it
-models.  The first discharge of spent fuel from fleets begins immediately on
+models.  The first discharge of spent fuel begins immediately on
 the first time step of operation rather than after a complete refueling cycle.
@@
 used were the standard ones provided with Cyclus in the included Cycamore
-library - also a pre-release of version 1.4.  A custom storage facility was
+library - also a pre-release of version 1.4.  A custom storage archetype was
 used because Cycamore did not yet provide one.  The standard cycamore reactor
@@
 There are various low level phenomena that have potential to affect the larger
-outcome of simulations with respect to facility and time step modeling
+outcome of simulations with respect to facility discretization and time step modeling
 decisions. Several effects are described below that arise from these

### 2015-10-02 paper.tex
more edits

@@
 
-The scenario used approximates a transition from a the current once-through light water
-reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle known as
-EG-23 in the DOE's Advanced Nuclear Fuel Cycle Options report
+The scenario modeled approximates a transition from a the current once-through light water
+reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle within evaluation group 23 (EG-23)
+in the DOE's Advanced Nuclear Fuel Cycle Options report
 \cite{wigeland_nuclear_2014}. Scenario and transition details were patterned
@@
 annual growth rate for 200 years.  The decommissioning of the initial 100
-reactors is staggered over years 15 to 55.  Enrichment feeds LWRs with fresh
-fuel with infinite capacity. In year 35, fast reactors become available for
+reactors is staggered over years 15 to 55.  Fresh fuel for LWRs is provided directly from enrichment with infinite supply and throughput. In year 35, fast reactors become available for
 deployment and no more thermal reactors are built.  LWR spent fuel separations
@@
 infinite capacity.  All spent fuel is stored/cooled for 7 years before it becomes
-available for separation. Figure \ref{fig:flow} shows the material flow
+available for separation: 5 years for cooling and an additional 2 years to approximate non-instantaneous separations and fabrication times. Figure \ref{fig:flow} shows the material flow
 relationships between the facilities.  Compositions used for the simulations
@@
 library - also a pre-release of version 1.4.  A custom storage archetype was
-used because Cycamore did not yet provide one.  The standard cycamore reactor
-was used for both LWR and SRF reactor types in cases MI and QI.  Correspondingly
-the custom fleet-based reactor described earlier was used for both reactor
+used because Cycamore did not yet provide one.  The standard cycamore reactor archetype
+was used for both LWR and SFR reactor types in cases MI and QI.  Correspondingly
+the custom fleet-based reactor archetype described earlier was used for both reactor
 types in cases MF and QF.  Additional configuration for the reactor types is
@@
 with individual reactors, cycle and outage times must be adjusted to be a
-multiple of 3 months.  In order to make the comparison more fair, the cases
+multiple of 3 months.  In order to make the comparison more valid, the cases
 with monthly time steps are also adjusted to have identical cycle and outage
 times.  The invariants preserved with respect to reactor behavior between all
-four cases are shown in Table \ref{tab:invar}.  Table \ref{tab:reactors} shows
+four cases are shown in Table \ref{tab:invar}.  Table \ref{tab:reactor-detail} shows
 the computed/selected configuration for both LWR and SFR reactor types for all
@@
 
-    \label{tab:reactors}
+    \label{tab:reactor-detail}
 \end{table*}
@@
 In each of the four cases, the initial 100 LWRs are modified to have a zero
-length refueling outage with an explicit 900 MWe power capacity.  Because the
+length refueling outage with an explicit 900 MWe net power output capacity.  Because the
 individual reactor model doesn't have any way to start a reactor mid-cycle,

### 2015-10-02 paper.tex
improve cycle staggering bad example plot

@@
 durations (i.e. cases MF and QF).  Because the fleet reactors have perfect fuel
-sharing (described below) and no outage noise, differences observed will
+sharing (described below) and no noise from refueling outages, differences observed will
 likely be due to quantized shutdown.
@@
 fluctuations in the difference between deployed power capacity and generated
-power.
+power.  The figure shows power jumping above and below the installed net
+capacity (includes capacity factor) from time step to time step depending on
+how many reactors are online/offline together.
 
@@
 does not have enough material on hand, some number of reactors will need to
-wait until enough fuel can be pumped through fabrication.  Avoiding such
+wait until enough fuel can be fabricated.  Avoiding such
 constraints would require the fabrication facility to maintain suboptimally
@@
         their cycles synced as well.  While it may look like separate curves,
-        this plot shows only a single time series.
+        generated power curve (purple) represents only a single time series with many discontinuities.
     }

### 2015-10-02 paper.tex
edits up to results section

@@
 inventories. This by itself is not problematic because this is balanced by
-correspondingly larger inventory top-up quantities.  The other component of
+correspondingly larger inventory top-up quantities.  However, another component of
 this effect is the simultaneity of incoming and outgoing material flows on a
@@
 inventory buffers in order to avoid material shortages.  Another aspect of
-this effect is that a single material object can only move between a single
-pair of facilities per time step. Larger time steps result in longer minimum
+this effect is that at least one time step is required for a material object to move between two facilities. Larger time steps result in longer minimum
 bounds on time to traverse paths between facilities. Although direct measures
@@
 cycle analysis in general, particularly when considering optimization or
-sensitivity studies where running hundreds of thousands of simulations may be
+sensitivity studies where running thousands or millions of simulations may be
 desirable.  The time step duration has less impact on performance because the
-amount of work being done in the simulation is the same; in the monthly time
-step cases most facilities (i.e. all reactors) are only perform actions every
+amount of work being done in the simulation is similar; in the monthly time
+step cases most facilities (i.e. all reactors) only perform actions every
 refueling cycle rather than every time step.

### 2015-10-02 paper.tex
edits up to page 21 in manuscript from dad

@@
 
-All results data, analysis tools and instructions are publicly available for
-download and use at \url{http://dx.doi.org/10.6084/m9.figshare.1546775}
-\cite{Carlsen2015}.
+All results data, analysis tools, and related instructions are publicly
+available for download and use at
+\url{http://dx.doi.org/10.6084/m9.figshare.1546775} \cite{Carlsen2015}.
 
@@
 one of the "wasted" batches could have been given to a reactor enabling it to
-operate.  Multiplying each of these cumulative batches by reactor capacity
+operate.  Multiplying each of these cumulative batches by the fast reactor power capacity
 (i.e. 450 MWe) provides a way to compare fuel sharing inefficiency with the
-overall fuel outage.  This results in a cumulative offline power caused by
+overall fuel outage.  This results in a cumulative energy deficit caused by
 poor fuel sharing of roughly 2,000 $\text{GWe} \cdot \text{months}$ and 6,600
 $\text{GWe} \cdot \text{months}$ for cases MI and QI respectively.  These
-approximations are a significant portion of the difference between individual
+approximations account for a significant portion of the difference between individual
 and fleet cases in Figure \ref{fig:unfueled} suggesting that most of the
@@
 assemblies.  Spare fresh fuel inventory (item 2 above) will have the global
-effect of increasing total outage severity; on average more batches of fresh
+effect of increasing total shortage severity; on average more batches of fresh
 fuel will be sitting unused.  However, it can potentially reduce the frequency
@@
 alleviate the poor fuel sharing is to modify the individual reactor model to
-lower the preference value for requests for fresh fuel depending on how many
+adjust the preference value on requests for fresh fuel depending on how many
 assemblies it needs to have a full core.  The more assemblies it needs, the
 lower it will set its request preference.  This will have the effect of
-allowing the resource exchange resolution in Cyclus to globally prefer sending
+allowing the DRE in Cyclus to prefer sending
 fuel to reactors that need less to operate.
@@
         by the number of time steps that reactor had to delay the start of its
-        next cycle (due to fuel shortage).  Fleet reactors have perfect fuel
-        sharing by design and so are not included here. The large difference
+        next cycle (due to fuel shortage).  Fleet reactors implicitly have perfect fuel
+        sharing and so are not included here. The large difference
         between cases MI and QI is mostly due to the \emph{inventory drawdown
@@
 this effect.  The black dots in the figure represent impulse flows of
-available Pu out of the fuel fab and into fast reactors.  Blue lines represent
+available Pu from fuel fabrication and into fast reactors.  Blue lines represent
 the flows into Pu inventory available for making fuel. The in-flows are not
@@
 requires the floating Pu inventory to be maintained at a higher level during
-the shortage.  This can be clearly seen in the case QF plot which has a higher
+the shortage.  This can be clearly seen in the Figure \ref{fig:puinv} case QF plot which has a higher
 standing inventory than case MF during the fuel shortage years.  The case QF Pu
@@
 the 30 tonne level.  As expected, the case QF withdrawals are also three times
-higher case MF withdrawals exactly matching the time step duration ratio
+higher than the case MF withdrawals, exactly matching the time step duration ratio
 between the two cases.

### 2015-10-05 paper.tex
finish edits from dad

@@
 offline, but those extra reactors are also discharging less aggregate spent
-fuel for recycling.  This farther increases the discrepency between the two
+fuel for recycling.  This increases the discrepency between the two
 cases. Eventually the shortage in case MF ends slightly before it does in case
-QF. This feedback effect occurs because the reactors in question have a
+QF. This feedback effect occurs because the reactors have a
 conversion ratio greater than 1.0.
@@
 fraction of out-flow points that are higher than the bulk majority. These
-higher out-flows occur once every 21 months when new reactors are built. The
+higher out-flows occur once every 21 months when new reactors are built because
 new reactors require a full core of fresh fuel rather than the single batch
@@
         cases MI and QI are omitted because they are noisy and obscure other
-        useful information.  A larger time step results in larger per time
+        useful information.  When out-flows meet inventory, fuel shortages are taking place. A larger time step results in larger per time
         step material flows. These larger flows mean more material is not
@@
         more supply constraints than occur with smaller time steps -
-        illustrating the \emph{inventory drawdown effect}. Another result of
+        illustrating the \emph{inventory drawdown effect}. Another consequence of
         this effect is the larger standing Pu inventory during the shortage
@@
 rather than sitting idle.  Cases QI and QF with their longer time steps show
-higher inventories indicating a more severe shortage.  Post shortage, they
-also fall a bit behind on building up Pu stocks due to the delayed
+higher inventories indicating a more severe shortage.  Post shortage, cases QI and QF
+also fall a bit behind with respect to building up Pu stocks due to the delayed
 contributions of more unfueled reactors to the separated Pu pool.  The
@@
         fuel shortages begin.  These differences gradually decay away starting
-        in year 140 after the fuel shortages have ebbed.  Higher aggregate Pu
-        inventory  during the shortage is correlated with a longer during and
-        post shortage recovery period.
+        in year 140 after fuel shortages have ebbed.  Higher aggregate Pu
+        inventory  during the shortage is correlated with both a longer shortage and
+        a longer post-shortage recovery period.
     }
@@
 quality if things such as cycle staggering are not handled appropriately.  The
-fleet and discreet models used here are just two points in a multi-dimensional
-spectrum of modeling choices. If fuel sharing is of interest, but cycle
-staggering is not, individual reactors can be used with the power capacity
-lowered and the outage reduced to zero duration making the capacity factor
-included in the reactor's power capacity.
+fleet and discrete models used here are just two points in a multi-dimensional
+spectrum of modeling choices. For example, if fuel sharing is of interest but cycle
+staggering is not, individual reactors could be used with the power capacity
+lowered and the outage reduced to zero duration, including the capacity factor
+in the reactor's power capacity.
 
@@
 even when facilities do not have individual inventory or throughput limits,
-the time step can still have a significant effect. In recycle loops there is a
-constraint on how fast material is introduced through the loop (i.e. the
-discharge rate of spent fuel from reactors).  The \emph{inventory drawdown
-effect} can impact the aggregate throughput of the recycling loop by affecting
+the time step can still have a significant effect. The \emph{inventory drawdown
+effect} can impact the aggregate throughput of supply-constrained recycle loops by affecting
 minimum bounds on idling material quantities.

### 2015-10-06 paper.tex
working on NT journal manuscript detail requirements

@@
 \usepackage[acronym,toc]{glossaries}
-\newacronym{MIT}{MIT}{the Massachusetts Institute of Technology}
 \newacronym{UW}{UW}{University of Wisconsin}
@@
 %%%%%%%% cover info %%%%%%%%%%%%%%%
-\title{Facility and Time Step Discretization Effects in Fuel Cycle Simulation}
+\title{Challenging Fuel Cycle Modeling Assumptions: Facility and Time Step Discretization Effects}
 \author{Robert~W.~Carlsen, Paul~P.H.~Wilson}
@@
 % No more than three keywords, though each can be a phrase
-\keywords{nuclear fuel cycle, simulation, TODO:third}
+\keywords{Nuclear Fuel Cycle, Simulation, \Cyclus}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-
-TODO: add table titles
-TODO: adjust figure fonts to meet NT guidlines
-
-\newpage
-
-\begin{abstract}
-
+\myabstract{
 Due to the diversity of fuel cycle simulator modeling assumptions, direct
@@
 counterparts.
+}
 
-\end{abstract}
-
-\pagebreak
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
 
 \section{Introduction}
+
+TODO: add table titles
+TODO: adjust figure fonts to meet NT guidlines
 
@@
 
-\subsection{CYCLUS}
+\subsection{\Cyclus}
 

### 2015-10-06 paper.tex
turn table captions into titles

@@
 types in cases MF and QF.  Additional configuration for the reactor types is
-shown in Table \ref{tab:reactors}.
+shown in Table \ref{tab:reactors} (cycle length includes the refueling outage).
 
@@
 \begin{table}[h]
+    \label{tab:reactors}
+    \caption{Reactor Facility Parameters}
     \centering
@@
     \captionsetup{justification=centering}
-    \caption[Reactor facility parameters]{Reactor facility parameters. Cycle length includes the refueling outage.}
-    \label{tab:reactors}
 \end{table}
@@
 \begin{table}
+    \label{tab:invar}
+    \captionsetup{justification=centering}
+    \caption{Reactor Parameter Invariants}
     \centering
@@
     \end{tabular}
-    \captionsetup{justification=centering}
-    \caption[Reactor parameter invariants]{
-        Reactor configuration invariants maintained across all four
-        simulation cases.
-    }
-
-    \label{tab:invar}
 \end{table}
@@
 \begin{table*}
+    \label{tab:reactor-detail}
     \centering
+    \captionsetup{justification=centering}
+    \caption{Reactor Parameters by Case}
     \begin{tabular}{ |r | c c | c c | }
@@
     \end{tabular}
-    \captionsetup{justification=centering}
-    \caption[Reactor parameters by case]{
-        Reactor configuration for each of the four simulation cases.
-    }
-
-    \label{tab:reactor-detail}
 \end{table*}

### 2015-10-07 paper.tex
label must go last in table or referencing doesn't work right
remove done todos
provide more definition of fleet and continuous

@@
 
-TODO: add table titles
-TODO: adjust figure fonts to meet NT guidlines
-
 The diversity of assumptions embedded within various fuel cycle simulators pose challenges for direct
@@
     \item Groups of facilities (usually all facilities of a given type) are
-        treated as a single, aggregate entity.
+        treated as a single, aggregate entity. Therefore, all facilities experience
+        identical operational histories.
 
-    \item Resource flows are often (approximately) continuous (limited by time
-        step discretization).
+    \item Resource flows are modeled as continuous, within the limitations of time
+        step discretization.  Therefore, material is acquired and discharge in
+        uniform amounts at every time step.
 
@@
 \begin{table}[h]
-    \label{tab:reactors}
     \caption{Reactor Facility Parameters}
@@
     \captionsetup{justification=centering}
+    \label{tab:reactors}
 \end{table}
@@
 \begin{table}
-    \label{tab:invar}
     \captionsetup{justification=centering}
@@
     \end{tabular}
+    \label{tab:invar}
 \end{table}
@@
 \begin{table*}
-    \label{tab:reactor-detail}
     \centering
@@
     \end{tabular}
+    \label{tab:reactor-detail}
 \end{table*}
@@
 There are various low level phenomena that have potential to affect the larger
-outcome of simulations with respect to facility discretization and time step modeling
-decisions. Several effects are described below that arise from these
-different modeling choices.
+outcome of simulations with respect to facility discretization and time step
+modeling decisions. Several effects arising from these different modeling
+choices are present in many fuel cycle simulators to varying degrees.  Some of
+these effects are described in detail below.
 
@@
         every 3rd year refuel together.  All initial 100 LWRs start out with
-        their cycles synced as well.  While it may look like separate curves,
-        generated power curve (purple) represents only a single time series with many discontinuities.
+        their cycles synced as well.
     }
@@
 every "wasted" batch of fuel, could have been given to a reactor that only
-needed one batch to operate.
+needed one batch to operate.  Equation \ref{eq:badshare} describes such an
+approximation.
+
+\begin{equation}
+    N_{wasted}(t) = \sum\limits_{r \in R_t} \left(H[t-S_{sched}(t,r)] \cdot [1-O(t,r)]  \sum\limits_{t'=F_{prev}(t,r)}^{t} N_b(t',r) \right)
+    \label{eq:badshare}
+\end{equation}
+
+where $R_t$ is the set of all reactors, $H$ is the Heaviside function,
+$S_{sched}(t,r)$ is the original scheduled startup time of reactor $r$ after
+the most recent refueling outage that started before $t$, $F_{prev}(t,r)$ is
+the start of the most recent refueling outage for reactor $r$ before or on
+time $t$, $O(t,r)$ is 1 if reactor $r$ is producing power at time $t$ and zero
+otherwise, and $N_b(t,r)$ is the number of new fuel batches received by
+reactor $r$ since time $t$.
 
@@
 inventories. This by itself is not problematic because this is balanced by
-correspondingly larger inventory top-up quantities.  However, another component of
-this effect is the simultaneity of incoming and outgoing material flows on a
-given time step.  Within a particular time step facilities do not know about
-incoming material when they make commitments for outgoing material.  For
-longer time steps, larger incoming quanta of material are not available for
-making offers. In general longer time steps create a need for larger floating
-inventory buffers in order to avoid material shortages.  Another aspect of
-this effect is that at least one time step is required for a material object to move between two facilities. Larger time steps result in longer minimum
-bounds on time to traverse paths between facilities. Although direct measures
-of this effect are difficult, some of its consequences (e.g. higher standing
-inventories during shortages) can be quantified by comparing scenarios with
-different time step durations (e.g. cases MF and QF).
+correspondingly larger inventory top-up quantities.  However, another
+component of this effect is the simultaneity of incoming and outgoing material
+flows on a given time step.  Within a particular time step facilities do not
+know about incoming material when they make commitments for outgoing material.
+For longer time steps, larger incoming quanta of material are not available
+for making offers. In general longer time steps create a need for larger
+floating inventory buffers in order to avoid material shortages.  Another
+aspect of this effect is that at least one time step is required for a
+material object to move between two facilities. Larger time steps result in
+longer minimum bounds on time to traverse paths between facilities. Although
+direct measures of this effect are difficult, some of its consequences (e.g.
+higher standing inventories during shortages) can be quantified by comparing
+scenarios with different time step durations (e.g.  cases MF and QF).
 
@@
 
+For the fleet reactors, fuel shortage is exactly the
+difference between generated power and installed power capacity. Measuring the
+fuel shortage for individual reactor cases, however, requires careful
+accounting of the difference between reactors that are offline for a normal
+refueling outage and reactors that are offline because they have insufficient
+fuel. Equation \ref{eq:shortage} shows how this is calculated for cases MI and
+QI.
+
+\begin{equation}
+    P_{outage}(t) = \sum\limits_{r \in R_t} C_r \cdot H[t-S_{sched}(t,r)] \cdot [1-O(t,r)]
+    \label{eq:shortage}
+\end{equation}
+
+where $R_t$ is the set of all reactors, $C_r$ is the power capacity of reactor
+$r$, $H$ is the Heaviside function, $S_{sched}(t,r)$ is the original scheduled
+startup time of reactor $r$ after the most recent refueling outage that
+started before $t$, and $O(t,r)$ is 1 if reactor $r$ is producing power at
+time $t$ and zero otherwise.
+
 Figure \ref{fig:unfueled} shows the fuel shortage more explicitly for each of
-the four cases. The individual reactor modeling cases result in more
-cumulative outage than corresponding fleet-based cases. The quarterly time
-step of cases QI and QF also make the outage significantly more severe. Cases MI
-and QI exhibit approximately twice as much offline power as their corresponding
-fleet based simulations in cases MF and QF.  The monthly time step simulations
-also have roughly twice as much offline power as their corresponding quarterly
-cases.
+the four cases with the case MI and QI results being the cumulative version of
+Equation \ref{eq:shortage}.  The individual reactor modeling cases result in
+more cumulative outage than corresponding fleet-based cases. The quarterly
+time step of cases QI and QF also make the outage significantly more severe.
+Cases MI and QI exhibit approximately twice as much offline power as their
+corresponding fleet based simulations in cases MF and QF.  The monthly time
+step simulations also have roughly twice as much offline power as their
+corresponding quarterly cases.
 
@@
 Quantifying imperfect fuel sharing is a bit difficult, but Figure
-\ref{fig:badshare} gives one way to approximate the \emph{fuel
-sharing effect}.  Fleet reactor modeling by design exhibits perfect fuel
-sharing by design (zero inefficiency) and so is not included in this figure.
-Every batch that is given to a reactor that ends up not being able to start its
-cycle at the scheduled time (due to fuel shortage) adds one to this cumulative
-for each time step the cycle start is delayed.
+\ref{fig:badshare} provides one way to see the \emph{fuel sharing effect}
+showing wasted batch-months computed using Equation \ref{eq:badshare}.  Fleet
+reactor modeling by design exhibits perfect fuel sharing by design (zero
+inefficiency) and so is not included in this figure.  Every batch that is
+given to a reactor that ends up not being able to start its cycle at the
+scheduled time (due to fuel shortage) adds one to this cumulative for each
+time step the cycle start is delayed.
 
@@
 requires the floating Pu inventory to be maintained at a higher level during
-the shortage.  This can be clearly seen in the Figure \ref{fig:puinv} case QF plot which has a higher
+the shortage.  This can be clearly seen in Figures \ref{fig:puinv} and \ref{fig:puinv-compare} where case QF has a higher
 standing inventory than case MF during the fuel shortage years.  The case QF Pu
@@
 cases. Eventually the shortage in case MF ends slightly before it does in case
-QF. This feedback effect occurs because the reactors have a
+QF (shown most clearly in Figure \ref{fig:puinv-compare-zoom}). This feedback effect occurs because the reactors have a
 conversion ratio greater than 1.0.
@@
 \ref{fig:puinv} are a characteristic of the synchronization of reactor
-refueling caused by the shortage.  They begin at the end of the fuel shortage
-window when many reactors come online together and are resonances that the
+refueling caused by the shortage.  A close-up view of these can be seen in
+Figure \ref{fig:puinv-MI}. They begin at the end of the fuel shortage window
+when many reactors come online together in groups and are resonances that the
 inventory level jumps between.  These resonances gradually disappear as the
@@
 The in-flows shown for cases MF and QF show a periodic pulsing that begins
-during the shortage in about year 120.  This pusling actually has nothing to
+during the shortage in about year 120 and is shown in greater detail in Figure \ref{fig:puinv-MF}. This pulsing actually has nothing to
 do with the shortage and is caused by the retirement of fast reactors.  The
@@
 
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/puinv-MI.eps}
+    \caption[Separated Pu inventory and flow: Case MI Zoom]{
+        Separated Pu inventory and flows for case MI.  At locations where the
+        outflow dots lie on/near the inventory level, fuel shortages are
+        taking place.  This begins around year 105 and continues until about
+        year 140 where Pu withdrawals become lower than standing inventory.
+        The horizontal bands in out-flows starting around year 140 are caused
+        by many previously offline reactors coming online together in groups.
+        These fuel inventory level resonances gradually disappear as the
+        simulation progresses.
+    }
+    \label{fig:puinv-MI}
+\end{figure*}
+
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/puinv-MF.eps}
+    \caption[Separated Pu inventory and flow: Case MF Zoom]{
+        Separated Pu inventory and flows for case MF.  At locations where the
+        outflow dots meet the inventory level, fuel shortages are taking
+        place.  This begins around year 105 and continues until about year 140
+        where Pu withdrawals become lower than standing inventory.  The
+        periodic pulsing that begins just after year 120 is caused by the
+        first fast reactor retirements; the first fast reactors deployed starting in
+        year 35 retire annually after their 80 year lifetime and discharge an entire core's
+        worth of fuel rather than the usual single batch for refueling.
+    }
+    \label{fig:puinv-MF}
+\end{figure*}
+
 Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
-\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated plutonium
-inventory during the shortage generally indicates fewer unfueled reactors.
-Lower inventory levels mean available Pu is being utilized more efficiently
-rather than sitting idle.  Cases QI and QF with their longer time steps show
-higher inventories indicating a more severe shortage.  Post shortage, cases QI and QF
-also fall a bit behind with respect to building up Pu stocks due to the delayed
-contributions of more unfueled reactors to the separated Pu pool.  The
-reactors in all the scenarios request fresh fuel only when they need it. While
-this is an unrealistic behavior for real-world facilities, it is closer to a
-globally optimum fuel management strategy.  Keeping on-hand fresh fuel batches
-would also have the same effect of increasing idling plutonium quantities and
-would make the fuel shortage worse overall.
+\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated
+plutonium inventory during the shortage generally indicates fewer unfueled
+reactors.  Lower inventory levels mean available Pu is being utilized more
+efficiently rather than sitting idle.  Cases QI and QF with their longer time
+steps show higher inventories indicating a more severe shortage.  Cases QI and
+QF take longer to recover from the shortage and fall a bit behind with respect
+to building up Pu stocks due to the delayed contributions of more unfueled
+reactors to the separated Pu pool.  This can be seen most clearly in Figure
+\ref{fig:puinv-compare-zoom}.  Cases MI and QF recover Pu inventory
+post-shortage at about the same time and rate - this is consistent with their
+approximately equivalent cumulative offline power capacity curves in Figure
+\ref{fig:unfueled}.
 
@@
         fuel shortages begin.  These differences gradually decay away starting
-        in year 140 after fuel shortages have ebbed.  Higher aggregate Pu
-        inventory  during the shortage is correlated with both a longer shortage and
-        a longer post-shortage recovery period.
+        in year 140 after fuel shortages have ebbed.
     }
@@
 \end{figure*}
+
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare-zoom.eps}
+    \caption{
+        Separated Pu inventories for all four cases zoomed in on the fuel
+        shortage years.  Higher aggregate Pu inventory  during the shortage is
+        correlated with both a longer shortage and a slower post-shortage
+        recovery of Pu inventory.
+    }
+    \label{fig:puinv-compare-zoom}
+\end{figure*}
+
+The reactors in all the scenarios request fresh fuel only when they need it.
+While this is an unrealistic behavior for real-world facilities, it is closer
+to a globally optimum fuel management strategy.  Keeping on-hand fresh fuel
+batches would also have the same effect of increasing idling plutonium
+quantities and would make the fuel shortage worse overall.
 

### 2015-10-07 paper.tex
fix typo
add some visual queues to avoid the need for explaining naming scheme

@@
 reactor modeling.  The difference between these two modeling styles in terms
-of their affects on output of fuel cycle simulators in general is not well
+of their effects on output of fuel cycle simulators in general is not well
 known. This analysis is designed to provide both quantitative and
@@
 
-    \item \emph{Case MI}: Monthly time steps with individual reactor modeling
-    \item \emph{Case MF}: Monthly time steps with fleet reactor modeling
-    \item \emph{Case QI}: Quarterly (3-month) time steps with individual reactor modeling
-    \item \emph{Case QF}: Quarterly (3-month) time steps with fleet reactor modeling
+    \item \emph{Case MI}: \underline{M}onthly time steps with \underline{I}ndividual reactor modeling
+    \item \emph{Case MF}: \underline{M}onthly time steps with \underline{F}leet reactor modeling
+    \item \emph{Case QI}: \underline{Q}uarterly (3-month) time steps with \underline{I}ndividual reactor modeling
+    \item \emph{Case QF}: \underline{Q}uarterly (3-month) time steps with \underline{F}leet reactor modeling
 
 \end{enumerate}
-
-The case labels consist of two letters each: "M" for monthly time steps, "Q"
-for quarterly time steps, "I" for individual reactors, and "F" for fleet
-reactors.
 

### 2015-10-07 paper.tex
context to need for fleet reactor model

@@
 
-A fleet reactor archetype was created for Cyclus.  The flexibility of Cyclus
+A fleet reactor archetype was created for Cyclus.  Although the modeling of
+individual, discrete facilities was a motivation for Cyclus, its flexibility
 makes it relatively straight forward to create plug-in archetypes to model

### 2015-10-07 paper.tex
fix typo

--- 2015-10-07 paper.tex
@@
 
-A fleet reactor archetype was created for Cyclus.  The flexibility of Cyclus
-makes it relatively straight forward to create plug-in archetypes to model
+A fleet reactor archetype was created for Cyclus.  Although the modeling of
+individual, discrete facilities was a motivation for Cyclus, its flexibility
+makes it relatively straightforward to create plug-in archetypes to model
 anywhere along the spectrum from individual, discrete facilities to

### 2015-10-07 paper.tex
fix typo

--- 2015-10-07 paper.tex
@@
 individual, discrete facilities was a motivation for Cyclus, its flexibility
-makes it relatively straight forward to create plug-in archetypes to model
+makes it relatively straightforward to create plug-in archetypes to model
 anywhere along the spectrum from individual, discrete facilities to

### 2015-10-07 paper.tex
word choice

--- 2015-10-07 paper.tex
@@
 anywhere along the spectrum from individual, discrete facilities to
-continuous flow fleet facilities.  In order to contrast well with individual,
-single-facility granularity modeling, the fleet archetype created here falls more
-toward the aggregate, continuous flow side of the spectrum.
+continuous flow fleet facilities.  In order to contrast well with the individual,
+single-facility granularity modeling that is part of the standard Cyclus archetypes, 
+a fleet archetype was created for this work with aggregate facilitiy behavior and
+continuous flow.
 

### 2015-10-07 paper.tex
Introduce variable for number of operating reactors to simplify discussion.

--- 2015-10-07 paper.tex
@@
 
-    \item Reactor capacity is deployed and retired in single reactor units
-        (e.g. 1 GWe LWRs).
+    \item Reactor capacity is deployed and retired in increments of single
+      reactor units, each with capable of producing power $P_r$ (e.g. 1 GWe
+      LWRs).  $N_{r}$ reactors are deployed at any time capable of producing
+      $P_r \cdot N_r$ power.
+
+    \item In the event of fuel shortages, the number of reactors that can
+      operate is modeled as
+
+      \begin{equation}
+        N_{o} = N_{r} \cdot \frac{C_{inv}}{C_{cap}},
+        \label{eqn:operating-reactors}
+      \end{equation}
+
+       where $C_{inv}$ is the total amount of fuel in all reactor cores (kg),
+       and $C_{cap}$ is the total fuel capacity for all reactor cores (kg).
 
@@
             D =
-            \frac{B}{L} \cdot N_{r} \cdot \frac{C_{inv}}{C_{cap}}
+            \frac{B}{L} \cdot N_{o} =  \frac{B}{L} \cdot N_{r}\cdot \frac{C_{inv}}{C_{cap}}
             \label{eqn:fleet-discharge}
@@
         (kg) for a single reactor's batch, $L$ is the cycle length (including
-        refueling time) in time steps, $N_{r}$ is the number of reactors in
-        the fleet, $C_{inv}$ is the total amount of fuel in all reactor cores
-        (kg), and $C_{cap}$ is the total fuel capacity for all reactor cores
-        (kg).  The last fractional term accounts for lower fuel burning rates
-        due to unfueled reactors.
+        refueling time) in time steps.  In the event of fuel shortages, the
+        discharge rate is reduced because some reactors are not operating.
 
@@
 
-    \item Generated power is computed similarly to core discharge rates by
-        multiplying the total fleet power capacity by the same last fractional
-        term of Equation \ref{eqn:fleet-discharge} to account for unfueled
-        capacity in the fleet.
+    \item Generated power at any timestep, $P_G$, is based on the number of
+      operating reactors:
+
+      \begin{equation}
+        P_{G} = P_r \cdot N_o = P_r \cdot N_r \cdot \frac{C_{inv}}{C_{cap}}
+      \end{equation}
 

### 2015-10-07 paper.tex
Remove postpositive past participle acting as adjective

@@
 
-The scenario modeled approximates a transition from a the current once-through light water
+The chosen scenario approximates a transition from a the current once-through light water
 reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle within evaluation group 23 (EG-23)

### 2015-10-07 paper.tex
Add comment on flow diagram
Add comment on referencing ongoing FCO work
tweak discussion of standard archetypes

--- 2015-10-07 paper.tex
@@
 after ongoing work by the DOE in their Fuel Cycle Options campaign.
+%% Not sure if we should make such an explicit reference to the ongoing FCO
+%% transition work until they publish something. On one hand, it reveals
+%% things they are secretive about. On the other hand, we should give them due
+%% credit.
 
@@
     \caption[Material flow paths]{Potential material flow paths between facility types.}
+    %% consider a Cyclist-derived visualization here?
     \label{fig:flow}
@@
 
-A pre-release of Cyclus version 1.4 was used for this analysis.  Archetypes
-used were the standard ones provided with Cyclus in the included Cycamore
+A pre-release of Cyclus version 1.4 was used for this analysis, with the
+standard archetypes provided with Cyclus in the included Cycamore
 library - also a pre-release of version 1.4.  A custom storage archetype was
-used because Cycamore did not yet provide one.  The standard cycamore reactor archetype
+developed because Cycamore did not yet provide one.  The standard cycamore reactor archetype
 was used for both LWR and SFR reactor types in cases MI and QI.  Correspondingly

### 2015-10-07 paper.tex
comment on 1.4 and release
Cycamore capitalization
comment on Table I, SFR cycle length

--- 2015-10-07 paper.tex
@@
 
+%% is it worth waiting until the 1.4 release?  Citing an ambiguous pre-release makes
+%% this a little harder to reproduce.
 A pre-release of Cyclus version 1.4 was used for this analysis, with the
@@
 library - also a pre-release of version 1.4.  A custom storage archetype was
-developed because Cycamore did not yet provide one.  The standard cycamore reactor archetype
+developed because Cycamore did not yet provide one.  The standard Cycamore reactor archetype
 was used for both LWR and SFR reactor types in cases MI and QI.  Correspondingly
@@
         Lifetime (yr)         & 80 & 80 \\
-        Cycle length (months) & 18 & 14 \\
+        Cycle length (months) & 18 & 14 \\  %% why list this SFR cycle length if it's never used?
         Batch size (kg)       & 29565 & 7490 \\
@@
 cases.  However, a few small case-dependent changes to the scenarios relating to
-reactor configuration are necessary.  For the cases using quarterly time steps
-with individual reactors, cycle and outage times must be adjusted to be a
+reactor configuration are necessary.  For case \emph{QI}, 
+cycle and outage times must be adjusted to be a
 multiple of 3 months.  In order to make the comparison more valid, the cases

### 2015-10-07 paper.tex
Add comment on tweaking cycles to multiples of 3
Fix LWR fleet batch size = core size

@@
 cycle and outage times must be adjusted to be a
-multiple of 3 months.  In order to make the comparison more valid, the cases
+multiple of 3 months. 
+%% this tweak is a distraction.  Why not simply declare the SFR cycle length &
+%% outage as 12/3 from the beginning.  Is this because the fuel recipes you
+%% are using assume the other parameters and you don't want to sweep that
+%% under the rug?
+ In order to make the comparison more valid, the cases
 with monthly time steps are also adjusted to have identical cycle and outage
@@
         Refueling outage (months)         & 3         & 0                 & 3         & 0 \\
-        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 29,565            & 8,025     & 40,125 \\
+        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 88,695            & 8,025     & 40,125 \\
         Power capacity (MWe)              & 1,080     & 900               & 450       & 360 \\

### 2015-10-07 paper.tex
All capacity factors are much closer to 80%
Update notation for consistency

--- 2015-10-07 paper.tex
@@
 parameters for each case were also selected to keep capacity factors for
-individual reactors very roughly around 90\%.
+individual reactors very roughly around 80\%.
 
@@
 \emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
-cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot C$
-of power for that time step where $C$ is the power capacity of a single
+cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot P_r$
+of power for that time step where $P_r$ is the power capacity of a single
 reactor of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is
-short a half batch of fuel, then it will produce no power ($0 \cdot C$) that
+short a half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
 time step.  Time step duration can also have this effect.  For both fleet and
@@
 fuel with a 3 month time step, then it will produce $\frac{2.5}{3} \cdot 3
-\text{mo} \cdot C $ of energy.  In contrast, a monthly time step might allow
+\text{mo} \cdot P_r$ of energy.  In contrast, a monthly time step might allow
 the same reactor to produce $[\frac{5}{6} \cdot 1 \text{mo} + 1 \cdot 2
-\text{mo}] \cdot C $ of energy if the fuel shortage actually ended after only
+\text{mo}] \cdot P_r $ of energy if the fuel shortage actually ended after only
 1 month.  Quantifying this effect directly is difficult.  One way of observing

### 2015-10-07 paper.tex
Consistent fractions
Bind units to number

--- 2015-10-07 paper.tex
@@
 of power for that time step where $P_r$ is the power capacity of a single
-reactor of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is
+reactor of the fleet (\emph{e.g.}~1000~MWe).  If an individually modeled reactor is
 short a half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
@@
 \text{mo} \cdot P_r$ of energy.  In contrast, a monthly time step might allow
-the same reactor to produce $[\frac{5}{6} \cdot 1 \text{mo} + 1 \cdot 2
+the same reactor to produce $[\frac{2.5}{3} \cdot 1 \text{mo} + 1 \cdot 2
 \text{mo}] \cdot P_r $ of energy if the fuel shortage actually ended after only

### 2015-10-07 paper.tex
timestep doesn't introduce quantized shutdown but amplifies it
Expand example to list form

@@
 short a half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
-time step.  Time step duration can also have this effect.  For both fleet and
+time step.  Time step duration can amplify this effect.  For both fleet and
 individual reactor modeling, power goes offline in full time step increments.
-For example, if a 3-batch fleet based reactor is short $\frac{1}{2}$ batch of
-fuel with a 3 month time step, then it will produce $\frac{2.5}{3} \cdot 3
-\text{mo} \cdot P_r$ of energy.  In contrast, a monthly time step might allow
-the same reactor to produce $[\frac{2.5}{3} \cdot 1 \text{mo} + 1 \cdot 2
-\text{mo}] \cdot P_r $ of energy if the fuel shortage actually ended after only
-1 month.  Quantifying this effect directly is difficult.  One way of observing
+Consider an example of a 3-batch reactor with a fuel shortage of $\frac{1}{2}$ batch of
+fuel that would be resolve in 1 month.  Depending on the modeling choice, each reactor
+would produce the following amounts of energy over a 3 month period:
+
+\begin{itemize}
+  \item case \emph{QF}: $\frac{2.5}{3} \cdot 3 \text{mo} \cdot P_r = 2.5P_r$
+  \item case \emph{MF}: $[\frac{2.5}{3} \cdot 1 \text{mo} + 1 \cdot 2
+    \text{mo}] \cdot P_r = 2.83P_r$
+  \item case \emph{QI}: $0 \cdot 3 \text{mo} \cdot P_r = 0$
+  \item case \emph{MI}: $[0\cdot 1 \text{mo} + 1 \cdot 2\text{mo}] \cdot P_r = 2 P_r$
+\end{itemize}
+
+Quantifying this effect directly is difficult.  One way of observing
 it is to compare fleet reactor scenarios that have different time step

### 2015-10-07 paper.tex
Refine discussion of waves of reactors

@@
 there are corresponding waves of deployment, and these waves are echoed every
-80 years (the reactor lifetime).  These waves of deployment cause waves in
-recycled fuel availability that cause many of the reactors offline during the
+80 years (the reactor lifetime).  These waves of deployment cause surges in
+recycled fuel availability when they begin to discharge their fuel
+%% these need to be LWR waves, right, or they'd just make things worse?
+that cause many of the reactors offline during the
 shortage to receive fuel and come online together.  The net effect is that the

### 2015-10-13 paper.tex
Add comment on how to measure fuel sharing effect
add equations for shortage and fuel sharing

@@
     \item Groups of facilities (usually all facilities of a given type) are
-        treated as a single, aggregate entity. Therefore, all facilities experience
-        identical operational histories.
+        treated as a single, aggregate entity.
 
-    \item Resource flows are modeled as continuous, within the limitations of time
-        step discretization.  Therefore, material is acquired and discharge in
-        uniform amounts at every time step.
+    \item Resource flows are often (approximately) continuous (limited by time
+        step discretization).
 
@@
 reactor modeling.  The difference between these two modeling styles in terms
-of their effects on output of fuel cycle simulators in general is not well
+of their affects on output of fuel cycle simulators in general is not well
 known. This analysis is designed to provide both quantitative and
@@
 
-    \item \emph{Case MI}: \underline{M}onthly time steps with \underline{I}ndividual reactor modeling
-    \item \emph{Case MF}: \underline{M}onthly time steps with \underline{F}leet reactor modeling
-    \item \emph{Case QI}: \underline{Q}uarterly (3-month) time steps with \underline{I}ndividual reactor modeling
-    \item \emph{Case QF}: \underline{Q}uarterly (3-month) time steps with \underline{F}leet reactor modeling
+    \item \emph{Case MI}: Monthly time steps with individual reactor modeling
+    \item \emph{Case MF}: Monthly time steps with fleet reactor modeling
+    \item \emph{Case QI}: Quarterly (3-month) time steps with individual reactor modeling
+    \item \emph{Case QF}: Quarterly (3-month) time steps with fleet reactor modeling
 
 \end{enumerate}
+
+The case labels consist of two letters each: "M" for monthly time steps, "Q"
+for quarterly time steps, "I" for individual reactors, and "F" for fleet
+reactors.
 
@@
 
-A fleet reactor archetype was created for Cyclus.  Although the modeling of
-individual, discrete facilities was a motivation for Cyclus, its flexibility
-makes it relatively straightforward to create plug-in archetypes to model
+A fleet reactor archetype was created for Cyclus.  The flexibility of Cyclus
+makes it relatively straight forward to create plug-in archetypes to model
 anywhere along the spectrum from individual, discrete facilities to
-continuous flow fleet facilities.  In order to contrast well with the individual,
-single-facility granularity modeling that is part of the standard Cyclus archetypes, 
-a fleet archetype was created for this work with aggregate facilitiy behavior and
-continuous flow.
+continuous flow fleet facilities.  In order to contrast well with individual,
+single-facility granularity modeling, the fleet archetype created here falls more
+toward the aggregate, continuous flow side of the spectrum.
 
@@
 
-    \item Reactor capacity is deployed and retired in increments of single
-      reactor units, each with capable of producing power $P_r$ (e.g. 1 GWe
-      LWRs).  $N_{r}$ reactors are deployed at any time capable of producing
-      $P_r \cdot N_r$ power.
-
-    \item In the event of fuel shortages, the number of reactors that can
-      operate is modeled as
-
-      \begin{equation}
-        N_{o} = N_{r} \cdot \frac{C_{inv}}{C_{cap}},
-        \label{eqn:operating-reactors}
-      \end{equation}
-
-       where $C_{inv}$ is the total amount of fuel in all reactor cores (kg),
-       and $C_{cap}$ is the total fuel capacity for all reactor cores (kg).
+    \item Reactor capacity is deployed and retired in single reactor units
+        (e.g. 1 GWe LWRs).
 
@@
             D =
-            \frac{B}{L} \cdot N_{o} =  \frac{B}{L} \cdot N_{r}\cdot \frac{C_{inv}}{C_{cap}}
+            \frac{B}{L} \cdot N_{r} \cdot \frac{C_{inv}}{C_{cap}}
             \label{eqn:fleet-discharge}
@@
         (kg) for a single reactor's batch, $L$ is the cycle length (including
-        refueling time) in time steps.  In the event of fuel shortages, the
-        discharge rate is reduced because some reactors are not operating.
+        refueling time) in time steps, $N_{r}$ is the number of reactors in
+        the fleet, $C_{inv}$ is the total amount of fuel in all reactor cores
+        (kg), and $C_{cap}$ is the total fuel capacity for all reactor cores
+        (kg).  The last fractional term accounts for lower fuel burning rates
+        due to unfueled reactors.
 
@@
 
-    \item Generated power at any timestep, $P_G$, is based on the number of
-      operating reactors:
-
-      \begin{equation}
-        P_{G} = P_r \cdot N_o = P_r \cdot N_r \cdot \frac{C_{inv}}{C_{cap}}
-      \end{equation}
+    \item Generated power is computed similarly to core discharge rates by
+        multiplying the total fleet power capacity by the same last fractional
+        term of Equation \ref{eqn:fleet-discharge} to account for unfueled
+        capacity in the fleet.
 
@@
 
-The chosen scenario approximates a transition from a the current once-through light water
+The scenario modeled approximates a transition from a the current once-through light water
 reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle within evaluation group 23 (EG-23)
@@
 after ongoing work by the DOE in their Fuel Cycle Options campaign.
-%% Not sure if we should make such an explicit reference to the ongoing FCO
-%% transition work until they publish something. On one hand, it reveals
-%% things they are secretive about. On the other hand, we should give them due
-%% credit.
 
@@
     \caption[Material flow paths]{Potential material flow paths between facility types.}
-    %% consider a Cyclist-derived visualization here?
     \label{fig:flow}
@@
 
-%% is it worth waiting until the 1.4 release?  Citing an ambiguous pre-release makes
-%% this a little harder to reproduce.
-A pre-release of Cyclus version 1.4 was used for this analysis, with the
-standard archetypes provided with Cyclus in the included Cycamore
+A pre-release of Cyclus version 1.4 was used for this analysis.  Archetypes
+used were the standard ones provided with Cyclus in the included Cycamore
 library - also a pre-release of version 1.4.  A custom storage archetype was
-developed because Cycamore did not yet provide one.  The standard Cycamore reactor archetype
+used because Cycamore did not yet provide one.  The standard cycamore reactor archetype
 was used for both LWR and SFR reactor types in cases MI and QI.  Correspondingly
@@
         Lifetime (yr)         & 80 & 80 \\
-        Cycle length (months) & 18 & 14 \\  %% why list this SFR cycle length if it's never used?
+        Cycle length (months) & 18 & 14 \\
         Batch size (kg)       & 29565 & 7490 \\
@@
 cases.  However, a few small case-dependent changes to the scenarios relating to
-reactor configuration are necessary.  For case \emph{QI}, 
-cycle and outage times must be adjusted to be a
-multiple of 3 months. 
-%% this tweak is a distraction.  Why not simply declare the SFR cycle length &
-%% outage as 12/3 from the beginning.  Is this because the fuel recipes you
-%% are using assume the other parameters and you don't want to sweep that
-%% under the rug?
- In order to make the comparison more valid, the cases
+reactor configuration are necessary.  For the cases using quarterly time steps
+with individual reactors, cycle and outage times must be adjusted to be a
+multiple of 3 months.  In order to make the comparison more valid, the cases
 with monthly time steps are also adjusted to have identical cycle and outage
@@
         Refueling outage (months)         & 3         & 0                 & 3         & 0 \\
-        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 88,695            & 8,025     & 40,125 \\
+        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 29,565            & 8,025     & 40,125 \\
         Power capacity (MWe)              & 1,080     & 900               & 450       & 360 \\
@@
 parameters for each case were also selected to keep capacity factors for
-individual reactors very roughly around 80\%.
+individual reactors very roughly around 90\%.
 
@@
 \emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
-cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot P_r$
-of power for that time step where $P_r$ is the power capacity of a single
-reactor of the fleet (\emph{e.g.}~1000~MWe).  If an individually modeled reactor is
-short a half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
-time step.  Time step duration can amplify this effect.  For both fleet and
+cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot C$
+of power for that time step where $C$ is the power capacity of a single
+reactor of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is
+short a half batch of fuel, then it will produce no power ($0 \cdot C$) that
+time step.  Time step duration can also have this effect.  For both fleet and
 individual reactor modeling, power goes offline in full time step increments.
-Consider an example of a 3-batch reactor with a fuel shortage of $\frac{1}{2}$ batch of
-fuel that would be resolve in 1 month.  Depending on the modeling choice, each reactor
-would produce the following amounts of energy over a 3 month period:
-
-\begin{itemize}
-  \item case \emph{QF}: $\frac{2.5}{3} \cdot 3 \text{mo} \cdot P_r = 2.5P_r$
-  \item case \emph{MF}: $[\frac{2.5}{3} \cdot 1 \text{mo} + 1 \cdot 2
-    \text{mo}] \cdot P_r = 2.83P_r$
-  \item case \emph{QI}: $0 \cdot 3 \text{mo} \cdot P_r = 0$
-  \item case \emph{MI}: $[0\cdot 1 \text{mo} + 1 \cdot 2\text{mo}] \cdot P_r = 2 P_r$
-\end{itemize}
-
-Quantifying this effect directly is difficult.  One way of observing
+For example, if a 3-batch fleet based reactor is short $\frac{1}{2}$ batch of
+fuel with a 3 month time step, then it will produce $\frac{2.5}{3} \cdot 3
+\text{mo} \cdot C $ of energy.  In contrast, a monthly time step might allow
+the same reactor to produce $[\frac{5}{6} \cdot 1 \text{mo} + 1 \cdot 2
+\text{mo}] \cdot C $ of energy if the fuel shortage actually ended after only
+1 month.  Quantifying this effect directly is difficult.  One way of observing
 it is to compare fleet reactor scenarios that have different time step
@@
 there are corresponding waves of deployment, and these waves are echoed every
-80 years (the reactor lifetime).  These waves of deployment cause surges in
-recycled fuel availability when they begin to discharge their fuel
-%% these need to be LWR waves, right, or they'd just make things worse?
-that cause many of the reactors offline during the
+80 years (the reactor lifetime).  These waves of deployment cause waves in
+recycled fuel availability that cause many of the reactors offline during the
 shortage to receive fuel and come online together.  The net effect is that the

### 2015-10-20 paper.tex
Comment on functional form
add some commas
tmp - adding zoom figs

@@
 
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/puinv-MI.eps}
+    \caption[Separated Pu inventory and flow: Case MI Zoom]{
+        TODO
+    }
+    \label{fig:puinv-MI}
+\end{figure*}
+
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/puinv-MF.eps}
+    \caption[Separated Pu inventory and flow: Case MF Zoom]{
+        TODO
+    }
+    \label{fig:puinv-MF}
+\end{figure*}
+
 Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
@@
     \label{fig:puinv-compare}
+\end{figure*}
+
+\begin{figure*}[!h]
+    \centering
+    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare-zoom.eps}
+    \caption{
+        TODO
+    }
+    \label{fig:puinv-compare-zoom}
 \end{figure*}

### 2015-10-20 paper.tex
add figure captions and refs in text. zoom in on sync-cycle fig

@@
     \caption[Separated Pu inventory and flow: Case MI Zoom]{
-        TODO
+        Separated Pu inventory and flows for case MI.  At locations where the
+        outflow dots lie on/near the inventory level, fuel shortages are
+        taking place.  This begins around year 105 and continues until about
+        year 140 where Pu withdrawals become lower than standing inventory.
+        The horizontal bands in out-flows starting around year 140 are caused
+        by many previously offline reactors coming online together in groups.
+        These fuel inventory level resonances gradually disappear as the
+        simulation progresses.
     }
@@
     \caption[Separated Pu inventory and flow: Case MF Zoom]{
-        TODO
+        Separated Pu inventory and flows for case MF.  At locations where the
+        outflow dots meet the inventory level, fuel shortages are taking
+        place.  This begins around year 105 and continues until about year 140
+        where Pu withdrawals become lower than standing inventory.  The
+        periodic pulsing that begins just after year 120 is caused by the
+        first fast reactor retirements; the first fast reactors deployed starting in
+        year 35 retire annually after their 80 year lifetime and discharge an entire core's
+        worth of fuel rather than the usual single batch for refueling.
     }
@@
 Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
-\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated plutonium
-inventory during the shortage generally indicates fewer unfueled reactors.
-Lower inventory levels mean available Pu is being utilized more efficiently
-rather than sitting idle.  Cases QI and QF with their longer time steps show
-higher inventories indicating a more severe shortage.  Post shortage, cases QI and QF
-also fall a bit behind with respect to building up Pu stocks due to the delayed
-contributions of more unfueled reactors to the separated Pu pool.  The
-reactors in all the scenarios request fresh fuel only when they need it. While
-this is an unrealistic behavior for real-world facilities, it is closer to a
-globally optimum fuel management strategy.  Keeping on-hand fresh fuel batches
-would also have the same effect of increasing idling plutonium quantities and
-would make the fuel shortage worse overall.
+\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated
+plutonium inventory during the shortage generally indicates fewer unfueled
+reactors.  Lower inventory levels mean available Pu is being utilized more
+efficiently rather than sitting idle.  Cases QI and QF with their longer time
+steps show higher inventories indicating a more severe shortage.  Cases QI and
+QF take longer to recover from the shortage and fall a bit behind with respect
+to building up Pu stocks due to the delayed contributions of more unfueled
+reactors to the separated Pu pool.  This can be seen most clearly in Figure
+\ref{fig:puinv-compare-zoom}.  Cases MI and QF recover Pu inventory
+post-shortage at about the same time and rate - this is consistent with their
+approximately equivalent cumulative offline power capacity curves in Figure
+\ref{fig:unfueled}.
 
@@
     \caption{
-        TODO
+        Separated Pu inventories for all four cases zoomed in on the fuel
+        shortage years.  Higher aggregate Pu inventory  during the shortage is
+        correlated with both a longer shortage and a slower post-shortage
+        recovery of Pu inventory.
     }
@@
 \end{figure*}
+
+The reactors in all the scenarios request fresh fuel only when they need it.
+While this is an unrealistic behavior for real-world facilities, it is closer
+to a globally optimum fuel management strategy.  Keeping on-hand fresh fuel
+batches would also have the same effect of increasing idling plutonium
+quantities and would make the fuel shortage worse overall.
 

### 2015-10-20 paper.tex

@@
     \item Groups of facilities (usually all facilities of a given type) are
-        treated as a single, aggregate entity.
-
-    \item Resource flows are often (approximately) continuous (limited by time
-        step discretization).
+        treated as a single, aggregate entity. Therefore, all facilities experience
+        identical operational histories.
+
+    \item Resource flows are modeled as continuous, within the limitations of time
+        step discretization.  Therefore, material is acquired and discharge in
+        uniform amounts at every time step.
 
@@
 
-A fleet reactor archetype was created for Cyclus.  The flexibility of Cyclus
-makes it relatively straight forward to create plug-in archetypes to model
+A fleet reactor archetype was created for Cyclus.  Although the modeling of
+individual, discrete facilities was a motivation for Cyclus, its flexibility
+makes it relatively straightforward to create plug-in archetypes to model
 anywhere along the spectrum from individual, discrete facilities to
-continuous flow fleet facilities.  In order to contrast well with individual,
-single-facility granularity modeling, the fleet archetype created here falls more
-toward the aggregate, continuous flow side of the spectrum.
+continuous flow fleet facilities.  In order to contrast well with the individual,
+single-facility granularity modeling that is part of the standard Cyclus archetypes, 
+a fleet archetype was created for this work with aggregate facilitiy behavior and
+continuous flow.
 
@@
 
-A pre-release of Cyclus version 1.4 was used for this analysis.  Archetypes
-used were the standard ones provided with Cyclus in the included Cycamore
+%% is it worth waiting until the 1.4 release?  Citing an ambiguous pre-release makes
+%% this a little harder to reproduce.
+A pre-release of Cyclus version 1.4 was used for this analysis, with the
+standard archetypes provided with Cyclus in the included Cycamore
 library - also a pre-release of version 1.4.  A custom storage archetype was
-used because Cycamore did not yet provide one.  The standard cycamore reactor archetype
+developed because Cycamore did not yet provide one.  The standard Cycamore reactor archetype
 was used for both LWR and SFR reactor types in cases MI and QI.  Correspondingly
@@
 cases.  However, a few small case-dependent changes to the scenarios relating to
-reactor configuration are necessary.  For the cases using quarterly time steps
-with individual reactors, cycle and outage times must be adjusted to be a
-multiple of 3 months.  In order to make the comparison more valid, the cases
+reactor configuration are necessary.  For case \emph{QI}, 
+cycle and outage times must be adjusted to be a
+multiple of 3 months. 
+%% this tweak is a distraction.  Why not simply declare the SFR cycle length &
+%% outage as 12/3 from the beginning.  Is this because the fuel recipes you
+%% are using assume the other parameters and you don't want to sweep that
+%% under the rug?
+ In order to make the comparison more valid, the cases
 with monthly time steps are also adjusted to have identical cycle and outage
@@
 \emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
-cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot C$
-of power for that time step where $C$ is the power capacity of a single
-reactor of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is
-short a half batch of fuel, then it will produce no power ($0 \cdot C$) that
-time step.  Time step duration can also have this effect.  For both fleet and
+cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot P_r$
+of power for that time step where $P_r$ is the power capacity of a single
+reactor of the fleet (\emph{e.g.}~1000~MWe).  If an individually modeled reactor is
+short a half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
+time step.  Time step duration can amplify this effect.  For both fleet and
 individual reactor modeling, power goes offline in full time step increments.
-For example, if a 3-batch fleet based reactor is short $\frac{1}{2}$ batch of
-fuel with a 3 month time step, then it will produce $\frac{2.5}{3} \cdot 3
-\text{mo} \cdot C $ of energy.  In contrast, a monthly time step might allow
-the same reactor to produce $[\frac{5}{6} \cdot 1 \text{mo} + 1 \cdot 2
-\text{mo}] \cdot C $ of energy if the fuel shortage actually ended after only
-1 month.  Quantifying this effect directly is difficult.  One way of observing
+Consider an example of a 3-batch reactor with a fuel shortage of $\frac{1}{2}$ batch of
+fuel that would be resolve in 1 month.  Depending on the modeling choice, each reactor
+would produce the following amounts of energy over a 3 month period:
+
+\begin{itemize}
+  \item case \emph{QF}: $\frac{2.5}{3} \cdot 3 \text{mo} \cdot P_r = 2.5P_r$
+  \item case \emph{MF}: $[\frac{2.5}{3} \cdot 1 \text{mo} + 1 \cdot 2
+    \text{mo}] \cdot P_r = 2.83P_r$
+  \item case \emph{QI}: $0 \cdot 3 \text{mo} \cdot P_r = 0$
+  \item case \emph{MI}: $[0\cdot 1 \text{mo} + 1 \cdot 2\text{mo}] \cdot P_r = 2 P_r$
+\end{itemize}
+
+Quantifying this effect directly is difficult.  One way of observing
 it is to compare fleet reactor scenarios that have different time step
@@
 \end{equation}
+%% What is the value of this function during a normal refueling outage,
+%% immediately before S_(sched)(t)?  It seems you need another Heavyside
+%% function to drop this back down to an expected value of 0 again.
+%%
+%%          ___________________
+%%          |                 |
+%%  ________|                 |_____________
 
@@
 showing wasted batch-months computed using Equation \ref{eq:badshare}.  Fleet
-reactor modeling by design exhibits perfect fuel sharing by design (zero
+reactor modeling, by design, exhibits perfect fuel sharing by design (zero
 inefficiency) and so is not included in this figure.  Every batch that is
@@
 time step the cycle start is delayed.
+
+%% What about a different measure of fuel sharing effect?  Perfect sharing
+%% in a fleet reactor results in N_o reactors operating.  For individual
+%% reactors, that would mean floor(N_o) reactors.  Can't we measure the
+%% actual number of reactors that are operating, N_actual, and accumulate
+%% floor(N_o) - N_actual as a measure of the effect of fuel sharing?  Would
+%% this be substantially different?
 

### 2015-10-20 paper.tex
clean up @gonuke's edits

@@
         treated as a single, aggregate entity. Therefore, all facilities experience
-        identical operational histories.
+        a single, identical history.
 
     \item Resource flows are modeled as continuous, within the limitations of time
-        step discretization.  Therefore, material is acquired and discharge in
+        step discretization.  Therefore, material is acquired and discharged in
         uniform amounts at every time step.
@@
     \item Reactor capacity is deployed and retired in increments of single
-      reactor units, each with capable of producing power $P_r$ (e.g. 1 GWe
+      reactor units, each capable of producing power $P_r$ (e.g. 1 GWe
       LWRs).  $N_{r}$ reactors are deployed at any time capable of producing
-      $P_r \cdot N_r$ power.
+      up to $P_r \cdot N_r$ power.
 
@@
         Lifetime (yr)         & 80 & 80 \\
-        Cycle length (months) & 18 & 14 \\  %% why list this SFR cycle length if it's never used?
+        Cycle length (months) & 18 & 15 \\
         Batch size (kg)       & 29565 & 7490 \\
@@
 cases.  However, a few small case-dependent changes to the scenarios relating to
-reactor configuration are necessary.  For case \emph{QI}, 
-cycle and outage times must be adjusted to be a
-multiple of 3 months. 
-%% this tweak is a distraction.  Why not simply declare the SFR cycle length &
-%% outage as 12/3 from the beginning.  Is this because the fuel recipes you
-%% are using assume the other parameters and you don't want to sweep that
-%% under the rug?
- In order to make the comparison more valid, the cases
-with monthly time steps are also adjusted to have identical cycle and outage
-times.  The invariants preserved with respect to reactor behavior between all
+reactor configuration are necessary.
+The invariants preserved with respect to reactor behavior between all
 four cases are shown in Table \ref{tab:invar}.  Table \ref{tab:reactor-detail} shows
@@
     \captionsetup{justification=centering}
-    \caption{Reactor Parameters by Case}
+    \caption{Selected Reactor Parameters by Case}
     \begin{tabular}{ |r | c c | c c | }
@@
         Refueling outage (months)         & 3         & 0                 & 3         & 0 \\
-        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 88,695            & 8,025     & 40,125 \\
+        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 29,565            & 8,025     & 8,025 \\
         Power capacity (MWe)              & 1,080     & 900               & 450       & 360 \\
@@
 parameters for each case were also selected to keep capacity factors for
-individual reactors very roughly around 80\%.
+individual reactors very roughly around 90\%, although they are actually 
+closer to 80\% due to integer variable constraints on the case MI and QI
+reactor configuration.
 
@@
 reactor of the fleet (\emph{e.g.}~1000~MWe).  If an individually modeled reactor is
-short a half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
+short one half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
 time step.  Time step duration can amplify this effect.  For both fleet and
@@
 \begin{itemize}
-  \item case \emph{QF}: $\frac{2.5}{3} \cdot 3 \text{mo} \cdot P_r = 2.5P_r$
-  \item case \emph{MF}: $[\frac{2.5}{3} \cdot 1 \text{mo} + 1 \cdot 2
+  \item case \emph{QF}: $\frac{2.5}{3} \cdot 3 \text{ mo} \cdot P_r = 2.5P_r$
+  \item case \emph{MF}: $[\frac{2.5}{3} \cdot 1 \text{ mo} + 1 \cdot 2
     \text{mo}] \cdot P_r = 2.83P_r$
-  \item case \emph{QI}: $0 \cdot 3 \text{mo} \cdot P_r = 0$
-  \item case \emph{MI}: $[0\cdot 1 \text{mo} + 1 \cdot 2\text{mo}] \cdot P_r = 2 P_r$
+  \item case \emph{QI}: $0 \cdot 3 \text{ mo} \cdot P_r = 0$
+  \item case \emph{MI}: $[0\cdot 1 \text{ mo} + 1 \cdot 2\text{mo}] \cdot P_r = 2 P_r$
 \end{itemize}
@@
 showing wasted batch-months computed using Equation \ref{eq:badshare}.  Fleet
-reactor modeling, by design, exhibits perfect fuel sharing by design (zero
+reactor modeling, by design, exhibits perfect fuel sharing (zero
 inefficiency) and so is not included in this figure.  Every batch that is

### 2015-10-20 paper.tex
add github PR comments for provenance
spell check

@@
 individual reactors with either monthly or quarterly (3-month) time steps.
-The Cyclus fuel cycle simulator's plug-in facility cability along with its
+The Cyclus fuel cycle simulator's plug-in facility capability along with its
 market-like dynamic material routing allow it to be used as a level playing
 field for comparing the scenarios.  When under supply-constraint pressure, the
-four cases exhibit noticably different behavior.  Fleet-based modeling is more
+four cases exhibit noticeably different behavior.  Fleet-based modeling is more
 efficient in supply-constrained environments at the expense of losing insight
@@
 duration.  A better understanding of how these affect fuel cycle dynamics can
-allow for making more deliberate decisions with respect to tradeoffs such as
+allow for making more deliberate decisions with respect to trade-offs such as
 computational investment vs. realism.
@@
 more expressive than more traditional fuel cycle simulators.  Cyclus was
-designed with a plugin-style architecture that allows different models for
+designed with a plug-in style architecture that allows different models for
 facility types, known as \emph{archetypes}, to be swapped in and out easily.
@@
 flow problem that is optimized to resolve which transactions actually occur.
-The plugin architecture combined with the DRE creates a powerful and flexible
+The plug-in architecture combined with the DRE creates a powerful and flexible
 environment for the natural comparison of different modeling paradigms.
@@
 
-Cyclus was designed to operate natively in unix/linux type environments with
+Cyclus was designed to operate natively in Unix/Linux type environments with
 flexible and scriptable usage and input file format.  Cyclus stores its
@@
 single-facility granularity modeling that is part of the standard Cyclus archetypes, 
-a fleet archetype was created for this work with aggregate facilitiy behavior and
+a fleet archetype was created for this work with aggregate facility behavior and
 continuous flow.
@@
 The fleet reactor, although a single entity modeling many reactors, is
-actually designed to look like many individual reactors to the cyclus
-simulator kernel. This "trick" allows other plugins for managing facility
+actually designed to look like many individual reactors to the Cyclus
+simulator kernel. This "trick" allows other plug-ins for managing facility
 deployments to transparently work with the fleet reactor without having to
@@
 
-    \item Generated power at any timestep, $P_G$, is based on the number of
+    \item Generated power at any time step, $P_G$, is based on the number of
       operating reactors:
@@
 whole-reactor quantized shutdown.  The full implementation of the fleet-based
-reactor is publicy available for download and use \cite{Carlsen2015}.
+reactor is publicly available for download and use \cite{Carlsen2015}.
 
@@
 
-The chosen scenario approximates a transition from a the current once-through light water
+The chosen scenario approximates a transition from the current once-through light water
 reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle within evaluation group 23 (EG-23)
@@
 cycles. The same deployment schedule is used for all four cases; power
-capacity is built along a curve starting at 100GWe in year zero following the
+capacity is built along a curve starting at 100 GWe in year zero following the
 1\% exponential growth curve.  This particular deployment schedule is chosen
@@
     \caption[The cycle staggering effect]{
-        In this scenario, reactors are deployed annualy with a refueling cycle
+        In this scenario, reactors are deployed annually with a refueling cycle
         length (including outage) of 18 months.  So all reactors deployed
@@
 using individual reactor facilities takes about 20 seconds.  This order of
-magnitude difference is an important tradeoff to keep in mind when doing fuel
+magnitude difference is an important trade-off to keep in mind when doing fuel
 cycle analysis in general, particularly when considering optimization or
@@
 shortage to receive fuel and come online together.  The net effect is that the
-fuel shortage degrades reactor cycle stagering overall.  As the simulation
+fuel shortage degrades reactor cycle staggering overall.  As the simulation
 continues, however, new reactors continue to be deployed and reactors with
-synchornized cycles retire resulting in a gradual return to better
+synchronized cycles retire resulting in a gradual return to better
 staggering visible by the end of the 200 years.
@@
         modeling, refueling cycles become much more synchronized during and
-        after the fuel shortage resuting in large power fluctuations - an artifact of the
+        after the fuel shortage resulting in large power fluctuations - an artifact of the
         \emph{cycle staggering effect}.
@@
         the shortage increases cycle synchronization significantly. The
-        \emph{quantized shutdown effect} can also be seen in the discrepency
+        \emph{quantized shutdown effect} can also be seen in the discrepancy
         between quarterly and monthly time steps for fleet reactor modeling
@@
 
+%% Paul Wilson says:
+%%
 %% What about a different measure of fuel sharing effect?  Perfect sharing
@@
 %% this be substantially different?
+%%
+%% rwcarlsen responds:
+%%
+%% This could be interesting, but it is quantifying the significance of the
+%% quantized shutdown effect discussed in the power production section of
+%% results and doesn't belong here with the fuel sharing discussion - which is
+%% only pseudo-related. Fuel sharing effects are when alternative sharing of
+%% fuel between individual reactors could have resulted in different
+%% online/offline capacity levels. These effects step on each other a tiny bit,
+%% but I think it still makes much more sense to discuss/treat them separately.
 
@@
 and fleet cases in Figure \ref{fig:unfueled} suggesting that most of the
-discrepency in cumulative outage between fleet and individual modeling is
+discrepancy in cumulative outage between fleet and individual modeling is
 caused by inefficient fuel sharing in the individual reactor cases.
@@
 offline, but those extra reactors are also discharging less aggregate spent
-fuel for recycling.  This increases the discrepency between the two
+fuel for recycling.  This increases the discrepancy between the two
 cases. Eventually the shortage in case MF ends slightly before it does in case
@@
 associated understanding of how outcomes are affected by added fidelity.
-Cyclus' flexibility for accomodating different modeling choices uniquely
+Cyclus' flexibility for accommodating different modeling choices uniquely
 enables many interesting comparisons.  Continued work in this direction has

### 2015-10-21 paper.tex
change S to tau and delta t_out in equations

@@
 \begin{equation}
-    N_{wasted}(t) = \sum\limits_{r \in R_t} \left(H[t-S_{sched}(t,r)] \cdot [1-O(t,r)]  \sum\limits_{t'=F_{prev}(t,r)}^{t} N_b(t',r) \right)
+    N_{wasted}(t) = \sum\limits_{r \in R_t} \left(H[t-(\tau(t,r) + \Delta\;t_{out})] \cdot [1-O(t,r)]  \sum\limits_{t'=F_{prev}(t,r)}^{t} N_b(t',r) \right)
     \label{eq:badshare}
@@
 where $R_t$ is the set of all reactors, $H$ is the Heaviside function,
-$S_{sched}(t,r)$ is the original scheduled startup time of reactor $r$ after
-the most recent refueling outage that started before $t$, $F_{prev}(t,r)$ is
+$\tau(t,r)$ is the beginning of the most recent refueling outage of reactor $r$ before $t$,
+$\Delta\; t_{out}$ is the normal refueling outage duration, $F_{prev}(t,r)$ is
 the start of the most recent refueling outage for reactor $r$ before or on
@@
 \begin{equation}
-    P_{outage}(t) = \sum\limits_{r \in R_t} C_r \cdot H[t-S_{sched}(t,r)] \cdot [1-O(t,r)]
+    P_{outage}(t) = \sum\limits_{r \in R_t} C_r \cdot H[t-(\tau(t,r) + \Delta\;t_{out})] \cdot [1-O(t,r)]
     \label{eq:shortage}
@@
 where $R_t$ is the set of all reactors, $C_r$ is the power capacity of reactor
-$r$, $H$ is the Heaviside function, $S_{sched}(t,r)$ is the original scheduled
-startup time of reactor $r$ after the most recent refueling outage that
-started before $t$, and $O(t,r)$ is 1 if reactor $r$ is producing power at
-time $t$ and zero otherwise.
+$r$, $H$ is the Heaviside function, $\tau(t,r)$ is the beginning of the most
+recent refueling outage of reactor $r$ before $t$, $\Delta\; t_{out}$ is the
+normal refueling outage duration, and $O(t,r)$ is 1 if reactor $r$ is
+producing power at time $t$ and zero otherwise.
 

### 2015-10-21 paper.tex
a few minor tweaks

@@
 relationships between the facilities.  Compositions used for the simulations
-were simple containing primarily $^{238}U$, $^{235}U$, $^{239}Pu$, $^{241}Am$,
+were simple containing $^{238}U$, $^{235}U$, $^{239}Pu$, $^{241}Am$,
 and small quantities of a few lanthanides.  Full composition details can be
@@
 \begin{equation}
-    P_{outage}(t) = \sum\limits_{r \in R_t} C_r \cdot H[t-(\tau(t,r) + \Delta\;t_{out})] \cdot [1-O(t,r)]
+    P_{outage}(t) = \sum\limits_{r \in R_t} P_r \cdot H[t-(\tau(t,r) + \Delta\;t_{out})] \cdot [1-O(t,r)]
     \label{eq:shortage}
@@
 
-where $R_t$ is the set of all reactors, $C_r$ is the power capacity of reactor
+where $R_t$ is the set of all reactors, $P_r$ is the power capacity of reactor
 $r$, $H$ is the Heaviside function, $\tau(t,r)$ is the beginning of the most

### 2015-10-21 paper.tex
fix table I to just use case MI values - to avoid complicated discussion of why selected params are different between table I and III.

@@
         Cycle length (months) & 18 & 15 \\
-        Batch size (kg)       & 29565 & 7490 \\
+        Batch size (kg)       & 29565 & 8025 \\
         Batches per core      & 3 & 5 \\
@@
 The described scenario is used as closely as possible in each of the four
-cases.  However, a few small case-dependent changes to the scenarios relating to
-reactor configuration are necessary.
-The invariants preserved with respect to reactor behavior between all
-four cases are shown in Table \ref{tab:invar}.  Table \ref{tab:reactor-detail} shows
-the computed/selected configuration for both LWR and SFR reactor types for all
-four cases.
+cases with appropriate adjustments for case-specific differences (e.g. monthly
+vs. quarterly time steps).  Invariants preserved with respect to reactor
+behavior between all four cases are shown in Table \ref{tab:invar}.  Table
+\ref{tab:reactor-detail} shows the computed/selected configuration for both
+LWR and SFR reactor types that was used in Cyclus input files for all four
+cases.
 
@@
 Cyclus' flexibility for accommodating different modeling choices uniquely
-enables many interesting comparisons.  Continued work in this direction has
-great potential to to improve the integrity of fuel cycle analysis generally.
+enables many interesting comparisons.  Those in the field of fuel cycle
+analysis in general should be cognizant of how modeling choices are affecting
+results and conclusions through exercises like the one here.  Even if the
+outcome is that a particular modeling decision has little impact on results,
+work in this direction has great potential to improve the integrity of fuel
+cycle analysis generally.
 

### 2015-10-21 paper.tex
commas in vals in tables

@@
         Cycle length (months) & 18 & 15 \\
-        Batch size (kg)       & 29565 & 8025 \\
+        Batch size (kg)       & 29,565 & 8,025 \\
         Batches per core      & 3 & 5 \\
@@
                                                                              &          &          \\
-        Discharge Rate ($\frac{\text{kg} \cdot \text{HM}}{\text{month}}$)           & 1642.5   & 535      \\
+        Discharge Rate ($\frac{\text{kg} \cdot \text{HM}}{\text{month}}$)           & 1,642.5   & 535      \\
                                                                              &          &          \\
@@
                                                                              &          &          \\
-        Core Size  ($\text{kg} \cdot \text{HM}$)                                    & 88695    & 40125    \\
+        Core Size  ($\text{kg} \cdot \text{HM}$)                                    & 88,695    & 40,125    \\
                                                                              &          &          \\

### 2015-10-22 paper.tex
no small caps cyclus
elliot's edits

@@
 % No more than three keywords, though each can be a phrase
-\keywords{Nuclear Fuel Cycle, Simulation, \Cyclus}
+\keywords{Nuclear Fuel Cycle, Simulation, Cyclus}
 
@@
 Economic Co-operation and Development (OECD) completed a benchmark study that
-is perhaps one of the most complete published comparisons performed.  Despite
-this, however, various results from the simulators were often significantly
-different due to inconsistency in modeling decisions involving reprocessing
-strategies, refueling behavior, reactor end-of-life handling, etc.  This work
-attempts to identify and quantify the effects of selected modeling choices
-that may sometimes be taken for granted in the fuel cycle modeling domain.
-Four scenarios are compared using combinations of either fleet-based or
-individual reactors with either monthly or quarterly (3-month) time steps.
-The Cyclus fuel cycle simulator's plug-in facility capability along with its
-market-like dynamic material routing allow it to be used as a level playing
-field for comparing the scenarios.  When under supply-constraint pressure, the
-four cases exhibit noticeably different behavior.  Fleet-based modeling is more
-efficient in supply-constrained environments at the expense of losing insight
-on issues such as realistically suboptimal fuel distribution and challenges in
-reactor refueling cycle staggering.  Finer-grained time steps also enable more
+is perhaps the most complete published comparisons performed.  Despite this,
+various results from the simulators were often significantly different due to
+inconsistency in modeling decisions involving reprocessing strategies,
+refueling behavior, reactor end-of-life handling, etc.  This work identifies
+and quantifies the effects of selected modeling choices that may sometimes be
+taken for granted in the fuel cycle modeling domain.  Four scenarios are
+compared using combinations of either fleet-based or individual reactors with
+either monthly or quarterly (3-month) time steps.  The scenarios approximate a
+transition from the current once-through light water reactor (LWR) fleet to a
+full sodium fast reactor (SFR) fuel cycle The Cyclus fuel cycle simulator's
+plug-in facility capability along with its market-like dynamic material
+routing allow it to be used as a level playing field for comparing the
+scenarios.  When under supply-constraint pressure, the four cases exhibit
+noticeably different behavior.  Fleet-based modeling is more efficient in
+supply-constrained environments at the expense of losing insight on issues
+such as realistically suboptimal fuel distribution and challenges in reactor
+refueling cycle staggering.  Finer-grained time steps also enable more
 efficient material use in supply-constrained environments.  Large simulations
 with fleet-based reactors run much more quickly than their individual reactor
-counterparts.
+counterparts.  Gaining a better understanding of how these and other modeling
+choices affect fuel cycle dynamics will enable making more deliberate
+decisions with respect to trade-offs such as computational investment vs.
+realism.
 }
@@
 discretization (i.e. individual facilities vs. aggregate fleets) and time step
-duration.  A better understanding of how these affect fuel cycle dynamics can
+duration.  A better understanding of how these affect fuel cycle dynamics will
 allow for making more deliberate decisions with respect to trade-offs such as
@@
 \cite{forrester_industrial_1961} to modeling. These simulators model several
-"stocks" representing isotopes or isotope groups that have levels that change
-over time due to "flows" between them.  The flows are determined by equations
+\emph{stocks} representing isotopes or isotope groups that have quantities that change
+over time due to \emph{flows} between them.  The flows are determined by equations
 that are a function of the state of the system (i.e. the system's stocks).  A
@@
 changes within the entire system. The effects of the time step duration are
-also investigated together with fleet vs facility modeling.
+also investigated together with fleet vs. facility modeling.
 
@@
 
-\subsection{\Cyclus}
+\subsection{Cyclus}
 
@@
 Cyclus' module plug-in based architecture in concert with its DRE has the ability to
-model fleet-based facilities.  Additionally, however, Cyclus can model
+model fleet-based facilities.  Additionally, Cyclus can model
 facilities individually with discrete flows enabling the investigation of
-interesting real-world effects such as competition, reactor cycle staggering,
+real-world effects such as competition, reactor cycle staggering,
 and individual reactor outages.
@@
         available up to the total amount to fill all reactor cores in the
-        fleet.  No extra fuel is kept on-hand - it is all ordered
+        fleet.  No extra fuel is kept on-hand --- it is all ordered
         just-in-time.
@@
     \item When a reactor in the fleet is retired, a full core of material is
-        discharged - even if there is some fraction of the fleet with unfueled
+        discharged --- even if there is some fraction of the fleet with unfueled
         cores.
@@
 deployment and no more thermal reactors are built.  LWR spent fuel separations
-begins in year 15 with 2000 MTHM/yr capacity and increases to 3000 MTHM/yr in
+begins in year 15 with 2,000 MTHM/yr capacity and increases to 3,000 MTHM/yr in
 year 25.  Fast reactor spent fuel separations and fuel fabrication have
@@
     \centering
-    \includegraphics[width=0.47\textwidth]{exp2/flow.eps}
+    \includegraphics[keepaspectratio=true,width=0.45\textwidth,height=.9\textheight]{exp2/flow.eps}
     \caption[Material flow paths]{Potential material flow paths between facility types.}
@@
 %% this a little harder to reproduce.
-A pre-release of Cyclus version 1.4 was used for this analysis, with the
-standard archetypes provided with Cyclus in the included Cycamore
-library - also a pre-release of version 1.4.  A custom storage archetype was
-developed because Cycamore did not yet provide one.  The standard Cycamore reactor archetype
+A pre-release of Cyclus version 1.4 was used for this analysis.  A standard library of facility plug-ins, Cycamore, is provided with Cyclus.
+Like Cyclus, a pre-release of Cycamore version 1.4 was used.  A custom storage facility archetype was
+developed because Cycamore did not yet provide one.  Code for the fleet reactor and
+storage archetypes is publicly available and can be downloaded \cite{Carlsen2015}. The standard Cycamore reactor archetype
 was used for both LWR and SFR reactor types in cases MI and QI.  Correspondingly
@@
                                           & \multicolumn{2}{c |}{LWR}       & \multicolumn{2}{c |}{SFR} \\
-                                          & Cases MI,QI & Cases MF,QF (fleet) & Cases MI,QI & Cases MF,QF (fleet)  \\
+                                          & Cases MI, QI & Cases MF, QF & Cases MI, QI & Cases MF, QF  \\
         \hline                       
@@
 
-In each of the four cases, the initial 100 LWRs are modified to have a zero
-length refueling outage with an explicit 900 MWe net power output capacity.  Because the
-individual reactor model doesn't have any way to start a reactor mid-cycle,
+In each of the four cases, the initial 100 LWRs are modified to have a zero-length
+refueling outage with an explicit 900 MWe net power output capacity.  Because the
+individual reactor model does not have any way to start a reactor mid-cycle,
 this was done to avoid having all initial LWRs refuel at the same time.  The
@@
 
-There are various low level phenomena that have potential to affect the larger
+There are various low-level phenomena that have potential to affect the larger
 outcome of simulations with respect to facility discretization and time step
@@
 of power for that time step where $P_r$ is the power capacity of a single
-reactor of the fleet (\emph{e.g.}~1000~MWe).  If an individually modeled reactor is
+reactor of the fleet (e.g.~1000~MWe).  If an individually modeled reactor is
 short one half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
@@
 Consider an example of a 3-batch reactor with a fuel shortage of $\frac{1}{2}$ batch of
-fuel that would be resolve in 1 month.  Depending on the modeling choice, each reactor
+fuel that would be resolved in 1 month.  Depending on the modeling choice, each reactor
 would produce the following amounts of energy over a 3 month period:
@@
   \item case \emph{QI}: $0 \cdot 3 \text{ mo} \cdot P_r = 0$
-  \item case \emph{MI}: $[0\cdot 1 \text{ mo} + 1 \cdot 2\text{mo}] \cdot P_r = 2 P_r$
+  \item case \emph{MI}: $(0\cdot 1 \text{ mo} + 1 \cdot 2\text{ mo}) \cdot P_r = 2 P_r$
 \end{itemize}
@@
         In this scenario, reactors are deployed annually with a refueling cycle
-        length (including outage) of 18 months.  So all reactors deployed
+        length (including outage) of 18 months.  As a result, all reactors deployed
         every 3rd year refuel together.  All initial 100 LWRs start out with
@@
 distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  A system
-being short 3 batches might mean that one reactor has no batches and isn't
+being short 3 batches might mean that one reactor has no batches and is not
 operating or that 3 reactors are each missing a single batch.  This is
 referred to as the \emph{fuel sharing effect}.  Fleet reactors, by design,
-have perfect fuel sharing - all fuel is distributed so that the minimum
+have perfect fuel sharing --- all fuel is distributed so that the minimum
 possible amount of capacity is offline during shortages. The most correct
@@
 
+\noindent
 where $R_t$ is the set of all reactors, $H$ is the Heaviside function,
@@
 single 200 year simulation with several hundred fleet-based reactors takes
-about 2 seconds to run on a typical desktop machine where the same simulation
+about 2 seconds to run on a computer with an Intel Core(TM) i7-4770 3.40GHz CPU where the same simulation
 using individual reactor facilities takes about 20 seconds.  This order of
@@
         modeling, refueling cycles become much more synchronized during and
-        after the fuel shortage resulting in large power fluctuations - an artifact of the
+        after the fuel shortage resulting in large power fluctuations --- an artifact of the
         \emph{cycle staggering effect}.
@@
         reactor outage than necessary. The consistent jumping up and down for
-        the fleet reactor cases (even pre-shortage) is due to the mis-alignment of reactors going
+        the fleet reactor cases (even pre-shortage) is due to the misalignment of reactors going
         offline (multiples of 12 months) and the build period (21 months).
@@
 
+\noindent
 where $R_t$ is the set of all reactors, $P_r$ is the power capacity of reactor
@@
 given to a reactor that ends up not being able to start its cycle at the
-scheduled time (due to fuel shortage) adds one to this cumulative for each
+scheduled time (due to fuel shortage) adds one to this cumulative total for each
 time step the cycle start is delayed.
@@
 
-One useful over-estimate of the fuel sharing inefficiency is to assume every
+One useful overestimate of the fuel sharing inefficiency is to assume every
 one of the "wasted" batches could have been given to a reactor enabling it to
@@
 effect of increasing total shortage severity; on average more batches of fresh
-fuel will be sitting unused.  However, it can potentially reduce the frequency
+fuel will be idling unused.  However, it can potentially reduce the frequency
 of outages for individual reactors in some cases.
@@
 points during the shortage where the outflow point lies exactly on top of the
-inventory point - indicating that all available inventory is being
+inventory point --- indicating that all available inventory is being
 transferred.
@@
 must be maintained at a higher limit as described above.  The small difference
-in outage start time has feed-back effects.  Not only are extra reactors
+in outage start time has feedback effects.  Not only are extra reactors
 offline, but those extra reactors are also discharging less aggregate spent
@@
 when many reactors come online together in groups and are resonances that the
-inventory level jumps between.  These resonances gradually disappear as the
+inventory level jumps between.  This artifact is associated with poor cycle staggering
+like that shown earlier in Figure \ref{fig:sync-cycle}.
+These resonances gradually disappear as the
 simulation progresses farther in time beyond the shortage and refueling cycles
@@
     \caption[Separated Pu inventory and flow: Case MI Zoom]{
-        Separated Pu inventory and flows for case MI.  At locations where the
+        Zoom view of separated Pu inventory and flows from Figure \ref{fig:puinv} for case MI.  At locations where the
         outflow dots lie on/near the inventory level, fuel shortages are
@@
     \caption[Separated Pu inventory and flow: Case MF Zoom]{
-        Separated Pu inventory and flows for case MF.  At locations where the
+        Zoom view of separated Pu inventory and flows from Figure \ref{fig:puinv} for case MF.  At locations where the
         outflow dots meet the inventory level, fuel shortages are taking
@@
 \ref{fig:puinv-compare-zoom}.  Cases MI and QF recover Pu inventory
-post-shortage at about the same time and rate - this is consistent with their
+post-shortage at about the same time and rate --- this is consistent with their
 approximately equivalent cumulative offline power capacity curves in Figure
@@
 to a globally optimum fuel management strategy.  Keeping on-hand fresh fuel
-batches would also have the same effect of increasing idling plutonium
+batches would also have the same effect of increasing idling Pu
 quantities and would make the fuel shortage worse overall.

### 2015-10-23 coverletter/cover.tex
final fixes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}
 \begin{document}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor A.C. Klein\\
Editor, Nuclear Technology}
%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\address{Robert W. Carlsen\\
229 N Midvale Blvd Apt 1\\
Madison, WI 53705}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Professor Klein,}

Please find enclosed a manuscript entitled: ``Challenging Fuel Cycle Modeling
Assumptions: Facility and Time Step Discretization Effects'' which I am
submitting for exclusive consideration of publication as a research article in
Nuclear Technology.  This manuscript describes work investigating impacts of
common fuel cycle simulator modeling choices in order to enable more
deliberate simulation design and help improve fuel cycle analysis integrity
generally.

Thank you for your consideration of this work. Please address all
correspondence concerning this manuscript to me and feel free to correspond
with me by e-mail.

\closing{
Happy Halloween,

Robert Carlsen
}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}

### 2015-10-23 paper.tex
final fixes

@@
 Economic Co-operation and Development (OECD) completed a benchmark study that
-is perhaps the most complete published comparisons performed.  Despite this,
+is perhaps the most complete published comparison performed.  Despite this,
 various results from the simulators were often significantly different due to
@@
 taken for granted in the fuel cycle modeling domain.  Four scenarios are
-compared using combinations of either fleet-based or individual reactors with
+compared using combinations of either fleet-based or individually modeled reactors with
 either monthly or quarterly (3-month) time steps.  The scenarios approximate a
-transition from the current once-through light water reactor (LWR) fleet to a
-full sodium fast reactor (SFR) fuel cycle The Cyclus fuel cycle simulator's
+transition from the current U.S. once-through light water reactor (LWR) fleet to a
+full sodium fast reactor (SFR) fuel cycle. The Cyclus fuel cycle simulator's
 plug-in facility capability along with its market-like dynamic material
@@
 
-TODO: thank ross maybe for vision-sourced inspiration about modeling effects
-
 This research is being performed in part using funding received from the DOE

### 2016-02-03 diff.sh
add script for generating latex diffs

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+oldref=$1
+olddir=/tmp/latexdiff1
+
+newref=HEAD
+newdir=/tmp/latexdiff2
+
+if [[ $# -ge 2 ]]; then
+    newref=$2
+fi
+
+echo "Running 'latexdiff $oldref $newref'..."
+
+
+rm -rf $olddir $newdir
+mkdir -p $olddir
+mkdir -p $newdir
+bash -c "cd $olddir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &>/dev/null && git checkout $oldref &>/dev/null"
+bash -c "cd $newdir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &>/dev/null && git checkout $newref &>/dev/null"
+latexdiff --flatten $olddir/fleetpaper/paper.tex $newdir/fleetpaper/paper.tex > diff.tex
+pdflatex diff.tex &> /dev/null
+bibtex diff.aux &>/dev/null
+pdflatex diff.tex &> /dev/null
+bibtex diff.aux &>/dev/null
+pdflatex diff.tex &> /dev/null
+
+rm -rf $olddir $newdir diff.tex diff.aux diff.blg diff.glo diff.ist diff.log diff.bbl diff.acn diff.out diff.xwm
+

### 2016-02-04 paper.tex
address reviewer 1 comments

@@
 various results from the simulators were often significantly different due to
-inconsistency in modeling decisions involving reprocessing strategies,
+inconsistencies in modeling decisions involving reprocessing strategies,
 refueling behavior, reactor end-of-life handling, etc.  This work identifies
 and quantifies the effects of selected modeling choices that may sometimes be
-taken for granted in the fuel cycle modeling domain.  Four scenarios are
-compared using combinations of either fleet-based or individually modeled reactors with
-either monthly or quarterly (3-month) time steps.  The scenarios approximate a
-transition from the current U.S. once-through light water reactor (LWR) fleet to a
-full sodium fast reactor (SFR) fuel cycle. The Cyclus fuel cycle simulator's
-plug-in facility capability along with its market-like dynamic material
-routing allow it to be used as a level playing field for comparing the
-scenarios.  When under supply-constraint pressure, the four cases exhibit
-noticeably different behavior.  Fleet-based modeling is more efficient in
-supply-constrained environments at the expense of losing insight on issues
-such as realistically suboptimal fuel distribution and challenges in reactor
-refueling cycle staggering.  Finer-grained time steps also enable more
-efficient material use in supply-constrained environments.  Large simulations
-with fleet-based reactors run much more quickly than their individual reactor
-counterparts.  Gaining a better understanding of how these and other modeling
-choices affect fuel cycle dynamics will enable making more deliberate
-decisions with respect to trade-offs such as computational investment vs.
-realism.
+taken for granted in the fuel cycle simulation domain.  Four scenarios are
+compared using combinations of either fleet-based or individually modeled
+reactors with either monthly or quarterly (3-month) time steps.  The scenarios
+approximate a transition from the current U.S. once-through light water
+reactor (LWR) fleet to a full sodium fast reactor (SFR) fuel cycle. The Cyclus
+fuel cycle simulator's plug-in facility capability along with its market-like
+dynamic material routing allow it to be used as a level playing field for
+comparing the scenarios.  When under supply-constraint pressure, the four
+cases exhibit noticeably different behavior.  Fleet-based modeling is more
+efficient in supply-constrained environments at the expense of losing insight
+on issues such as realistically suboptimal fuel distribution and challenges in
+reactor refueling cycle staggering.  Finer-grained time steps also enable more
+efficient material use in supply-constrained environments resulting in much lower
+standing inventories of separated Pu.  Large simulations with fleet-based
+reactors run much more quickly than their individual reactor counterparts.
+Gaining a better understanding of how these and other modeling choices affect
+fuel cycle dynamics will enable making more deliberate decisions with respect
+to trade-offs such as computational investment vs.  realism.
 }
@@
 
+This work identifies and quantifies the effects of selected modeling choices
+that may sometimes be taken for granted in the fuel cycle simulation domain.
+Four scenarios are compared using combinations of either fleet-based or
+individually modeled reactors with either monthly or quarterly (3-month) time
+steps.  Time step duration and the degree facility discretization are design
+decisions that are handled in a variety of unique ways among nuclear fuel
+cycle simulators, and a better understanding of their effects can serve to
+increase confidence in results and can help make inter-simulator comparisons
+more meaningful.  Reducing error and noise artifacts associated with modeling
+decisions will provide a more stable foundation for understanding the
+significance of external market, political, and social forces that often
+dominate the industry in the real world.
+
 \subsection{Motivation and Background}
+
+Informal comparison of results from the VISION and Cyclus simulators when
+running equivalent scenarios revealed some surprising discrepencies.  One of
+the most apparent was a large difference in separated Pu levels between the
+two simulators --- something that some in the DOE and nuclear community are
+very sensitive to.  For one particular scenario, the Cyclus results also
+showed no fuel shortages for reactors, while the VISION results indicated
+noticable shortages.  This observation prompted a closer investigation to
+determine if the differences were caused by mistakes in simulator input, bugs
+in the simulator(s), or differing modeling choices between the simulators.
+This investigation eventaully developed into this work which shows important
+effects of some modeling choices that are operative in many fuel cycle
+simulators.
 
@@
 sharing (described below) and no noise from refueling outages, differences observed will
-likely be due to quantized shutdown.
+partly be due to quantized shutdown.
 
@@
 shortage lasts until about year 140. The power generated in case QF is slightly
-lower than in case MF during the fuel shortage years. This is primarily due to
+lower than in case MF during the fuel shortage years. This is partly due to
 the \emph{quantized shutdown effect}.  The longer 3-month time step in case QF 
@@
         the shortage increases cycle synchronization significantly. The
-        \emph{quantized shutdown effect} can also be seen in the discrepancy
+        \emph{quantized shutdown effect} causes part of the discrepancy
         between quarterly and monthly time steps for fleet reactor modeling
@@
 Increasing time step duration increases facilities' standing inventory
-requirement during supply constrained periods.  Facilities can provide no more
-inventory than they can hold on a single time step.  So increasing time step
-duration for a facility with unconstrained supply generally means higher
-standing inventories are required in order to achieve equal throughput.
-Facilities with large inventory buffer that are unconstrained are not affected
-significantly by the time step duration when operating in isolation.  However,
-even when facilities do not have individual inventory or throughput limits,
-the time step can still have a significant effect. The \emph{inventory drawdown
-effect} can impact the aggregate throughput of supply-constrained recycle loops by affecting
-minimum bounds on idling material quantities.
+requirement during supply constrained periods.  Here we observe a factor of
+three difference between standing inventories of separated Pu for a large
+portion of the simulation --- something certain analyses are very sensitive
+to.  Facilities can provide no more inventory than they can hold on a single
+time step.  So increasing time step duration for a facility with unconstrained
+supply generally means higher standing inventories are required in order to
+achieve equal throughput.  Facilities with large inventory buffer that are
+unconstrained are not affected significantly by the time step duration when
+operating in isolation.  However, even when facilities do not have individual
+inventory or throughput limits, the time step can still have a significant
+effect. The \emph{inventory drawdown effect} can impact the aggregate
+throughput of supply-constrained recycle loops by affecting minimum bounds on
+idling material quantities.
 
@@
 For example, someone interested in fuel-shortage driven reactor outages would
-observe a factor of four span among the four cases shown in the detail
-comparison.  However, someone more interested in the overall transition
+observe a factor of four span among the four cases shown in this comparison
+exercise.  However, someone more interested in the overall transition
 mechanics would not see as significant a discrepancy.  In an unconstrained
 supply setting, varying the time step duration or discretization of facility
-modeling have a relatively minor impact on the overall simulation results.
+modeling has a smaller impact on the overall simulation results.
 Differences become more significant in material constrained regimes.
-Robustness in constrained regimes is especially useful in the context of
+Robustness in constrained regimes can be especially useful in the context of
 sensitivity and optimization analysis.
 
-The value of realism in fuel cycle modeling is greatly be enhanced by an
+The value of realism in fuel cycle modeling is greatly enhanced by an
 associated understanding of how outcomes are affected by added fidelity.
@@
 outcome is that a particular modeling decision has little impact on results,
-work in this direction has great potential to improve the integrity of fuel
-cycle analysis generally.
+efforts to measure and document these impacts can provide confidence in the
+appropriateness of modeling choices, reducing the need to rely on intuition
+for cycle analysis work.
 

### 2016-02-04 diff.sh
update diff.sh to include abstract

@@ -18,7 +18,7 @@ mkdir -p $olddir
 mkdir -p $newdir
 bash -c "cd $olddir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &>/dev/null && git checkout $oldref &>/dev/null"
 bash -c "cd $newdir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &>/dev/null && git checkout $newref &>/dev/null"
-latexdiff --flatten $olddir/fleetpaper/paper.tex $newdir/fleetpaper/paper.tex > diff.tex
+latexdiff --flatten --append-context2cmd="myabstract" $olddir/fleetpaper/paper.tex $newdir/fleetpaper/paper.tex > diff.tex
 pdflatex diff.tex &> /dev/null
 bibtex diff.aux &>/dev/null
 pdflatex diff.tex &> /dev/null

### 2016-02-05 paper.tex
spell check and move+tweak tar archive mention

@@
 Informal comparison of results from the VISION and Cyclus simulators when
-running equivalent scenarios revealed some surprising discrepencies.  One of
+running equivalent scenarios revealed some surprising discrepancies.  One of
 the most apparent was a large difference in separated Pu levels between the
@@
 showed no fuel shortages for reactors, while the VISION results indicated
-noticable shortages.  This observation prompted a closer investigation to
+noticeable shortages.  This observation prompted a closer investigation to
 determine if the differences were caused by mistakes in simulator input, bugs
 in the simulator(s), or differing modeling choices between the simulators.
-This investigation eventaully developed into this work which shows important
+This investigation eventually developed into this work which shows important
 effects of some modeling choices that are operative in many fuel cycle
@@
 \section{Results}
-
-All results data, analysis tools, and related instructions are publicly
-available for download and use at
-\url{http://dx.doi.org/10.6084/m9.figshare.1546775} \cite{Carlsen2015}.
 
@@
 
+Raw data, custom code, instructions for reproducing results, and other
+artifacts are available for download and use at
+\url{http://dx.doi.org/10.6084/m9.figshare.1546775} \cite{Carlsen2015}.
+
 \section{Conclusion}

### 2016-02-05 paper.tex
address more comments

@@
 facility types, known as \emph{archetypes}, to be swapped in and out easily.
+While the reactor archetypes used in this analysis do not implement
+internal physics calculations, plug-in archetypes such as Bright-lite
+\cite{flanagan_bright-lite_2014} have been developed for Cyclus that provide
+varying degrees modeling fidelity.
+% response to technical comment 2.
 
@@
 
-A single fuel cycle transition scenario is used with a few modeling variations
-resulting in four different cases:
+Two particular design choices are selected for investigation: time step
+duration and facility discretization.  A single fuel cycle transition scenario
+is run in a set of simulations using combinations of the time step duration
+and facility discretization  --- four four different cases:
 
@@
 
+%% response to technical comment 1.
+These four cases span a spectrum of fidelity with one end being more realistic
+(i.e. smaller time steps, individual facilities) and the other being more
+appropriate for scoping studies (i.e.  larger time steps, fleet facilities).
 Conducting this analysis involved developing some additional capability
@@
 makes it relatively straightforward to create plug-in archetypes to model
-anywhere along the spectrum from individual, discrete facilities to
-continuous flow fleet facilities.  In order to contrast well with the individual,
-single-facility granularity modeling that is part of the standard Cyclus archetypes, 
-a fleet archetype was created for this work with aggregate facility behavior and
-continuous flow.
+anywhere along the spectrum from individual, discrete facilities to continuous
+flow fleet facilities.  In order to contrast well with the individual,
+single-facility granularity modeling that is part of the standard Cyclus
+archetypes, a fleet archetype was created for this work with aggregate
+facility behavior and continuous flow.  Although it would certainly be
+interesting to compare many intermediate levels of facility discretization
+(e.g. 1-reactor groups, 10-reactor groups, single-fleet, etc.), such a
+detailed analysis is beyond the scope of this work.
+% address technical comment 14
 
@@
     \centering
-    \includegraphics[keepaspectratio=true,width=0.45\textwidth,height=.9\textheight]{exp2/flow.eps}
-    \caption[Material flow paths]{Potential material flow paths between facility types.}
-    %% consider a Cyclist-derived visualization here?
+    \includegraphics[keepaspectratio=true,width=0.45\textwidth,height=.89\textheight]{exp2/flow.eps}
+    \caption[Material flow paths]{
+        Material flow paths between facility types. The DU source is separate
+        from enrichment tails for modeling simplicity, although enrichment
+        tails could be used for added realism.  The primary fissile isotope in
+        the fissile separations stream is $_{239}Pu$.
+    }
+    % response to technical comment 9
     \label{fig:flow}
@@
 
-In each of the four cases, the initial 100 LWRs are modified to have a zero-length
-refueling outage with an explicit 900 MWe net power output capacity.  Because the
-individual reactor model does not have any way to start a reactor mid-cycle,
-this was done to avoid having all initial LWRs refuel at the same time.  The
-parameters for each case were also selected to keep capacity factors for
+In each of the four cases, the initial 100 LWRs are modified to have a
+zero-length refueling outage with an explicit 900 MWe net power output
+capacity.  Because the individual reactor model does not have any way to start
+a reactor mid-cycle, this is a modeling trick that was done to avoid having
+all initial LWRs refuel at the same time.  This adjustment serves to improve
+the realism associated with the effects being investigated (i.e. fleet-average
+power generation) even though it is unrealistic in other ways.
+% response to technical comment 11
+
+The parameters for each case were also selected to keep capacity factors for
 individual reactors very roughly around 90\%, although they are actually
@@
 
+% response to technical comment 12
 Instead of annual deployments, a longer 21-month deployment period is used
 because it provides better natural staggering for the reactor refueling
-cycles. The same deployment schedule is used for all four cases; power
-capacity is built along a curve starting at 100 GWe in year zero following the
-1\% exponential growth curve.  This particular deployment schedule is chosen
-because it causes significant fuel shortages providing interesting features to
-investigate and compare.
+cycles. The 630-month lowest common multiple (LCM) for 21-months (deployment
+period), 15 months (SFR cycle), and 18 months (LWR cycle) is much larger than
+the 180-month LCM when using a 12-month build period providing a more even
+power generation profile.  The same deployment schedule is used for all four
+cases; power capacity is built along a curve starting at 100 GWe in year zero
+following the 1\% exponential growth curve.  Although many alternate
+deployment schedules exist that avoid SFR fuel shortages (e.g. a more gradual
+shift toward SFRs), this particular schedule is deliberately chosen because it
+causes significant fuel shortages.  This is done in order to force the
+simulation and facility models into more extreme circumstances in order to
+expose behavioral differences among the design choices investigated.
+% response to general comment paragraph
+% response to technical comment 13
 

### 2016-02-05 paper.tex
remove improper usage of 'due to'

@@
 is perhaps the most complete published comparison performed.  Despite this,
-various results from the simulators were often significantly different due to
+various results from the simulators were often significantly different because of
 inconsistencies in modeling decisions involving reprocessing strategies,
@@
 comparisons performed.  Despite this, however, various results from the
-different simulators were often significantly different due to inconsistent
+different simulators were often significantly different because of inconsistent
 modeling decisions involving reprocessing strategies, refueling behavior,
@@
 \emph{stocks} representing isotopes or isotope groups that have quantities that change
-over time due to \emph{flows} between them.  The flows are determined by equations
+over time because of \emph{flows} between them.  The flows are determined by equations
 that are a function of the state of the system (i.e. the system's stocks).  A
@@
 individual reactors very roughly around 90\%, although they are actually
-closer to 80\% due to integer variable constraints on the case MI and QI
+closer to 80\% because of integer variable constraints on the case MI and QI
 reactor configuration.
@@
 sharing (described below) and no noise from refueling outages, differences observed will
-partly be due to quantized shutdown.
+partly be a result of quantized shutdown.
 
@@
 shortage lasts until about year 140. The power generated in case QF is slightly
-lower than in case MF during the fuel shortage years. This is partly due to
+lower than in case MF during the fuel shortage years. This is partly a result of
 the \emph{quantized shutdown effect}.  The longer 3-month time step in case QF 
@@
         reactor outage than necessary. The consistent jumping up and down for
-        the fleet reactor cases (even pre-shortage) is due to the misalignment of reactors going
+        the fleet reactor cases (even pre-shortage) is caused by the misalignment of reactors going
         offline (multiples of 12 months) and the build period (21 months).
@@
     \includegraphics[width=1.0\columnwidth]{exp2/unfueled.eps}
-    \caption[Cumulative offline power due to fuel shortage]{
-        Cumulative offline power for reactors that had delayed cycle start due
-        to fuel shortage. The difference between individual and fleet based
+    \caption[Cumulative offline power caused by fuel shortage]{
+        Cumulative offline power for reactors that had delayed cycle start caused
+        by fuel shortage. The difference between individual and fleet based
         modeling is primarily a result of the \emph{fuel sharing effect}. The
         difference between monthly and quarterly time steps is primarily
-        caused by the \emph{inventory drawdown effect}.
+        a result of the \emph{inventory drawdown effect}.
     }
@@
 given to a reactor that ends up not being able to start its cycle at the
-scheduled time (due to fuel shortage) adds one to this cumulative total for each
+scheduled time (caused by fuel shortage) adds one to this cumulative total for each
 time step the cycle start is delayed.
@@
         by the number of time steps that reactor had to delay the start of its
-        next cycle (due to fuel shortage).  Fleet reactors implicitly have perfect fuel
+        next cycle (caused by fuel shortage).  Fleet reactors implicitly have perfect fuel
         sharing and so are not included here. The large difference
-        between cases MI and QI is mostly due to the \emph{inventory drawdown
+        between cases MI and QI is mostly a result of the \emph{inventory drawdown
         effect}. 
@@
 The large differences seen between monthly and quarterly time step cases in
-Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily caused by the
+Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily a result of the
 \emph{inventory drawdown effect}.  Figure \ref{fig:puinv} helps to visualize
@@
 Throughout most of the shortage, the aggregate outflow of Pu from separated
-inventory is roughly the same in cases MF and QF.  However, due to the longer
+inventory is roughly the same in cases MF and QF.  However, because of the longer
 time steps in case QF, the shortage starts slightly sooner because Pu inventory
@@
 QF take longer to recover from the shortage and fall a bit behind with respect
-to building up Pu stocks due to the delayed contributions of more unfueled
+to building up Pu stocks because of the delayed contributions of more unfueled
 reactors to the separated Pu pool.  This can be seen most clearly in Figure

### 2016-02-05 paper.tex
i.e./e.g. -> i.e.,/e.g.,

@@
 quantify the effects of selected modeling choices relative to facility modeling
-discretization (i.e. individual facilities vs. aggregate fleets) and time step
+discretization (i.e., individual facilities vs. aggregate fleets) and time step
 duration.  A better understanding of how these affect fuel cycle dynamics will
@@
 over time because of \emph{flows} between them.  The flows are determined by equations
-that are a function of the state of the system (i.e. the system's stocks).  A
+that are a function of the state of the system (i.e., the system's stocks).  A
 constructed system is then solved in discrete or continuous time depending on
@@
 to as "fleets".  As the simulation steps through time, the levels of stocks
-(e.g. reactor fresh fuel inventory, repository waste inventories, etc.) are
+(e.g., reactor fresh fuel inventory, repository waste inventories, etc.) are
 adjusted according to the calculated flow values.  Fuel cycle simulators that model
@@
 These four cases span a spectrum of fidelity with one end being more realistic
-(i.e. smaller time steps, individual facilities) and the other being more
-appropriate for scoping studies (i.e.  larger time steps, fleet facilities).
+(i.e., smaller time steps, individual facilities) and the other being more
+appropriate for scoping studies (i.e.,  larger time steps, fleet facilities).
 Conducting this analysis involved developing some additional capability
@@
 interesting to compare many intermediate levels of facility discretization
-(e.g. 1-reactor groups, 10-reactor groups, single-fleet, etc.), such a
+(e.g., 1-reactor groups, 10-reactor groups, single-fleet, etc.), such a
 detailed analysis is beyond the scope of this work.
@@
     \item Reactor capacity is deployed and retired in increments of single
-      reactor units, each capable of producing power $P_r$ (e.g. 1 GWe
+      reactor units, each capable of producing power $P_r$ (e.g., 1 GWe
       LWRs).  $N_{r}$ reactors are deployed at any time capable of producing
@@
 The described scenario is used as closely as possible in each of the four
-cases with appropriate adjustments for case-specific differences (e.g. monthly
+cases with appropriate adjustments for case-specific differences (e.g., monthly
 vs. quarterly time steps).  Invariants preserved with respect to reactor
@@
 all initial LWRs refuel at the same time.  This adjustment serves to improve
-the realism associated with the effects being investigated (i.e. fleet-average
+the realism associated with the effects being investigated (i.e., fleet-average
 power generation) even though it is unrealistic in other ways.
@@
 following the 1\% exponential growth curve.  Although many alternate
-deployment schedules exist that avoid SFR fuel shortages (e.g. a more gradual
+deployment schedules exist that avoid SFR fuel shortages (e.g., a more gradual
 shift toward SFRs), this particular schedule is deliberately chosen because it
@@
 of power for that time step where $P_r$ is the power capacity of a single
-reactor of the fleet (e.g.~1000~MWe).  If an individually modeled reactor is
+reactor of the fleet (e.g.,~1000~MWe).  If an individually modeled reactor is
 short one half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
@@
 it is to compare fleet reactor scenarios that have different time step
-durations (i.e. cases MF and QF).  Because the fleet reactors have perfect fuel
+durations (i.e., cases MF and QF).  Because the fleet reactors have perfect fuel
 sharing (described below) and no noise from refueling outages, differences observed will
@@
 fuel batches distributed to reactors that ended up not having enough fuel to
-operate (i.e. a full core) that time step.  This approximation assumes that
+operate (i.e., a full core) that time step.  This approximation assumes that
 every "wasted" batch of fuel, could have been given to a reactor that only
@@
 longer minimum bounds on time to traverse paths between facilities. Although
-direct measures of this effect are difficult, some of its consequences (e.g.
+direct measures of this effect are difficult, some of its consequences (e.g.,
 higher standing inventories during shortages) can be quantified by comparing
-scenarios with different time step durations (e.g.  cases MF and QF).
+scenarios with different time step durations (e.g.,  cases MF and QF).
 
@@
 amount of work being done in the simulation is similar; in the monthly time
-step cases most facilities (i.e. all reactors) only perform actions every
+step cases most facilities (i.e., all reactors) only perform actions every
 refueling cycle rather than every time step.
@@
 operate.  Multiplying each of these cumulative batches by the fast reactor power capacity
-(i.e. 450 MWe) provides a way to compare fuel sharing inefficiency with the
+(i.e., 450 MWe) provides a way to compare fuel sharing inefficiency with the
 overall fuel outage.  This results in a cumulative energy deficit caused by
@@
 
-    \item Multiple fuel quanta comprise a single batch (i.e. multiple
+    \item Multiple fuel quanta comprise a single batch (i.e., multiple
         assemblies per batch). This functionality is natively supported in the
@@
 
-Also notable is that different modeling choices (e.g. time step duration,
+Also notable is that different modeling choices (e.g., time step duration,
 facility discretization, etc.) may have varying levels of significance

### 2016-02-05 paper.tex
...-based hyphens

@@
 the limitations of the software and constraints imposed by the model.  In
-system dynamics based simulators, the number of stocks and potential flows between
+system dynamics-based simulators, the number of stocks and potential flows between
 them is generally a static property of the modeled system, making it somewhat
@@
 
-System dynamics based fuel cycle simulators generally use a collection of
+System dynamics-based fuel cycle simulators generally use a collection of
 stocks to represent the state of groups of like-facilities commonly referred

### 2016-02-05 paper.tex
small editorial comment tweaks

@@
 the simulation, and the installed versions of Cyclus dependencies. This and
-other features of Cyclus lend themselves well to automation, research
-provenance and reproducibility, and large-scale computing.
+other features of Cyclus lend themselves well to automation, tracking research
+provenance, reproducibility, and large-scale computing.
 
@@
         every 3rd year refuel together.  All initial 100 LWRs start out with
-        their cycles synced as well.
+        their cycles synchronized as well.
     }

### 2016-02-05 diff.sh
update diff script to include bibliography changes

@@ -16,9 +16,11 @@ echo "Running 'latexdiff $oldref $newref'..."
 rm -rf $olddir $newdir
 mkdir -p $olddir
 mkdir -p $newdir
-bash -c "cd $olddir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &>/dev/null && git checkout $oldref &>/dev/null"
-bash -c "cd $newdir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &>/dev/null && git checkout $newref &>/dev/null"
+bash -c "cd $olddir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &>/dev/null && git checkout $oldref &>/dev/null && make &>/dev/null"
+bash -c "cd $newdir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &>/dev/null && git checkout $newref &>/dev/null && make &>/dev/null"
 latexdiff --flatten --append-context2cmd="myabstract" $olddir/fleetpaper/paper.tex $newdir/fleetpaper/paper.tex > diff.tex
+latexdiff $olddir/fleetpaper/paper.bbl $newdir/fleetpaper/paper.bbl > diff.bbl
+
 pdflatex diff.tex &> /dev/null
 bibtex diff.aux &>/dev/null
 pdflatex diff.tex &> /dev/null

### 2016-02-05 paper.tex
adjust flow fig caption

@@
     \caption[Material flow paths]{
-        Material flow paths between facility types. The DU source is separate
-        from enrichment tails for modeling simplicity, although enrichment
-        tails could be used for added realism.  The primary fissile isotope in
-        the fissile separations stream is $_{239}Pu$.
+        Material flow paths between facility types. An independent DU source separate
+        from enrichment tails is used for simplicity.  The primary fissile isotope in
+        the fissile separations stream is $^{239}Pu$.
     }

### 2016-02-08 paper.tex
explain recipe sources better. update references

@@
 simulation scenario was created that is roughly based on prior and ongoing
-work in the Department of Energy (DOE) fuel cycle options campaign.  A few
+work in the Department of Energy (DOE) fuel cycle options campaign \cite{hoffman-eg23-2016}.  A few
 phenomena related to the different modeling choices in each of the four cases
@@
 available for separation: 5 years for cooling and an additional 2 years to approximate non-instantaneous separations and fabrication times. Figure \ref{fig:flow} shows the material flow
-relationships between the facilities.  Compositions used for the simulations
-were simple containing $^{238}U$, $^{235}U$, $^{239}Pu$, $^{241}Am$,
-and small quantities of a few lanthanides.  Full composition details can be
-found in the Cyclus input files used \cite{Carlsen2015}.
+relationships between the facilities.
+
+In fuel cycle simulation, it is common to use simplified fresh and discharged
+fuel compositions that are consistent with results from externally performed
+loading and burnup calculations. Compositions used in this analysis follow
+this approach and are fixed (i.e. not dynamically computed in-situ).  These
+compositions are the same as those used in similar transition analyses by
+Hoffman, et al. \cite{hoffman-eg23-2016} and Littell, et al.
+\cite{littell-cyclus-2015} containing $^{238}U$, $^{235}U$, $^{239}Pu$,
+$^{241}Am$, and small quantities of a few other isotopes as place-holders
+representing fission products. Exact composition details can be found in the
+Cyclus input files used \cite{Carlsen2015}.
 

### 2016-02-08 paper.tex
update cyclus fundamentals paper reference

@@
 One of the major motivations for the development of the Cyclus fuel cycle
-simulator \cite{cyclus_2015} was creating a simulation environment that is
+simulator \cite{cyclus-fundamentals-2016} was creating a simulation environment that is
 more expressive than more traditional fuel cycle simulators.  Cyclus was

### 2016-02-08 paper.tex
minor tweaks

@@
 determine if the differences were caused by mistakes in simulator input, bugs
-in the simulator(s), or differing modeling choices between the simulators.
+in the simulator(s), or different (yet still reasonable) modeling choices of the simulators.
 This investigation eventually developed into this work which shows important
-effects of some modeling choices that are operative in many fuel cycle
-simulators.
+effects of that are operative in many fuel cycle simulators.
 
@@
 %% response to technical comment 1.
+\noindent
 These four cases span a spectrum of fidelity with one end being more realistic
@@
 
+\noindent
 Because the standard package of archetypes included with Cyclus does not

### 2016-02-08 paper.tex
rephrase paper scope mention

@@
 interesting to compare many intermediate levels of facility discretization
-(e.g., 1-reactor groups, 10-reactor groups, single-fleet, etc.), such a
-detailed analysis is beyond the scope of this work.
+(e.g., 1-reactor groups, 10-reactor groups, single-fleet, etc.), only the two
+extremes of continuous, single fleet and individual reactor modeling are
+investigated in order to maintain an appropriate scope.
 % address technical comment 14

### 2016-02-16 diff.sh
add hack to get line numbers in diff

@@ -21,11 +21,13 @@ bash -c "cd $newdir &>/dev/null && git clone $PWD &>/dev/null && cd fleetpaper &
 latexdiff --flatten --append-context2cmd="myabstract" $olddir/fleetpaper/paper.tex $newdir/fleetpaper/paper.tex > diff.tex
 latexdiff $olddir/fleetpaper/paper.bbl $newdir/fleetpaper/paper.bbl > diff.bbl
 
+sed -i 's/%lineno//' diff.tex
+
 pdflatex diff.tex &> /dev/null
 bibtex diff.aux &>/dev/null
 pdflatex diff.tex &> /dev/null
 bibtex diff.aux &>/dev/null
 pdflatex diff.tex &> /dev/null
 
-rm -rf $olddir $newdir diff.tex diff.aux diff.blg diff.glo diff.ist diff.log diff.bbl diff.acn diff.out diff.xwm
+rm -rf $olddir $newdir diff.aux diff.blg diff.glo diff.ist diff.log diff.bbl diff.acn diff.out diff.xwm
 

### 2016-04-21 submit-zip.sh
add script for generating final submission files with special naming reqs, etc.

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+dir=submit-zip
+prefix="NT15-138R1"
+manuscript=${prefix}_LaTeX_FinalText.tex
+
+rm -rf $dir
+mkdir -p $dir
+
+cp submission-makefile $dir/Makefile
+cp paper.tex $dir/$manuscript
+
+cp ans.bst $dir/${prefix}_bibstyle.bst
+sed -i 's/{ans}/{'"$prefix"'_bibstyle}/g' $dir/$manuscript
+
+cp refs.bib $dir/${prefix}_references.bib
+sed -i 's/{refs}/{'"$prefix"'_references}/g' $dir/$manuscript
+
+cp printmanuscript.cls $dir/${prefix}_LaTeX_class.cls
+sed -i 's/{style}/{'"$prefix"'_LaTeX_class}/g' $dir/$manuscript
+
+sed -i 's/{neup_logo_large.jpg}/{'"$prefix"'_neup_logo_large.jpg}/g' $dir/$manuscript
+cp neup_logo_large.jpg $dir/${prefix}_neup_logo_large.jpg
+
+srcs=("flow.eps" "sync-cycle.eps" "fuel-sharing.pdf" "power.eps" "power-rel.eps" "unfueled.eps" "badshare.eps" "puinv.eps" "puinv-MI.eps" "puinv-MF.eps" "puinv-compare.eps" "puinv-compare-zoom.eps")
+
+declare -i i=1
+for src in "${srcs[@]}"; do
+    dstname=$(echo "$src" | sed 's/.*\.\(.*\)/'"$prefix"'_Fig'"$i"'.\1/')
+    cp exp2/$src $dir/$dstname
+    sed -i 's/exp2\/'"$src"'/'"$dstname"'/g' $dir/$manuscript
+    i=$i+1
+done
+
+tar -czf fleetpaper-final.tar.gz submit-zip/*
+

### 2016-04-22 paper.tex
add hack to get line numbers in diff
remove watermark and add submittephone and institute address

@@
 
-\usepackage[printwatermark]{xwatermark}
-\usepackage{xcolor}
-\newwatermark[allpages,color=black!5,angle=45,scale=6,xpos=-40,ypos=40]{DRAFT}
-
 %%%%%%%% cover info %%%%%%%%%%%%%%%
@@
 
-\institute{University of Wisconsin - Madison, Department of Nuclear Engineering and Engineering Physics, Madison, WI 53706}
+\institute{University of Wisconsin - Madison, Department of Nuclear Engineering and Engineering Physics}
+\instituteaddress{1415 Engineering Drive, Madison, WI 53706}
+
 \submitter{Robert W. Carlsen}
@@
 \submitteremail{rwcarlsen@gmail.com}
+\submitterphone{208/705-0291}
 
@@
 \usepackage{float} % used to help position figures with the H modifier
+\usepackage[modulo]{lineno}
 
@@
 
+%lineno\begin{linenumbers}
+
 \section{Introduction}
@@
 \bibliography{refs}
+
+%lineno\end{linenumbers}
+
 \end{document}

### 2016-05-04 paper.tex
mods to format/style for arxiv.org submission

@@
 %%%%%%%% cover info %%%%%%%%%%%%%%%
-\title{Challenging Fuel Cycle Modeling Assumptions: Facility and Time Step Discretization Effects}
-\author{Robert~W.~Carlsen, Paul~P.H.~Wilson}
-
-\institute{University of Wisconsin - Madison, Department of Nuclear Engineering and Engineering Physics}
-\instituteaddress{1415 Engineering Drive, Madison, WI 53706}
+\title{Challenging Fuel Cycle Modeling Assumptions:\\Facility and Time Step Discretization Effects}
+\author{Robert~W.~Carlsen (rwcarlsen@gmail.com), Paul~P.H.~Wilson}
+
+\institute{University of Wisconsin -- Madison, Department of Nuclear Engineering and Engineering Physics, Madison, WI 53706}
+\instituteaddress{x}
 
@@
 \myabstract{
-Due to the diversity of fuel cycle simulator modeling assumptions, direct
+\noindent Due to the diversity of fuel cycle simulator modeling assumptions, direct
 comparison and benchmarking can be difficult.  In 2012 the Organisation for
